---
title: "Chapter 4 - Cross-species iNKT"
author: "Salomé Carcy"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true

---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 5, fig.height = 5,
                      warning=FALSE, message=FALSE,
                      root.dir = "~/Projects/phd/")
```


# IMPORT

## Import librairies
```{r import-librairies}
library(ggplot2)
library(RColorBrewer)
library(cowplot)
library(tidyverse)
library(dplyr)
library(Seurat)
library(SCpubr)
library(patchwork)
library(SummarizedExperiment)
library(MetaNeighbor)
library(gplots)
library(reshape2)

source("~/Projects/phd/scripts/colors_universal.R")
```

## Import data
```{r import-data}
# Seurat objects
seur_hu_thym <- list(
  "nkt"=readRDS("./data/seurat_objects/thymus_objects/seurat_thymus_nkt_with_gene_signatures.rds"),
  "mait"=readRDS("./data/seurat_objects/thymus_objects/seurat_thymus_mait_with_gene_signatures.rds"),
  "gdt"=readRDS("./data/seurat_objects/thymus_objects/seurat_thymus_gdt_with_gene_signatures.rds")
)

seur_ms_thym <- list(
  "nkt"=readRDS("./data/seurat_objects/harshakrovi2020/ms_nkt_seurobj.rds"),
  "mait"=readRDS("./data/seurat_objects/legoux2019/ms_mait_seurobj.rds"),
  "gdt"=readRDS("./data/seurat_objects/lee2020/ms_gdt_seurobj_lee.rds")
)
seur_ms_thym$nkt$cell_type <- factor(seur_ms_thym$nkt$cell_type, levels=c(
  "Stage0",
  "iNKTp",
  "iNKT2",
  "iNKT17",
  "iNKT1"
))

# Orthologs table
ortholog.df <- read.csv("~/Projects/phd/data/cross_species/ortholog_table_ms_hu_full.csv")
```




# FUNCTIONS
```{r define-functions}
lineage_genesignature <- function(seuratobj, genesignature, file_name="no"){
  p <- ggrastr::rasterise(
    do_FeaturePlot(
    seuratobj,
    features = genesignature,
    min.cutoff=0,
    use_viridis = T,
    viridis.palette = "E",
    legend.position="right",
    order=T,
    pt.size = 1.5,
    raster=F
    # raster.dpi=2048,
    # pt.size=10
  ),
  layers="Point",
  dpi=300
  )
  print(p)
  if(file_name != "no"){
    ggsave(paste0("~/Projects/phd/data/figures/chapter_04/figs_unfinished/section1/", file_name),
           plot=p,
           width=9, height=6)
  }
}


plot_DotPlot <- function(seurobj, group, features, scaling=T){
  # get plot data
  p <- Seurat::DotPlot(
    seurobj,
    group.by=group,
    features=features,
    scale=scaling
  )
  # plot in personalized way
  p <- ggplot(p$data, aes(x=id, y=features.plot, fill=avg.exp.scaled, size=pct.exp))+
    geom_point(color="black", shape=21)+
    # scale_fill_gradient2(low=scales::muted("blue"), high=scales::muted("red"), name="z-score\nnormalized\navg expression")+
    scale_size_continuous(range=c(0,6), limits=c(0,100), name="%cells\nexpressing\ngene")+
    theme_bw()+
    theme(axis.text.y=element_text(face="italic"),
          axis.text.x=element_text(angle=45, hjust=1))+
    labs(y="", x="")
  # different color scale if scaled or not
  if(scaling==T){
    p <- p + scale_fill_gradient2(low=scales::muted("blue"), high=scales::muted("red"), name="z-score\nnormalized\navg expression")
  } else{
    p <- p + viridis::scale_fill_viridis(option="B", direction=-1, name="normalized\navg expression")
  }
  return(p)
}

plot_mtn <-
  function(mtn_output,
           var1 = "human",
           var2 = "mouse",
           var1_nbcells,
           var2_nbcells,
           var1_order,
           var2_order,
           bp_ymax = 20,
           file_name="no",
           figsize=c(12, 12)
           ) {
    
  # REMODEL DATAFRAME
  mtn.df <- melt(mtn_output) %>%
    filter(str_detect(Var1,var1)) %>%
    mutate(Var1 = gsub(paste0(var1, "\\|"), "", Var1)) %>%
    filter(str_detect(Var2, var2)) %>%
    mutate(Var2 = gsub(paste0(var2, "\\|"), "", Var2)) %>%
    as_tibble() %>%
    dplyr::rename(auroc=value) %>%
    # add nb of cells for var1 and var2
    left_join(var1_nbcells, by="Var1") %>%
    left_join(var2_nbcells, by="Var2") %>%
    # transform var1 and var2 into factors for ordering
    mutate(Var1=factor(Var1, levels=var1_order),
           Var2=factor(Var2, levels=var2_order))
  
    
    # PROPORTION OF VAR1 THYMOCYTES IN EACH CLUSTER
    bp.x <-
      ggplot(
        data = mtn.df %>% select(Var1, propcells_var1) %>% distinct(),
        aes(x = Var1, y = propcells_var1)
      ) +
      geom_bar(stat = "identity", fill = "#bdbdbd") + theme_cowplot() +
      scale_x_discrete(position = "top") +
      scale_y_continuous(limits = c(0, bp_ymax), breaks = seq(0,bp_ymax, length.out=3)) +
      labs(y = "%cells") +
      theme(
        axis.text.x = element_text(angle = 45, hjust = 0),
        axis.ticks.y = element_blank(),
        axis.title.x = element_blank(),
        axis.line.x = element_blank(),
        legend.position = "none"
      )
    
    # PROPORTION OF VAR2 THYMOCYTES IN EACH CLUSTER
    bp.y <-
      ggplot(
        data = mtn.df %>% select(Var2, propcells_var2) %>% distinct(),
        aes(x = Var2, y = propcells_var2)
      ) +
      geom_bar(stat = "identity", fill = "#bdbdbd") +
      scale_x_discrete(position = "top") +
      scale_y_continuous(limits = c(0, bp_ymax), breaks = seq(0,bp_ymax, length.out=3)) +
      labs(y = "%cells") + coord_flip() + theme_cowplot() +
      theme(
        axis.title.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.line.y = element_blank(),
        legend.position = "none"
      )
    
    # BUBBLEPLOT
    hm.clean <-
      ggplot(mtn.df, aes(x = Var1, y = Var2)) +
      geom_point(aes(size = abs(auroc-0.5), fill = auroc), color="black", shape=21) +
      geom_text(
        data = mtn.df %>% filter(auroc > 0.8) %>% mutate(across("auroc", \(x) round(x, 2))),
        aes(label = auroc),
        color = "white"
      ) +
      scale_size_continuous(
        limits = c(0, 0.5),
        breaks = seq(0, 0.5, by = 0.1),
        range = c(1, 15)
      ) +
      scale_fill_gradient2(
        low = "#2166ac",
        mid = "white",
        high = "#a50f15",
        midpoint = 0.5,
        limits = c(0, 1),
        name = "AUROC",
        breaks = seq(0, 1, by = 0.2)
      ) +
      labs(x = paste(var1, "clusters"), y = paste(var2, "clusters"), size = "AUROC") +
      theme_cowplot() +
      theme(
        legend.position = "bottom",
        legend.key.width = unit(0.8, 'cm'),
        axis.text.x = element_text(angle = 45, hjust = 1)
      )
    
    # COMBINE
    p <- (bp.x+plot_spacer() + plot_layout(widths = c(5, 1))) / (hm.clean + bp.y + plot_layout(widths = c(5, 1))) + plot_layout(heights = c(1, 5))
    if(file_name!="no"){ggsave(paste0("./data/figures/chapter_04/figs_unfinished/section1/", file_name), plot=p, width=figsize[1], height=figsize[2])}
    return(p)
  }

# Get nb of cells in human and mouse clusters
nbcells_hu <- as.data.frame(table(seur_hu_thym$nkt$clusters_per_lineage)) %>%
    dplyr::rename(ncells_var1=Freq) %>%
    mutate(propcells_var1 = ncells_var1*100/dim(seur_hu_thym$nkt)[2]) 
nbcells_ms <- as.data.frame(table(seur_ms_thym$nkt$cell_type)) %>%
    dplyr::rename(Var2=Var1, ncells_var2=Freq) %>%
    mutate(propcells_var2 = ncells_var2*100/dim(seur_ms_thym$nkt)[2])

# Define order of mouse and human clusters
hu_levels <- paste0("iNKT_thymus_c", 0:6)
ms_levels <- rev(c(
  "Stage0",
  "iNKTp",
  "iNKT2",
  "iNKT17",
  "iNKT1"
))
```




# iNKT FIGURE

## iNKT UMAPs
```{r inkt-umap-mouse}
ggrastr::rasterise(
  do_DimPlot(
    seur_ms_thym$nkt,
    group.by = "cell_type",
    colors.use = cols_harsha_ms,
    legend.position = "right"
  ),
  layers = "Point",
  dpi = 300
) +
  labs(title = "Mouse") +
  theme(
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    axis.line = element_blank(),
    panel.border = element_rect(
      colour = "black",
      fill = NA,
      size = 1
    ),
    axis.title = element_text(hjust = 0)
  )
# ggsave("./data/figures/chapter_04/figs_unfinished/section1/ch4_fig3_inkt_umap_mouse.pdf", width=6, height=5)
```

```{r inkt-umap-human}
ggrastr::rasterise(
  do_DimPlot(
    seur_hu_thym$nkt,
    group.by="clusters_per_lineage",
    colors.use=cols_thym_nkt,
    legend.position="right"
    ),
  layers="Point",
  dpi=300
  ) +
  labs(title="Human")+
  theme(
    axis.ticks=element_blank(),
    axis.text=element_blank(),
    axis.line = element_blank(),
    panel.border = element_rect(colour = "black", fill=NA, size=1),
    axis.title=element_text(hjust=0)
    )
# ggsave("./data/figures/chapter_04/figs_unfinished/section1/ch4_fig3_inkt_umap_human.pdf", width=6.5, height=5)
```


## iNKT stages DotPlot
```{r inkt-stages-dotplot, fig.height=6}
# Mouse and Human
plot_DotPlot(
  seurobj = seur_ms_thym$nkt,
  group = "cell_type",
  features = c("Cd4", "Cd8a", "Cd8b1", "Cd24a", "Cd44", "Klrb1c"),
  scaling = F
) + ggtitle("Mouse") +
  plot_DotPlot(
    seurobj = seur_hu_thym$nkt,
    group = "clusters_per_lineage",
    features = c("CD4", "CD8A", "CD8B", "CD24", "CD44", "KLRB1"),
    scaling = F
  ) + ggtitle("Human")
# ggsave("./data/figures/chapter_04/figs_unfinished/section1/ch4_fig3_inkt_stages_dotplot.pdf", width=10, height=4.5)
```


## iNKT functional stages DotPlot
Here we select genes that define each mouse cluster based on [Harsha Krovi et al. 2020](https://www.nature.com/articles/s41467-020-20073-8) and [Baranek et al.](https://www.sciencedirect.com/science/article/pii/S2211124720311050). These are not "marker genes" _per se_, as we didn't obtain them by computing the most DE genes in each cluster. Instead we selected the genes highlighted in the literature as defining each iNKT developmental stage.
```{r inkt-marker-genes1}
marker_genes_inkt_ms <- c(
  # "Stat1",
  # "Stat4",
  # "Stat6",
  # "Stat3",
  # "Il2rb",
  # "Il15ra",
  # "Il7r",
  # iNKT0
  "Cd24a",
  "Cd69",
  "Egr2",
  "Hivep3",
  # iNKTp
  "Cdk1",
  "Mki67",
  # Immature & iNKT2
  "Ccr9",
  "Lef1",
  "Slamf6",
  "Gata3",
  "Il4",
  "Ccr7",
  # "Plac8",
  # "Izumo1r",
  "Il6ra",
  # Maturation
  "Zbtb16",
  "Cd44",
  # iNKT17
  "Rorc",
  "Rora",
  "Il17re",
  "Ccr6",
  "Tmem176a",
  # iNKT1
  "Tbx21",
  "Eomes",
  "Klrb1c",
  "Ifng",
  "Prf1",
  "Xcl1",
  # "Il2rb",
  "Slamf7",
  # Exit
  "Klf2",
  "S1pr1"
)

marker_genes_inkt_hu <- c(
  # "STAT1",
  # "STAT4",
  # "STAT6",
  # "STAT3",
  # "IL2RB",
  # "IL15RA",
  # "IL7R",
  # iNKT0
  "CD24",
  "CD69",
  "EGR2",
  "HIVEP3",
  # iNKTp
  "CDK1",
  "MKI67",
  # Immature & iNKT2
  "CCR9",
  "LEF1",
  "SLAMF6",
  "GATA3",
  "ID3", # placeholder because IL4 was not found
  "CCR7",
  "IL6R",
  # Maturation
  "ZBTB16",
  "CD44",
  # iNKT17
  "RORC",
  "RORA",
  "IL17RE",
  "CCR6",
  "IL23R", # placeholder because TMEM176A was not found
  # iNKT1
  "TBX21",
  "EOMES",
  "KLRB1",
  "IFNG",
  "PRF1",
  "XCL1",
  "SLAMF7",
  # Exit
  "KLF2",
  "S1PR1"
)
```

```{r inkt-marker-genes2, fig.height=8, fig.width=8}
plot_DotPlot(
  seurobj = seur_ms_thym$nkt,
  group = "cell_type",
  features = marker_genes_inkt_ms,
  scaling = T
) + ggtitle("Mouse") +
  plot_DotPlot(
    seurobj = seur_hu_thym$nkt,
    group = "clusters_per_lineage",
    features = marker_genes_inkt_hu,
    scaling = T
  ) + ggtitle("Human")
# ggsave("./data/figures/chapter_04/figs_unfinished/section1/ch4_fig3_inkt_stages_bigdotplot.pdf", width=8, height=7)
```



## iNKT gene signature
The iNKT gene signatures are provided in supplementary table S1 from the [Harsha Krovi et al. 2020](https://www.nature.com/articles/s41467-020-20073-8) paper. Let's first prepare the gene signatures, in mouse and human. Not all mouse genes may have human orthologs, so we want to make sure to score the signatures on the same lists of genes (removing any mouse genes that don't have any human ortholog).
```{r inkt-gene-signatures-prep}
# get mouse iNKT signature genes
harshakrovi_supptable <- read.csv("./data/litterature_gene_signatures/harshakrovi_tableS1.csv", skip=1)[,2:8]
harshakrovi_supptable <- harshakrovi_supptable %>%
  # keep only iNKT1, 2, 17 clusters (c4-10)
  filter(cluster %in% 4:10) %>%
  filter(p_val_adj < 0.05) %>% # they are all already below 0.05
  # cluster to iNKT subset
  mutate(inkt=case_when(
    cluster==4 ~ "iNKT2",
    cluster==5 ~ "iNKT17",
    cluster %in% 6:10 ~ "iNKT1"
    )) %>%
  # get DE genes per iNKT subset
  select(gene, inkt) %>%
  distinct()
# write_csv(harshakrovi_supptable, "~/Downloads/inkt_signatures.csv")

# make it into a list
harshakrovi_gene_signature_list <- lapply(split(harshakrovi_supptable, harshakrovi_supptable$inkt), function(x) x$gene)
names(harshakrovi_gene_signature_list) <- paste0("harshakrovi_", names(harshakrovi_gene_signature_list))
harshakrovi_gene_signature_list <- lapply(harshakrovi_gene_signature_list, function(x) x[x!=""]) # remove empty genes
harshakrovi_gene_signature_list <- lapply(harshakrovi_gene_signature_list, function(x) x[x %in% rownames(seur_ms_thym$nkt)]) # remove genes not found in our data
lapply(harshakrovi_gene_signature_list, function(x) length(x)) # sanity check (0 genes per signature lost)

# check how many genes in gene signature we can find in ortholog table
lapply(harshakrovi_gene_signature_list, function(x) table(x %in% ortholog.df$ms_symbol, useNA="ifany"))

# More than 80% of genes in gene signatures can be found in ortholog table, so let's translate them
nkt_signature_list_humangenes <- lapply(
  harshakrovi_gene_signature_list,
  function(x) ortholog.df %>% filter(ms_symbol %in% x & hu_symbol %in% rownames(seur_hu_thym$nkt)) %>% pull(hu_symbol)
  ) # remove genes not found in our data
lapply(nkt_signature_list_humangenes, function(x) length(x)) # sanity check (a few genes per signature lost, but still >10 genes per signature left)

# Now update the mouse signatures to keep the same genes as the ones in human signatures
nkt_signature_list_mousegenes <- lapply(
  harshakrovi_gene_signature_list, 
  function(x) ortholog.df %>% filter(ms_symbol %in% x & hu_symbol %in% rownames(seur_hu_thym$nkt)) %>% pull(ms_symbol)
  )
lapply(nkt_signature_list_mousegenes, function(x) length(x))
```

Now let's score these signatures and plot!
```{r inkt-gene-signatures-mouse, fig.height=5, fig.width=8}
# Score
seur_ms_thym$nkt <- AddModuleScore(
  object = seur_ms_thym$nkt,
  assay = "RNA",
  features = nkt_signature_list_mousegenes,
  name=names(nkt_signature_list_mousegenes),
  seed=1
  )
colnames(seur_ms_thym$nkt@meta.data)[8:10] <- names(nkt_signature_list_mousegenes)

# Plot and save
lineage_genesignature(
  seuratobj = seur_ms_thym$nkt,
  # file_name = "ch4_fig3_inkt_signatureNKT1_mouse.pdf",
  genesignature = "harshakrovi_iNKT1"
  )
lineage_genesignature(
  seuratobj = seur_ms_thym$nkt,
  # file_name = "ch4_fig3_inkt_signatureNKT2_mouse.pdf",
  genesignature = "harshakrovi_iNKT2"
  )
lineage_genesignature(
  seuratobj = seur_ms_thym$nkt,
  # file_name = "ch4_fig3_inkt_signatureNKT17_mouse.pdf",
  genesignature = "harshakrovi_iNKT17"
)
```

```{r inkt-gene-signatures-human, fig.height=5, fig.width=8}
seur_hu_thym$nkt <- AddModuleScore(
  object = seur_hu_thym$nkt,
  assay = "RNA",
  features = nkt_signature_list_humangenes,
  name=names(nkt_signature_list_humangenes),
  seed=1
  )
colnames(seur_hu_thym$nkt@meta.data)[49:51] <- names(nkt_signature_list_humangenes)

# Plot and save
lineage_genesignature(
  seuratobj = seur_hu_thym$nkt,
  genesignature = "harshakrovi_iNKT1",
  # file_name = "ch4_fig3_inkt_signatureNKT1_human.pdf"
)
lineage_genesignature(
  seuratobj = seur_hu_thym$nkt,
  genesignature = "harshakrovi_iNKT2",
  # file_name = "ch4_fig3_inkt_signatureNKT2_human.pdf"
)
lineage_genesignature(
  seuratobj = seur_hu_thym$nkt,
  genesignature = "harshakrovi_iNKT17",
  # file_name = "ch4_fig3_inkt_signatureNKT17_human.pdf"
)
```





# iNKT METANEIGHBOR

## Prepare data for metaneighbor
```{r nkt-prep-metaneighbor}
# Get the ortholog table with 1:1 orthologs only and orthology confidence of 1
ortholog_df_1to1 <- read.csv("~/Projects/phd/data/cross_species/ortholog_table_ms_hu_one2one.csv")

# Get counts, HVGs and metadata
ms.hvg <- VariableFeatures(FindVariableFeatures(seur_ms_thym$nkt, nfeatures=5000))
ms.counts <- seur_ms_thym$nkt[["RNA"]]@counts
ms.metadata <- seur_ms_thym$nkt@meta.data

hu.hvg <- VariableFeatures(FindVariableFeatures(seur_hu_thym$nkt, nfeatures=5000))
hu.counts <- seur_hu_thym$nkt[["RNA"]]@counts
hu.metadata <- seur_hu_thym$nkt@meta.data


# First, check whether genes can all be found in the ortholog table (many can't be found because I removed genes with orthology confidence=0)
table(unique(rownames(ms.counts)) %in% ortholog_df_1to1$ms_symbol) # 16,033 not
table(unique(rownames(hu.counts)) %in% ortholog_df_1to1$hu_symbol) # 5,238 not


# Subset the ortholog table to only genes that we can "translate"
dictionary <- ortholog_df_1to1 %>%
  as_tibble() %>%
  ## Intersection
  filter(ms_symbol %in% unique(rownames(ms.counts)) & hu_symbol %in% unique(rownames(hu.counts))) %>%
  ## Remove any symbols that are NAs
  filter(!is.na(ms_symbol)) %>%
  filter(!is.na(hu_symbol)) # %>%
  # ## Keep only 1:1 orthologs
  # group_by(ms_symbol) %>% filter(n_distinct(hu_symbol) == 1) %>% ungroup() %>%
  # group_by(hu_symbol) %>% filter(n_distinct(ms_symbol) == 1) %>% ungroup()
  

# Translate the mouse HVGs into "human gene" language
ms.hvg.translated <- pull(dictionary %>% filter(ms_symbol %in% ms.hvg), # not all ms HVGs are found in ortholog.df
                          hu_symbol)
hu.hvg.translated <- pull(dictionary %>% filter(hu_symbol %in% hu.hvg), # not all hu HVGs are found in ortholog.df
                          hu_symbol)
total.hvg <- unique(union(ms.hvg.translated, hu.hvg.translated))
length(total.hvg) # 5,437 genes


# Keep only ms and hu genes that have 1:1 orthologs
table(unique(rownames(ms.counts)) %in% dictionary$ms_symbol) # 11,799 genes should have a translation
table(unique(rownames(hu.counts)) %in% dictionary$hu_symbol) # 11,799 genes should have a translation
ms.counts <- ms.counts[rownames(ms.counts) %in% dictionary$ms_symbol,]
hu.counts <- hu.counts[rownames(hu.counts) %in% dictionary$hu_symbol,]


# Translate the mouse genes in count table into "human gene"
ms.dict <- dictionary %>%
  filter(ms_symbol %in% rownames(ms.counts)) %>%
  select(ms_symbol, hu_symbol) %>%
  # distinct() %>%
  # group_by(ms_symbol_data) %>% filter(n_distinct(hu_symbol)>1)
  distinct(ms_symbol, .keep_all=T)
ms.dict <- ms.dict[match(rownames(ms.counts), ms.dict$ms_symbol),]
table(ms.dict$ms_symbol == rownames(ms.counts)) # all true
table(is.na(ms.dict$hu_symbol)) # no NAs
# Translate
rownames(ms.counts) <- ms.dict$hu_symbol


# Merge everything into one
ms.metadata$study <- "Mouse"
ms.metadata <- ms.metadata[,c("cell_type", "study")]
colnames(ms.metadata)[1] <- "clusters_NKT"
head(ms.metadata)

hu.metadata$study <- "Human"
hu.metadata <- hu.metadata[,c("clusters_per_lineage", "study")]
colnames(hu.metadata)[1] <- "clusters_NKT"
head(hu.metadata)

ms.counts <- ms.counts[rownames(hu.counts),]
table(rownames(ms.counts)==rownames(hu.counts), useNA="ifany")
table(colnames(ms.metadata)==colnames(hu.metadata), useNA="ifany")
tot.counts   <- cbind(ms.counts, hu.counts)
tot.metadata <- rbind(ms.metadata, hu.metadata)
table(colnames(tot.counts)==rownames(tot.metadata), useNA="ifany")

# Convert seurat count matrix to SummarizedExperiment object
count <- SummarizedExperiment(assays=tot.counts,
                              colData=tot.metadata)
```


## Run metaneighbor
```{r nkt-run-metaneighbor}
# Run metaneighbor
mtn <- MetaNeighborUS(var_genes=total.hvg,
                      dat=count,
                      study_id=count$study,
                      cell_type=count$clusters_NKT,
                      fast_version=F) # diagonal looking really weird in slow version

# Plot heatmap
# pdf(
#   file = "./data/figures/chapter_04/figs_unfinished/suppfigures/nkt_crosspecies_metaneighbor_heatmap.pdf",
#   width = 5.5,
#   height = 5.5
#   )
heatmap.2(
  mtn,
  trace = "none",
  # superimpose a density histogram on color key
  density.info = "none",
  # color scale
  col = rev(colorRampPalette(brewer.pal(11, "RdYlBu"))(100)),
  breaks = seq(0, 1, length = 101),
  key.xlab = "AUROC",
  key.title = "",
  # text labels
  main = "Mouse vs Human NKT",
  cexRow = 0.6,
  cexCol = 0.6,
  # margins
  margins = c(7, 7)
)
# dev.off()
```


## Make nice figure out of metaneighbor
```{r nkt-plot-metaneighbor, fig.height=7, fig.width=7}
plot_mtn(
  mtn_output = mtn,
  var1 = "Human",
  var2 = "Mouse",
  var1_nbcells = nbcells_hu,
  var2_nbcells = nbcells_ms,
  var1_order = hu_levels,
  var2_order = ms_levels,
  bp_ymax = 60,
  # file_name = "ch4_fig3_inkt_metaneighbor.pdf",
  figsize = c(7, 8) # c(width, height)
)
```

Try to randomly sample cells from each cluster, to have equal number of cells per cluster.
```{r nkt-metaneighbor-subset-100sampling, fig.height=7, fig.width=7}
tot.metadata$clusters_NKT <- as.character(tot.metadata$clusters_NKT)

# Randomly sample 330 cells per mouse cluster
randomcells_list <- list()
mtn_list <- list()
for(i in 1:100){
  cat("\nRandom sampling #", i)
  randomcells_vector <- c()
  mtn_subset <- NULL
  
  # RANDOMLY SAMPLE CELLS
  for(cluster in unique(count$clusters_NKT)){
    # print(cluster)
    randomcells <- NULL
    
    # if it's a mouse cluster
    if(cluster %in% ms_levels){
      nbcells_to_sample <- 330
      allcells <- rownames(colData(count)[colData(count)$clusters_NKT==cluster,])
      # cat("Total nb of cells:", length(allcells), "\n")
      # set.seed(123)
      randomcells <- sample(allcells, nbcells_to_sample)
      # cat("Nb of sampled cells:", length(randomcells), "\n")
      # sanity check they all come from same cluster
      # cat("Cluster sampled cells come from:", unique(tot.metadata[randomcells, "clusters_NKT"]), "\n\n")
    }
    
    # if it's a human cluster
    else if(cluster %in% hu_levels){
      if(cluster %in% c("iNKT_thymus_c3", "iNKT_thymus_c4")){
        nbcells_to_sample <- 0
        randomcells <- rownames(colData(count)[colData(count)$clusters_NKT==cluster,])
        # cat("Nb of sampled cells:", length(randomcells), "\n")
        # cat("Cluster sampled cells come from:", unique(tot.metadata[randomcells, "clusters_NKT"]), "\n\n")
      }
      else{
        nbcells_to_sample <- 185
        allcells <- rownames(colData(count)[colData(count)$clusters_NKT==cluster,])
        # cat("Total nb of cells:", length(allcells), "\n")
        # set.seed(123)
        randomcells <- sample(allcells, nbcells_to_sample)
        # cat("Nb of sampled cells:", length(randomcells), "\n")
        # cat("Cluster sampled cells come from:", unique(tot.metadata[randomcells, "clusters_NKT"]), "\n\n")
      }
    }
    
    randomcells_vector <- c(randomcells_vector, randomcells)
    # print(length(randomcells_vector))
  }
  
  # Subset Summarized Experiment object to the randomly selected cells
  count_subset <- count[,randomcells_vector]
  table(colData(count_subset)$clusters_NKT)
  
  
  # re-run metaneighbor
  mtn_subset <- MetaNeighborUS(var_genes=total.hvg,
                               dat=count_subset,
                               study_id=count_subset$study,
                               cell_type=count_subset$clusters_NKT,
                               fast_version=T)
  
  # save
  randomcells_list[[i]] <- randomcells_vector
  mtn_list[[i]] <- mtn_subset
}

length(randomcells_list)
length(mtn_list)

# sanity check that different cells were sampled at each round
table(randomcells_list[[1]] %in% randomcells_list[[2]])

# get mean and sd of all the metaneighbor matrices
mtn_mean <- apply(simplify2array(mtn_list), 1:2, mean)
mtn_sd <- apply(simplify2array(mtn_list), 1:2, sd)

# check the distribution of SD (very low SD across random samplings, which is good!)
ggplot(melt(mtn_sd), aes(x=value))+
  geom_histogram(bins=200, color="black", fill="white")+
  scale_x_continuous(expand=c(0,0), limits=c(0,1))+
  scale_y_continuous(expand=c(0,0), limits=c(0,100))+
  labs(x="AUROC standard deviation", y="# AUROC values", title="Distribution of SD of AUROC values")+
  theme(panel.grid.minor=element_blank(),
        panel.grid.major=element_blank())
# ggsave("./data/figures/chapter_04/figs_unfinished/suppfigures/nkt_crosspecies_100randomcellsampling_sd_auroc_hist.pdf", width=7, height=6)
```

```{r nkt-metaneighbor-subset-100sampling-plot, fig.height=7, fig.width=7}
# pdf(
#   file = "./data/figures/chapter_04/figs_unfinished/suppfigures/nkt_crosspecies_100randomcellsampling_mean_auroc_heatmap.pdf",
#   width = 5.5,
#   height = 5.5
#   )
# heatmap
heatmap.2(
  mtn_mean,
  trace = "none",
  # superimpose a density histogram on color key
  density.info = "none",
  # color scale
  col = rev(colorRampPalette(brewer.pal(11, "RdYlBu"))(100)),
  breaks = seq(0, 1, length = 101),
  key.xlab = "AUROC",
  key.title = "",
  # text labels
  main = "Mouse vs Human NKT",
  cexRow = 0.6,
  cexCol = 0.6,
  # margins
  margins = c(7, 7)
)
# dev.off()

# nice bubbleplot
plot_mtn(
  mtn_output = mtn_mean,
  var1 = "Human",
  var2 = "Mouse",
  var1_nbcells = as.data.frame(table(colData(count_subset)[colData(count_subset)$study=="Human",])) %>%
    dplyr::rename(propcells_var1=Freq, Var1=clusters_NKT) %>%
    filter(propcells_var1 != 0),
  var2_nbcells = as.data.frame(table(colData(count_subset)[colData(count_subset)$study=="Mouse",])) %>%
    dplyr::rename(propcells_var2=Freq, Var2=clusters_NKT) %>%
    filter(propcells_var2 != 0),
  var1_order = hu_levels,
  var2_order = ms_levels,
  bp_ymax = 400,
  # file_name = "../suppfigures/nkt_crosspecies_100randomcellsampling_mean_auroc.pdf",
  figsize = c(7, 8) # c(width, height)
)
```





# SESSION INFO
```{r}
sessionInfo()
```


