---
title: "Chapter 4 - Cross-species MAIT"
author: "Salomé Carcy"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true

---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 5, fig.height = 5,
                      warning=FALSE, message=FALSE,
                      root.dir = "~/Projects/phd/")
```


# IMPORT

## Import librairies
```{r import-librairies}
library(ggplot2)
library(RColorBrewer)
library(cowplot)
library(tidyverse)
library(dplyr)
library(Seurat)
library(SCpubr)
library(patchwork)
library(SummarizedExperiment)
library(MetaNeighbor)
library(gplots)
library(reshape2)

source("~/Projects/phd/scripts/colors_universal.R")
```

## Import data
```{r import-data}
# Seurat objects
seur_hu_thym <- list(
  "nkt"=readRDS("./data/seurat_objects/thymus_objects/seurat_thymus_nkt_with_gene_signatures.rds"),
  "mait"=readRDS("./data/seurat_objects/thymus_objects/seurat_thymus_mait_with_gene_signatures.rds"),
  "gdt"=readRDS("./data/seurat_objects/thymus_objects/seurat_thymus_gdt_with_gene_signatures.rds")
)

seur_ms_thym <- list(
  "nkt"=readRDS("./data/seurat_objects/harshakrovi2020/ms_nkt_seurobj.rds"),
  "mait"=readRDS("./data/seurat_objects/legoux2019/ms_mait_seurobj.rds"),
  "gdt"=readRDS("./data/seurat_objects/lee2020/ms_gdt_seurobj_lee.rds")
)
# Set order of clusters for MAIT
seur_ms_thym$mait$cell_type <- factor(seur_ms_thym$mait$cell_type, levels=c(
  "MAIT0",
  "Cluster 7",
  "CyclingS",
  "CyclingG2M",
  "MAIT17a",
  "MAIT17b",
  "MAIT1"
))

# Orthologs table
ortholog.df <- read.csv("~/Projects/phd/data/cross_species/ortholog_table_ms_hu_full.csv")
```




# FUNCTIONS
```{r define-functions}
lineage_genesignature <- function(seuratobj, genesignature, file_name="no"){
  p <- ggrastr::rasterise(
    do_FeaturePlot(
    seuratobj,
    features = genesignature,
    min.cutoff=0,
    use_viridis = T,
    viridis.palette = "E",
    legend.position="right",
    order=T,
    pt.size = 1.5,
    raster=F
    # raster.dpi=2048,
    # pt.size=10
  ),
  layers="Point",
  dpi=300
  )
  print(p)
  if(file_name != "no"){
    ggsave(paste0("~/Projects/phd/data/figures/chapter_04/figs_unfinished/section1/", file_name),
           plot=p,
           width=9, height=6)
  }
}

plot_DotPlot <- function(seurobj, group, features, scaling=T){
  # get plot data
  p <- Seurat::DotPlot(
    seurobj,
    group.by=group,
    features=features,
    scale=scaling
  )
  # plot in personalized way
  p <- ggplot(p$data, aes(x=id, y=features.plot, fill=avg.exp.scaled, size=pct.exp))+
    geom_point(color="black", shape=21)+
    # scale_fill_gradient2(low=scales::muted("blue"), high=scales::muted("red"), name="z-score\nnormalized\navg expression")+
    scale_size_continuous(range=c(0,6), limits=c(0,100), name="%cells\nexpressing\ngene")+
    theme_bw()+
    theme(axis.text.y=element_text(face="italic"),
          axis.text.x=element_text(angle=45, hjust=1))+
    labs(y="", x="")
  # different color scale if scaled or not
  if(scaling==T){
    p <- p + scale_fill_gradient2(low=scales::muted("blue"), high=scales::muted("red"), name="z-score\nnormalized\navg expression")
  } else{
    p <- p + viridis::scale_fill_viridis(option="B", direction=-1, name="normalized\navg expression")
  }
  return(p)
}

plot_mtn <-
  function(mtn_output,
           var1 = "human",
           var2 = "mouse",
           var1_nbcells,
           var2_nbcells,
           var1_order,
           var2_order,
           bp_ymax = 20,
           file_name="no",
           figsize=12
           ) {
    
  # REMODEL DATAFRAME
  mtn.df <- melt(mtn_output) %>%
    filter(str_detect(Var1,var1)) %>%
    mutate(Var1 = gsub(paste0(var1, "\\|"), "", Var1)) %>%
    filter(str_detect(Var2, var2)) %>%
    mutate(Var2 = gsub(paste0(var2, "\\|"), "", Var2)) %>%
    as_tibble() %>%
    dplyr::rename(auroc=value) %>%
    # add nb of cells for var1 and var2
    left_join(var1_nbcells, by="Var1") %>%
    left_join(var2_nbcells, by="Var2") %>%
    # transform var1 and var2 into factors for ordering
    mutate(Var1=factor(Var1, levels=var1_order),
           Var2=factor(Var2, levels=var2_order))
  
    
    # PROPORTION OF VAR1 THYMOCYTES IN EACH CLUSTER
    bp.x <-
      ggplot(
        data = mtn.df %>% select(Var1, propcells_var1) %>% distinct(),
        aes(x = Var1, y = propcells_var1)
      ) +
      geom_bar(stat = "identity", fill = "#bdbdbd") + theme_cowplot() +
      scale_x_discrete(position = "top") +
      scale_y_continuous(limits = c(0, bp_ymax), breaks = seq(0,bp_ymax, length.out=3)) +
      labs(y = "%cells") +
      theme(
        axis.text.x = element_text(angle = 45, hjust = 0),
        axis.ticks.y = element_blank(),
        axis.title.x = element_blank(),
        axis.line.x = element_blank(),
        legend.position = "none"
      )
    
    # PROPORTION OF VAR2 THYMOCYTES IN EACH CLUSTER
    bp.y <-
      ggplot(
        data = mtn.df %>% select(Var2, propcells_var2) %>% distinct(),
        aes(x = Var2, y = propcells_var2)
      ) +
      geom_bar(stat = "identity", fill = "#bdbdbd") +
      scale_x_discrete(position = "top") +
      scale_y_continuous(limits = c(0, bp_ymax), breaks = seq(0,bp_ymax, length.out=3)) +
      labs(y = "%cells") + coord_flip() + theme_cowplot() +
      theme(
        axis.title.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.line.y = element_blank(),
        legend.position = "none"
      )
    
    # BUBBLEPLOT
    hm.clean <-
      ggplot(mtn.df, aes(x = Var1, y = Var2)) +
      geom_point(aes(size = abs(auroc-0.5), fill = auroc), color="black", shape=21) +
      geom_text(
        data = mtn.df %>% filter(auroc > 0.8) %>% mutate(across("auroc", \(x) round(x, 2))),
        aes(label = auroc),
        color = "white"
      ) +
      scale_size_continuous(
        limits = c(0, 0.5),
        breaks = seq(0, 0.5, by = 0.1),
        range = c(1, 15)
      ) +
      scale_fill_gradient2(
        low = "#2166ac",
        mid = "white",
        high = "#a50f15",
        midpoint = 0.5,
        limits = c(0, 1),
        name = "AUROC",
        breaks = seq(0, 1, by = 0.2)
      ) +
      labs(x = paste(var1, "clusters"), y = paste(var2, "clusters"), size = "AUROC") +
      theme_cowplot() +
      theme(
        legend.position = "bottom",
        legend.key.width = unit(0.8, 'cm'),
        axis.text.x = element_text(angle = 45, hjust = 1)
      )
    
    # COMBINE
    p <- (bp.x+plot_spacer() + plot_layout(widths = c(5, 1))) / (hm.clean + bp.y + plot_layout(widths = c(5, 1))) + plot_layout(heights = c(1, 5))
    if(file_name!="no"){ggsave(paste0("./data/figures/chapter_04/figs_unfinished/section1/", file_name), plot=p, width=figsize, height=figsize, device=cairo_pdf)}
    return(p)
  }

# Get nb of cells in human and mouse clusters
nbcells_hu <- as.data.frame(table(seur_hu_thym$mait$clusters_per_lineage)) %>%
    dplyr::rename(ncells_var1=Freq) %>%
    mutate(propcells_var1 = ncells_var1*100/dim(seur_hu_thym$mait)[2]) 
nbcells_ms <- as.data.frame(table(seur_ms_thym$mait$cell_type)) %>%
    dplyr::rename(Var2=Var1, ncells_var2=Freq) %>%
    mutate(propcells_var2 = ncells_var2*100/dim(seur_ms_thym$mait)[2])

# Define order of mouse and human clusters
hu_levels <- paste0("MAIT_thymus_c", 0:6)
ms_levels <- rev(c(
  "MAIT0",
  "Cluster 7",
  "CyclingS",
  "CyclingG2M",
  "MAIT17a",
  "MAIT17b",
  "MAIT1"
))
```




# MAIT FIGURE

## MAIT UMAPs
```{r mait-umap-mouse}
ggrastr::rasterise(
  do_DimPlot(
    seur_ms_thym$mait,
    group.by = "cell_type",
    colors.use = cols_legoux_ms,
    legend.position = "right"
  ),
  layers = "Point",
  dpi = 300
) +
  labs(title = "Mouse") +
  theme(
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    axis.line = element_blank(),
    panel.border = element_rect(
      colour = "black",
      fill = NA,
      size = 1
    ),
    axis.title = element_text(hjust = 0)
  )
# ggsave("./data/figures/chapter_04/figs_unfinished/section1/ch4_fig4_mait_umap_mouse.pdf", width=6, height=5)
```

```{r mait-umap-human}
ggrastr::rasterise(
  do_DimPlot(
    seur_hu_thym$mait,
    group.by = "clusters_per_lineage",
    colors.use = cols_thym_mait,
    legend.position = "right"
  ),
  layers = "Point",
  dpi = 300
) +
  labs(title = "Human") +
  theme(
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    axis.line = element_blank(),
    panel.border = element_rect(
      colour = "black",
      fill = NA,
      size = 1
    ),
    axis.title = element_text(hjust = 0)
  )
# ggsave("./data/figures/chapter_04/figs_unfinished/section1/ch4_fig4_mait_umap_human.pdf", width=6.5, height=5)
```


## MAIT stages DotPlot
```{r mait-stages-dotplot, fig.height=5, fig.width=10}
# Mouse and Human
plot_DotPlot(
  seurobj = seur_ms_thym$mait,
  group = "cell_type",
  features = c("Cd4", "Cd8a", "Cd8b1", "Cd24a", "Cd44", "Cd27", "Klrb1c", "Il18r1"),
  scaling = F
) + ggtitle("Mouse") +
  plot_DotPlot(
    seurobj = seur_hu_thym$mait,
    group = "clusters_per_lineage",
    features = c("CD4", "CD8A", "CD8B", "CD24", "CD44", "CD27", "KLRB1", "IL18R1"),
    scaling = F
  ) + ggtitle("Human")
# ggsave("./data/figures/chapter_04/figs_unfinished/section1/ch4_fig4_mait_stages_dotplot.pdf", width=10, height=4.5)
```


## MAIT functional stages DotPlot
Here we select genes that define each mouse cluster based on [Legoux et al. 2019](https://www.nature.com/articles/s41590-019-0465-3). These are not "marker genes" _per se_, as we didn't obtain them by computing the most DE genes in each cluster. Instead we selected the genes highlighted in the literature as defining each MAIT developmental stage (Legoux et al., Koay et al., etc.).
```{r mait-marker-genes1}
marker_genes_mait_ms <- c(
  # DP & positive selection (TCR induced genes)
  "Dntt",
  "Cd24a",
  "Cd69",
  "Egr2",
  "Hivep3",
  # Immature
  "Cd4",
  "Ccr9",
  "Sell",
  "Ccr7",
  "Lef1",
  "Slamf6",
  # Maturation
  "Zbtb16",
  "Cd44",
  # Prolif
  "Cdk1",
  "Mki67",
  # MAIT17
  "Rorc",
  "Rora",
  "Il17re",
  "Ccr6",
  # "Sdc1",
  # MAIT1
  "Tbx21",
  "Eomes",
  "Cxcr3",
  "Klrd1",
  "Ifng",
  "Prf1",
  "Slamf7",
  # "Ly6c1", "Ly6c2",
  # Exit
  "Klf2",
  "S1pr1"
)

marker_genes_mait_hu <- c(
  # DP & positive selection (TCR induced genes)
  "RAG1",
  "CD24",
  "CD69",
  "EGR2",
  "HIVEP3",
  # Immature
  "CD4",
  "CCR9",
  "SELL",
  "CCR7",
  "LEF1",
  "SLAMF6",
  # Maturation
  "ZBTB16",
  "CD44",
  # Prolif
  "CDK1",
  "MKI67",
  # MAIT17
  "RORC",
  "RORA",
  "IL17RE",
  "CCR6",
  # "SDC1",
  # MAIT1
  "TBX21",
  "EOMES",
  "CXCR3",
  "KLRD1",
  "IFNG",
  "PRF1",
  "SLAMF7",
  # "LY6H", "LY6S",
  # Exit
  "KLF2",
  "S1PR1"
)
```

```{r mait-marker-genes2, fig.height=8, fig.width=10}
# Mouse and Human
plot_DotPlot(
  seurobj = seur_ms_thym$mait,
  group = "cell_type",
  features = marker_genes_mait_ms,
  scaling = T
) + ggtitle("Mouse") +
  plot_DotPlot(
    seurobj = seur_hu_thym$mait,
    group = "clusters_per_lineage",
    features = marker_genes_mait_hu,
    scaling = T
  ) + ggtitle("Human")
# ggsave("./data/figures/chapter_04/figs_unfinished/section1/ch4_fig4_mait_stages_bigdotplot.pdf", width=10, height=7)
```



## MAIT gene signature
MAIT gene signatures came from supplementary table S1 of [Legoux et al. 2019](https://www.nature.com/articles/s41590-019-0465-3). Let's first prepare the gene signatures, in mouse and human. Not all mouse genes may have human orthologs, so we want to make sure to score the signatures on the same lists of genes (removing any mouse genes that don't have any human ortholog).
```{r mait-gene-signatures-prep}
# Get mouse gene signatures
legoux_gene_signature <- read.csv("./data/litterature_gene_signatures/mouse_tinn_signatures_combined.csv")[,1:2]
legoux_gene_signature_list <- as.list(legoux_gene_signature)
legoux_gene_signature_list <- lapply(legoux_gene_signature_list, function(x) x[x!=""]) # remove empty genes
legoux_gene_signature_list <- lapply(legoux_gene_signature_list, function(x) x[x %in% rownames(seur_ms_thym$mait)]) # remove genes not found in our data
lapply(legoux_gene_signature_list, function(x) length(x)) # sanity check (between 0-5 genes per signature lost)

# check how many genes in gene signature we can find in ortholog table
lapply(legoux_gene_signature_list, function(x) table(x %in% ortholog.df$ms_symbol, useNA="ifany"))

# More than 80% of genes in gene signatures can be found in ortholog table, so let's translate them
mait_signature_list_humangenes <- lapply(
  legoux_gene_signature_list,
  function(x) ortholog.df %>% filter(ms_symbol %in% x & hu_symbol %in% rownames(seur_hu_thym$mait)) %>% pull(hu_symbol)
  ) # remove genes not found in our data
lapply(mait_signature_list_humangenes, function(x) length(x)) # sanity check (a few genes per signature lost, but still >110 genes per signature left)

# Now update the mouse signatures to keep the same genes as the ones in human signatures
mait_signature_list_mousegenes <- lapply(
  legoux_gene_signature_list, 
  function(x) ortholog.df %>% filter(ms_symbol %in% x & hu_symbol %in% rownames(seur_hu_thym$mait)) %>% pull(ms_symbol)
  )
lapply(mait_signature_list_mousegenes, function(x) length(x))
```

Now let's score these signatures and plot!
```{r mait-gene-signatures-mouse, fig.height=5, fig.width=8}
seur_ms_thym$mait <- AddModuleScore(
  object = seur_ms_thym$mait,
  assay = "RNA",
  features = mait_signature_list_mousegenes,
  name=names(mait_signature_list_mousegenes),
  seed=1
  )
colnames(seur_ms_thym$mait@meta.data)[10:11] <- names(mait_signature_list_mousegenes)

# Plot and save
lineage_genesignature(
  seuratobj = seur_ms_thym$mait,
  # file_name = "ch4_fig4_mait_signatureMAIT1_mouse.pdf",
  genesignature = "legoux_mait1_thymus"
  )
lineage_genesignature(
  seuratobj = seur_ms_thym$mait,
  # file_name = "ch4_fig4_mait_signatureMAIT17_mouse.pdf",
  genesignature = "legoux_mait17_thymus"
)
```

```{r mait-gene-signatures-human, fig.height=5, fig.width=8}
seur_hu_thym$mait <- AddModuleScore(
  object = seur_hu_thym$mait,
  assay = "RNA",
  features = mait_signature_list_humangenes,
  name=names(mait_signature_list_humangenes),
  seed=1
  )
colnames(seur_hu_thym$mait@meta.data)[49:50] <- names(mait_signature_list_humangenes)

# Plot and save
lineage_genesignature(
  seuratobj = seur_hu_thym$mait,
  genesignature = "legoux_mait1_thymus",
  # file_name = "ch4_fig4_mait_signatureMAIT1_human.pdf"
)
lineage_genesignature(
  seuratobj = seur_hu_thym$mait,
  genesignature = "legoux_mait17_thymus",
  # file_name = "ch4_fig4_mait_signatureMAIT17_human.pdf"
)
```




# MAIT METANEIGHBOR

## Prepare data for metaneighbor
```{r mait-prep-metaneighbor}
# Get the ortholog table with 1:1 orthologs only and orthology confidence of 1
ortholog_df_1to1 <- read.csv("~/Projects/phd/data/cross_species/ortholog_table_ms_hu_one2one.csv")

# Get counts, HVGs and metadata
ms.hvg <- VariableFeatures(FindVariableFeatures(seur_ms_thym$mait, nfeatures=5000))
ms.counts <- seur_ms_thym$mait[["RNA"]]@counts
ms.metadata <- seur_ms_thym$mait@meta.data

hu.hvg <- VariableFeatures(FindVariableFeatures(seur_hu_thym$mait, nfeatures=5000))
# hu.hvg <- VariableFeatures(seur_hu_thym$mait)
hu.counts <- seur_hu_thym$mait[["RNA"]]@counts
hu.metadata <- seur_hu_thym$mait@meta.data


# First, check whether genes can all be found in the ortholog table (many can't be found because I removed genes with orthology confidence=0)
table(unique(rownames(ms.counts)) %in% ortholog_df_1to1$ms_symbol) # 1,781 not
table(unique(rownames(hu.counts)) %in% ortholog_df_1to1$hu_symbol) # 5,238 not


# Subset the ortholog table to only genes that we can "translate"
dictionary <- ortholog_df_1to1 %>%
  as_tibble() %>%
  ## Intersection
  filter(ms_symbol %in% unique(rownames(ms.counts)) & hu_symbol %in% unique(rownames(hu.counts))) %>%
  ## Remove any symbols that are NAs
  filter(!is.na(ms_symbol)) %>%
  filter(!is.na(hu_symbol))


# Translate the mouse HVGs into "human gene" language
ms.hvg.translated <- pull(dictionary %>% filter(ms_symbol %in% ms.hvg), # not all ms HVGs are found in ortholog.df
                          hu_symbol) # 3,919 ms HVG
hu.hvg.translated <- pull(dictionary %>% filter(hu_symbol %in% hu.hvg), # not all hu HVGs are found in ortholog.df
                          hu_symbol) # 2,919 hu HVG
total.hvg <- unique(union(ms.hvg.translated, hu.hvg.translated))
length(total.hvg) # 5,429 genes


# Keep only ms and hu genes that have 1:1 orthologs
table(unique(rownames(ms.counts)) %in% dictionary$ms_symbol) # 9,509 genes should have a translation
table(unique(rownames(hu.counts)) %in% dictionary$hu_symbol) # 9,509 genes should have a translation
ms.counts <- ms.counts[rownames(ms.counts) %in% dictionary$ms_symbol,]
hu.counts <- hu.counts[rownames(hu.counts) %in% dictionary$hu_symbol,]


# Translate the mouse genes in count table into "human gene"
ms.dict <- dictionary %>%
  filter(ms_symbol %in% rownames(ms.counts)) %>%
  select(ms_symbol, hu_symbol) %>%
  # distinct() %>%
  # group_by(ms_symbol_data) %>% filter(n_distinct(hu_symbol)>1)
  distinct(ms_symbol, .keep_all=T)
ms.dict <- ms.dict[match(rownames(ms.counts), ms.dict$ms_symbol),]
table(ms.dict$ms_symbol == rownames(ms.counts)) # all true
table(is.na(ms.dict$hu_symbol)) # no NAs
# Translate
rownames(ms.counts) <- ms.dict$hu_symbol
# table(rownames(ms.counts)%in%rownames(hu.counts)) # sanity check


# Merge everything into one
ms.metadata$study <- "Mouse"
ms.metadata <- ms.metadata[,c("cell_type", "study")]
colnames(ms.metadata)[1] <- "clusters_MAIT"
head(ms.metadata)

hu.metadata$study <- "Human"
hu.metadata <- hu.metadata[,c("clusters_per_lineage", "study")]
colnames(hu.metadata)[1] <- "clusters_MAIT"
head(hu.metadata)

ms.counts <- ms.counts[rownames(hu.counts),]
table(rownames(ms.counts)==rownames(hu.counts), useNA="ifany")
table(colnames(ms.metadata)==colnames(hu.metadata), useNA="ifany")
tot.counts   <- cbind(ms.counts, hu.counts)
tot.metadata <- rbind(ms.metadata, hu.metadata)
table(colnames(tot.counts)==rownames(tot.metadata), useNA="ifany")

# Convert seurat count matrix to SummarizedExperiment object
count <- SummarizedExperiment(assays=tot.counts,
                              colData=tot.metadata)

```


## Run metaneighbor
```{r mait-run-metaneighbor}
# Run metaneighbor
mtn <- MetaNeighborUS(var_genes=total.hvg,
                      dat=count,
                      study_id=count$study,
                      cell_type=count$clusters_MAIT,
                      fast_version=T) # diagonal looking really weird in slow version

# Plot heatmap
# pdf(
#   file = "./data/figures/chapter_04/figs_unfinished/suppfigures/mait_crosspecies_metaneighbor_heatmap.pdf",
#   width = 5.5,
#   height = 5.5
#   )
heatmap.2(
  mtn,
  trace = "none",
  # superimpose a density histogram on color key
  density.info = "none",
  # color scale
  col = rev(colorRampPalette(brewer.pal(11, "RdYlBu"))(100)),
  breaks = seq(0, 1, length = 101),
  key.xlab = "AUROC",
  key.title = "",
  # text labels
  main = "Mouse vs Human MAIT",
  cexRow = 0.6,
  cexCol = 0.6,
  # margins
  margins = c(7, 7)
)
# dev.off()
```


## Make nice figure out of metaneighbor
```{r mait-plot-metaneighbor, fig.height=7, fig.width=7}
plot_mtn(
  mtn_output = mtn,
  var1 = "Human",
  var2 = "Mouse",
  var1_nbcells = nbcells_hu,
  var2_nbcells = nbcells_ms,
  var1_order = hu_levels,
  var2_order = ms_levels,
  bp_ymax = 50,
  # file_name = "ch4_fig4_mait_metaneighbor.pdf",
  figsize = 8
)
```

The strong AUROC between the mouse "cycling" clusters and the human c1 and c3 clusters is peculiar. Let's re-run metaneighbor without cell cycle genes.
```{r mait-metaneighbor-wo-ccgenes, fig.height=7, fig.width=7}
# Run metaneighbor
mtn_wo_ccgenes <- MetaNeighborUS(var_genes=total.hvg[!total.hvg %in% unlist(cc.genes.updated.2019)],
                                 dat=count,
                                 study_id=count$study,
                                 cell_type=count$clusters_MAIT,
                                 fast_version=T) # diagonal looking really weird in slow version

# Plot heatmap
heatmap.2(
  mtn_wo_ccgenes,
  trace = "none",
  # superimpose a density histogram on color key
  density.info = "none",
  # color scale
  col = rev(colorRampPalette(brewer.pal(11, "RdYlBu"))(100)),
  breaks = seq(0, 1, length = 101),
  key.xlab = "AUROC",
  key.title = "",
  # text labels
  main = "Mouse vs Human MAIT",
  cexRow = 0.6,
  cexCol = 0.6,
  # margins
  margins = c(7, 7)
)

# Plot nicely
plot_mtn(
  mtn_output = mtn_wo_ccgenes,
  var1 = "Human",
  var2 = "Mouse",
  var1_nbcells = nbcells_hu,
  var2_nbcells = nbcells_ms,
  var1_order = hu_levels,
  var2_order = ms_levels,
  bp_ymax = 50,
  # file_name = "../suppfigures/mait_crosspecies_metaneighbor_wo_cellcycle_genes.pdf",
  figsize = 8
)
```







# SESSION INFO
```{r}
sessionInfo()
```


