---
title: "Chapter X - R Notebook"
author: "Salomé Carcy"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true

---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 5, fig.height = 5,
                      warning=FALSE, message=FALSE,
                      root.dir = "~/Projects/phd/")
```


# IMPORT

## Import librairies
```{r import-librairies}
library(ggplot2)
library(RColorBrewer)
library(cowplot)
library(tidyverse)
library(dplyr)
library(Seurat)
library(SCpubr)
library(patchwork)

source("~/Projects/phd/scripts/colors_universal.R")
```

## Import data
```{r import-data}
# Seurat objects
seur_hu_thym <- list(
  "nkt"=readRDS("./data/seurat_objects/thymus_objects/seurat_thymus_nkt_with_gene_signatures.rds"),
  "mait"=readRDS("./data/seurat_objects/thymus_objects/seurat_thymus_mait_with_gene_signatures.rds"),
  "gdt"=readRDS("./data/seurat_objects/thymus_objects/seurat_thymus_gdt_with_gene_signatures.rds")
)

seur_ms_thym <- list(
  "nkt"=readRDS("./data/seurat_objects/harshakrovi2020/ms_nkt_seurobj.rds"),
  "mait"=readRDS("./data/seurat_objects/legoux2019/ms_mait_seurobj.rds"),
  "gdt"=readRDS("./data/seurat_objects/lee2020/ms_gdt_seurobj_lee.rds")
)

# Orthologs table
ortholog.df <- read.csv("~/Projects/phd/data/cross_species/ortholog_table_ms_hu_one2one.csv")
```




# FUNCTIONS
```{r define-functions}
lineage_genesignature <- function(seuratobj, genesignature, file_name="no"){
  p <- ggrastr::rasterise(
    do_FeaturePlot(
    seuratobj,
    features = genesignature,
    min.cutoff=0,
    use_viridis = T,
    viridis.palette = "E",
    legend.position="right",
    order=T,
    pt.size = 1.5,
    raster=F
    # raster.dpi=2048,
    # pt.size=10
  ),
  layers="Point",
  dpi=300
  )
  if(file_name != "no"){
    ggsave(paste0("~/Projects/phd/data/figures/chapter_04/figs_unfinished/section1/", file_name),
           plot=p,
           width=9, height=6)
  }
}
```




# MAIT FIGURE

## MAIT UMAPs
```{r mait-umap-mouse}
seur_ms_thym$mait$cell_type <- factor(seur_ms_thym$mait$cell_type, levels=c(
  "MAIT0",
  "Cluster 7",
  "CyclingS",
  "CyclingG2M",
  "MAIT17a",
  "MAIT17b",
  "MAIT1"
))

ggrastr::rasterise(
  DimPlot(seur_ms_thym$mait, group.by="cell_type", cols=cols_legoux_ms),
  layers="Point",
  dpi=300
  ) +
  labs(title="Mouse")+
  theme(
    axis.ticks=element_blank(),
    axis.text=element_blank(),
    axis.line = element_blank(),
    panel.border = element_rect(colour = "black", fill=NA, size=1),
    axis.title=element_text(hjust=0)
    )
# ggsave("./data/figures/chapter_04/figs_unfinished/section1/ch4_fig3_mait_umap_mouse.pdf", width=6, height=5)
```

```{r mait-umap-human}
ggrastr::rasterise(
  DimPlot(seur_hu_thym$mait, group.by="clusters_per_lineage", cols=cols_thym_mait),
  layers="Point",
  dpi=300
  ) +
  labs(title="Human")+
  theme(
    axis.ticks=element_blank(),
    axis.text=element_blank(),
    axis.line = element_blank(),
    panel.border = element_rect(colour = "black", fill=NA, size=1),
    axis.title=element_text(hjust=0)
    )
# ggsave("./data/figures/chapter_04/figs_unfinished/section1/ch4_fig3_mait_umap_human.pdf", width=6.5, height=5)
```


## MAIT stages DotPlot
```{r mait-stages-dotplot}
# Mouse
do_DotPlot(
  seur_ms_thym$mait,
  group.by="cell_type",
  features=c("Cd4", "Cd8a", "Cd8b1", "Cd24a", "Cd44", "Cd27", "Klrb1", "Il18r1"),
  flip=T,
  use_viridis = T,
  viridis.palette="B"
)+
  theme(axis.text.y=element_text(face="italic"))
# ggsave("./data/figures/chapter_04/figs_unfinished/section1/ch4_fig4_mait_stages_dotplot_mouse.pdf", width=4, height=5)

# Human
do_DotPlot(
  seur_hu_thym$mait,
  group.by="clusters_per_lineage",
  features=c("CD4", "CD8A", "CD8B", "CD24", "CD44", "CD27", "KLRB1", "IL18R1"),
  flip=T,
  use_viridis = T,
  viridis.palette="B"
)+
  theme(axis.text.y=element_text(face="italic"))
# ggsave("./data/figures/chapter_04/figs_unfinished/section1/ch4_fig4_mait_stages_dotplot_human.pdf", width=4, height=5)
```


## MAIT functional stages DotPlot
Here we select genes that define each mouse cluster based on [Legoux et al. 2019](https://www.nature.com/articles/s41590-019-0465-3). These are not "marker genes" _per se_, as we didn't obtain them by computing the most DE genes in each cluster. Instead we selected the genes highlighted in the literature as defining each MAIT developmental stage (Legoux et al., Koay et al., etc.).
```{r mait-marker-genes1}
marker_genes_mait_ms <- c(
  # DP & positive selection (TCR induced genes)
  "Dntt",
  "Cd24a",
  "Cd69",
  "Egr2",
  "Hivep3",
  # Immature
  "Cd4",
  "Ccr9",
  "Sell",
  "Ccr7",
  "Lef1",
  "Slamf6",
  # Maturation
  "Zbtb16",
  "Cd44",
  # Prolif
  "Cdk1",
  "Mki67",
  # MAIT17
  "Rorc",
  "Rora",
  "Il17re",
  "Ccr6",
  # "Sdc1",
  # MAIT1
  "Tbx21",
  "Eomes",
  "Cxcr3",
  "Klrd1",
  "Ifng",
  "Prf1",
  "Slamf7",
  # "Ly6c1", "Ly6c2",
  # Exit
  "Klf2",
  "S1pr1"
)

marker_genes_mait_hu <- c(
  # DP & positive selection (TCR induced genes)
  "RAG1",
  "CD24",
  "CD69",
  "EGR2",
  "HIVEP3",
  # Immature
  "CD4",
  "CCR9",
  "SELL",
  "CCR7",
  "LEF1",
  "SLAMF6",
  # Maturation
  "ZBTB16",
  "CD44",
  # Prolif
  "CDK1",
  "MKI67",
  # MAIT17
  "RORC",
  "RORA",
  "IL17RE",
  "CCR6",
  # "SDC1",
  # MAIT1
  "TBX21",
  "EOMES",
  "CXCR3",
  "KLRD1",
  "IFNG",
  "PRF1",
  "SLAMF7",
  # "LY6H", "LY6S",
  # Exit
  "KLF2",
  "S1PR1"
)


```

```{r mait-marker-genes2, fig.height=8, fig.width=4}
do_DotPlot(
  seur_ms_thym$mait,
  group.by="cell_type",
  features=marker_genes_mait_ms,
  flip=T,
  use_viridis = T,
  viridis.palette="B"
)+
  theme(axis.text.y=element_text(face="italic"))
# ggsave("./data/figures/chapter_04/figs_unfinished/section1/ch4_fig4_mait_stages_bigdotplot_mouse.pdf", width=4, height=7)

do_DotPlot(
  seur_hu_thym$mait,
  group.by="clusters_per_lineage",
  features=marker_genes_mait_hu,
  flip=T,
  use_viridis = T,
  viridis.palette="B"
)+
  theme(axis.text.y=element_text(face="italic"))
# ggsave("./data/figures/chapter_04/figs_unfinished/section1/ch4_fig4_mait_stages_bigdotplot_human.pdf", width=4, height=7)
```

```{r}
FeaturePlot(seur_hu_thym$mait, features="CCR9")
```



## MAIT gene signature
```{r mait-gene-signatures-mouse, fig.height=8, fig.width=8}
legoux_gene_signature <- read.csv("./data/litterature_gene_signatures/mouse_tinn_signatures_combined.csv")[,1:2]
legoux_gene_signature_list <- as.list(legoux_gene_signature)
legoux_gene_signature_list <- lapply(legoux_gene_signature_list, function(x) x[x!=""]) # remove empty genes
legoux_gene_signature_list <- lapply(legoux_gene_signature_list, function(x) x[x %in% rownames(seur_ms_thym$mait)]) # remove genes not found in our data
lapply(legoux_gene_signature_list, function(x) length(x)) # sanity check (between 0-5 genes per signature lost)

seur_ms_thym$mait <- AddModuleScore(
  object = seur_ms_thym$mait,
  assay = "RNA",
  features = legoux_gene_signature_list,
  name=names(legoux_gene_signature_list),
  seed=1
  )
colnames(seur_ms_thym$mait@meta.data)[10:11] <- names(legoux_gene_signature_list)

# Plot and save
lineage_genesignature(
  seuratobj = seur_ms_thym$mait,
  # file_name = "ch4_fig3_mait_signatureMAIT1_mouse.pdf",
  genesignature = "legoux_mait1_thymus"
  )
lineage_genesignature(
  seuratobj = seur_ms_thym$mait,
  # file_name = "ch4_fig3_mait_signatureMAIT17_mouse.pdf",
  genesignature = "legoux_mait17_thymus"
)
```

```{r mait-gene-signatures-human, fig.height=8, fig.width=8}
# check how many genes in gene signature we can find in ortholog table
lapply(legoux_gene_signature_list, function(x) table(x %in% ortholog.df$ms_symbol, useNA="ifany"))

# More than 80% of genes in gene signatures can be found in ortholog table, so let's translate them
legoux_gene_signature_list_humangenes <- lapply(legoux_gene_signature_list, function(x) ortholog.df %>% filter(ms_symbol %in% x) %>% pull(hu_symbol))
legoux_gene_signature_list_humangenes <- lapply(legoux_gene_signature_list_humangenes, function(x) x[x %in% rownames(seur_hu_thym$mait)]) # remove genes not found in our data
lapply(legoux_gene_signature_list_humangenes, function(x) length(x)) # sanity check (between 8-16 genes per signature lost, but still >90 genes per signature left)

seur_hu_thym$mait <- AddModuleScore(
  object = seur_hu_thym$mait,
  assay = "RNA",
  features = legoux_gene_signature_list_humangenes,
  name=names(legoux_gene_signature_list_humangenes),
  seed=1
  )
colnames(seur_hu_thym$mait@meta.data)[49:50] <- names(legoux_gene_signature_list_humangenes)

# Plot and save
lineage_genesignature(seuratobj=seur_hu_thym$mait, genesignature = "legoux_mait1_thymus", file_name = "ch4_fig3_mait_signatureMAIT1_human.pdf")
lineage_genesignature(seuratobj=seur_hu_thym$mait, genesignature = "legoux_mait17_thymus", file_name = "ch4_fig3_mait_signatureMAIT17_human.pdf")
```






# SESSION INFO
```{r}
sessionInfo()
```


