---
title: "Chapter 4 - Tconv metaneighbor"
author: "Salomé Carcy"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true

---

This notebook is for figure 1 in chapter 4.

```{r, setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 5, fig.height = 5,
                      warning=FALSE, message=FALSE,
                      root.dir = "~/Projects/phd/")
```


# IMPORT

## Import librairies
```{r import-librairies}
library(ggplot2)
library(RColorBrewer)
library(cowplot)
library(tidyverse)
library(dplyr)
library(Seurat)
library(SCpubr)
library(patchwork)
library(MetaNeighbor)
library(SummarizedExperiment)
library(gplots)
library(reshape2)

source("~/Projects/phd/scripts/colors_universal.R")
```

## Import data
```{r import-data}
# Seurat tables
seur_chopp_human <- readRDS("./data/seurat_objects/chopp2020/HumanThymusChoppEtAl2020.rds")
seur_chopp_human@meta.data$cluster_figure <- Idents(seur_chopp_human)
seur_chopp_mouse <- readRDS("./data/seurat_objects/chopp2020/MouseThymusChoppEtAl2020.rds")

# Remove B cells
seur_chopp_human <- subset(seur_chopp_human, subset=cluster_figure != "hs-Bcell")
seur_chopp_mouse <- subset(seur_chopp_mouse, subset=cluster != "Bcell")

# Orthologs table
ortholog.df <- read.csv("~/Projects/phd/data/cross_species/ortholog_table_ms_hu_one2one.csv")
# ortholog.df <- read.csv("~/Projects/HumanThymusProject/data/cross-species/03_BiomartTable/big_ass_ortholog_table.csv") %>%
#   dplyr::rename(ms_symbol=ms_symbol_bmt) %>%
#   filter(hu_homology_type=="ortholog_one2one") %>%
#   filter(hu_orthology_confidence==1)
```




# PLOT UMAPs

## Clusters on UMAPs
```{r mouse}
names(cols_chopp_ms[names(cols_chopp_ms)!="Bcell"]) %in% unique(seur_chopp_mouse@meta.data$cluster)
seur_chopp_mouse@meta.data$cluster <- factor(
  seur_chopp_mouse@meta.data$cluster,
  levels=names(cols_chopp_ms[names(cols_chopp_ms)!="Bcell"])
  )

ggrastr::rasterise(
  DimPlot(seur_chopp_mouse, group.by="cluster", cols=cols_chopp_ms),
  layers="Point",
  dpi=300
  ) +
  labs(title="Mouse")+
  theme(
    axis.ticks=element_blank(),
    axis.text=element_blank(),
    axis.line = element_blank(),
    panel.border = element_rect(colour = "black", fill=NA, size=1),
    axis.title=element_text(hjust=0)
    )
# ggsave("./data/figures/chapter_04/figs_unfinished/section1/ch4_fig1_umap_mouse.pdf", width=6, height=5)
```

```{r human}
ggrastr::rasterise(
  DimPlot(seur_chopp_human, group.by="cluster_figure", cols=cols_chopp_hu),
  layers="Point",
  dpi=300
  ) +
  labs(title="Human")+
  theme(
    axis.ticks=element_blank(),
    axis.text=element_blank(),
    axis.line = element_blank(),
    panel.border = element_rect(colour = "black", fill=NA, size=1),
    axis.title=element_text(hjust=0)
    )
# ggsave("./data/figures/chapter_04/figs_unfinished/section1/ch4_fig1_umap_human.pdf", width=6.5, height=5)
```


## Gene signatures on UMAPs
Calculate gene signatures on human Chopp et al. data.
```{r gene-signatures}
# Get Park et al. gene signatures
park_markergenes <- read.csv("./data/litterature_gene_signatures/park_genesignatures.csv", row.names=1)
park_gene_signatures <- c(
  "DNp",
  "DNq",
  "DPp",
  "DPq",
  "abT_entry",
  "CD4T",
  "CD8T",
  "CD8aa1",
  "CD8aa2",
  "gdT",
  "T_agonist",
  "Treg_diff",
  "Treg",
  "NKT",
  "Th17"
)
park_gene_signatures_list <- as.list(park_markergenes[,park_gene_signatures])
names(park_gene_signatures_list) <- paste0("park_", names(park_gene_signatures_list))

# # Get Chopp et al. gene signatures
# chopp_signatures <- read.csv("~/Projects/phd/data/litterature_gene_signatures/chopp_gene_signatures.csv")
# chopp_gene_signatures_list <- as.list(chopp_signatures)
# names(chopp_gene_signatures_list) <- paste0("chopp_", names(chopp_gene_signatures_list))

# Personally curated gene signatures from litterature (already scored in integrated seurat object)
perso_gene_signatures <- list(
  "effector"=c("HOPX", "GZMB", "NKG7", "TBX21", "PRF1", "GZMA", "KLRD1", "EOMES", "CCR6", "RORC", "JUNB", "FOS", "RORA", "FOSB"),
  "naive"   =c("SATB1", "TCF7", "LEF1", "CCR7", "SELL", "FOXP1", "KLF2", "SOX4", "ID3", "BACH2"),
  "egress"  =c("KLF2", "CORO1A", "CCR7", "CXCR4", "CXCR6", "FOXO1", "CXCR3", "S1PR1", "S1PR4", "S100A4", "S100A6", "EMP3") # from Sanchez Sanchez
  )


# Combined gene lists
gene_signatures_list <- c(park_gene_signatures_list, perso_gene_signatures)
gene_signatures_list <- lapply(gene_signatures_list, function(x) x[x!=""]) # remove empty genes
gene_signatures_list <- lapply(gene_signatures_list, function(x) x[x %in% rownames(seur_chopp_human)]) # remove genes not found in chopp data
lapply(gene_signatures_list, function(x) length(x)) # sanity check (between 0-2 genes per signature lost)


# add module score to integrated object (better for choice of control genes)
seur_chopp_human <- AddModuleScore(
  object = seur_chopp_human,
  assay = "RNA",
  features = gene_signatures_list,
  name=names(gene_signatures_list),
  seed=1
  )
colnames(seur_chopp_human@meta.data)[11:28] <- names(gene_signatures_list)
```

Plot on UMAPs
```{r gene-signatures-plot, fig.height=5, fig.width=5}
do_FeaturePlot(
  seur_chopp_human,
  features="egress",
  use_viridis = T,
  viridis.palette = "E",
  order=T,
  min.cutoff=0
)
```

Plot as DotPlot
```{r gene-signatures-dotplot, fig.height=8}
do_DotPlot(
  seur_chopp_human,
  group.by="cluster_figure",
  features=colnames(seur_chopp_human@meta.data)[11:28],
  scale=F
)
```




# RUN METANEIGHBOR

## Prepare metadata
```{r mtn-prep-metadata}
# Mouse
ms.metadata <- seur_chopp_mouse@meta.data %>%
  dplyr::select(c(cluster, Genotype)) %>%
  dplyr::rename(cluster_figure=cluster) %>%
  mutate(species="mouse",
         replicate=case_when(Genotype=="B2M" ~ "ms_B2M_KO",
                             Genotype=="IIKO"~ "ms_H2Ab1_KO")) %>%
  dplyr::select(-Genotype)

# Human
hu.metadata <- seur_chopp_human@meta.data %>%
  dplyr::select(c(cluster_figure, orig.ident)) %>%
  mutate(species="human",
         replicate=paste0("hu_", orig.ident)) %>%
  dplyr::select(-orig.ident)
```


## Prepare counts matrices
Get counts matrices and add little prefix to cell barcodes because some are the same across species (and we don't want to confuse them).
```{r mtn-prep-counts-1}
# Get counts matrices
ms.counts <- seur_chopp_mouse[["RNA"]]@counts
hu.counts <- seur_chopp_human[["RNA"]]@counts

# Check they have different cell barcodes
table(colnames(ms.counts) %in% colnames(hu.counts), useNA="ifany")

# They don't, so let's add a little prefix in counts matrices and metadata dataframes
# table(rownames(ms.metadata) == colnames(ms.counts), useNA="ifany") # sanity check
# table(rownames(hu.metadata) == colnames(hu.counts), useNA="ifany") # sanity check
colnames(ms.counts) <- paste0("ms_", colnames(ms.counts))
colnames(hu.counts) <- paste0("hu_", colnames(hu.counts))

rownames(ms.metadata) <- paste0("ms_", rownames(ms.metadata))
rownames(hu.metadata) <- paste0("hu_", rownames(hu.metadata))

# table(colnames(ms.counts) %in% colnames(hu.counts), useNA="ifany") # sanity check
```

Then let's subset the orthologs dataframe to only genes we find in our human/mouse count matrices.
```{r mtn-prep-counts-2}
# First, check whether genes can all be found in the ortholog table. Note: many can't be found because
# (1) I removed genes without orthologs;
# (2) I removed genes with orthology confidence of 0;
# (3) I kept only one2one orthologs
table(unique(rownames(ms.counts)) %in% unique(ortholog.df$ms_symbol)) # 4,202 not
table(unique(rownames(hu.counts)) %in% ortholog.df$hu_symbol) # 9,285 not


# Subset the ortholog table to only genes that we can "translate"
dictionary <- ortholog.df %>%
  as_tibble() %>%
  ## Intersection
  filter(ms_symbol %in% unique(rownames(ms.counts)) & hu_symbol %in% unique(rownames(hu.counts))) %>%
  ## Remove any symbols that are NAs
  filter(!is.na(ms_symbol)) %>%
  filter(!is.na(hu_symbol))
dim(dictionary) # 10,763 genes
```

Now we'll translate mouse genes in mouse count matrix into "human" genes.
```{r mtn-prep-counts-3}
# sanity check how many genes in each count matrix is found in our "dictionary" (should be 10,763 based on previous code chunk)
# table(unique(rownames(ms.counts)) %in% dictionary$ms_symbol) # 10,763
# table(unique(rownames(hu.counts)) %in% dictionary$hu_symbol) # 10,763

# Subset count matrices to only ms and hu genes that have 1:1 orthologs
ms.counts <- ms.counts[rownames(ms.counts) %in% dictionary$ms_symbol,]
hu.counts <- hu.counts[rownames(hu.counts) %in% dictionary$hu_symbol,]
nrow(ms.counts)==nrow(hu.counts) # should be TRUE

# Translate the mouse genes in count matrix into "human genes"
ms.dict <- dictionary %>%
  filter(ms_symbol %in% rownames(ms.counts)) %>%
  select(ms_symbol, hu_symbol) %>%
  distinct(ms_symbol, .keep_all=T)
ms.dict <- ms.dict[match(rownames(ms.counts), ms.dict$ms_symbol),]
table(ms.dict$ms_symbol == rownames(ms.counts), useNA="ifany") # should be all TRUE
table(is.na(ms.dict$hu_symbol)) # should have no NAs

# Translate
rownames(ms.counts) <- ms.dict$hu_symbol
```


## Prepare HVGs
```{r mtn-prep-hvg}
# Mouse hvg
ms.hvg <- VariableFeatures(FindVariableFeatures(seur_chopp_mouse, nfeatures=5000))
length(ms.hvg)

# Human hvg
hu.hvg <- VariableFeatures(FindVariableFeatures(seur_chopp_human, nfeatures=5000))
length(hu.hvg)


# Translate the mouse HVGs into "human gene" language
ms.hvg.translated <- pull(dictionary %>% filter(ms_symbol %in% ms.hvg), # not all ms HVGs are found in ortholog.df
                          hu_symbol)
hu.hvg.translated <- pull(dictionary %>% filter(hu_symbol %in% hu.hvg), # not all hu HVGs are found in ortholog.df
                          hu_symbol)
total.hvg <- unique(union(ms.hvg.translated, hu.hvg.translated))
length(total.hvg) # 1,909 genes (if 2000 HVG each) / 4,549 genes (if 5000 HVGs each)
table(total.hvg %in% unique(ms.hvg.translated), useNA="ifany") # 3,494 come from mouse (2,370 are mouse-specific)
table(total.hvg %in% unique(hu.hvg.translated), useNA="ifany") # 2,179 come from human (1,055 are human-specific)
```


## Prepare SummarizedExperiment object
```{r mtn-prep-se}
# check metadata columns have same colnames
head(ms.metadata)
head(hu.metadata)

# reorder ms and hu count matrices to have same gene order
nrow(ms.counts)
nrow(hu.counts)
ms.counts <- ms.counts[rownames(hu.counts),]

# last sanity checks
table(rownames(ms.counts)==rownames(hu.counts), useNA="ifany") # all TRUE
table(colnames(ms.counts) %in% colnames(hu.counts), useNA="ifany") # all FALSE
table(colnames(ms.counts)==rownames(ms.metadata), useNA="ifany") # all TRUE
table(colnames(hu.counts)==rownames(hu.metadata), useNA="ifany") # all TRUE

# merge count matrices and metadata dataframes
tot.counts   <- cbind(ms.counts, hu.counts)
tot.metadata <- rbind(ms.metadata, hu.metadata)
table(colnames(tot.counts)==rownames(tot.metadata), useNA="ifany") # all TRUE

# create SummarizedExperiment object
se <- SummarizedExperiment(assays=tot.counts, colData=tot.metadata)
```


## Run metaneighbor
```{r mtn-run}
mtn <- MetaNeighborUS(
  var_genes = total.hvg,
  dat = se,
  study_id = se$species,
  cell_type = se$cluster_figure,
  fast_version = T
)
```

Quick plot
```{r mtn-heatmap, fig.height=10, fig.width=10}
heatmap.2(mtn,
          # trace
          trace="none",
          # dendrogram
          # Rowv=FALSE,
          # Colv=FALSE,
          # dendrogram="none",
          # superimpose a density histogram on color key
          density.info="none",
          # color scale
          col=rev(colorRampPalette(brewer.pal(11,"RdYlBu"))(100)),
          breaks=seq(0,1,length=101),
          key.xlab = "AUROC",
          key.title="",
          # keysize = 1.2,
          # text labels
          main="Mouse vs Human Tconv (union: 4549 HVGs)",
          cexRow=0.6,
          cexCol=0.6,
          # margins
          margins=c(9,9))
```




# PLOT METANEIGHBOR

## Define functions
```{r}
plot_mtn <-
  function(mtn_output,
           var1 = "human",
           var2 = "mouse",
           var1_nbcells,
           var2_nbcells,
           var1_order,
           var2_order,
           bp_ymax = 20,
           file_name="no",
           figsize=12
           ) {
    
  # REMODEL DATAFRAME
  mtn.df <- melt(mtn_output) %>%
    filter(str_detect(Var1,var1)) %>%
    mutate(Var1 = gsub(paste0(var1, "\\|"), "", Var1)) %>%
    filter(str_detect(Var2, var2)) %>%
    mutate(Var2 = gsub(paste0(var2, "\\|"), "", Var2)) %>%
    as_tibble() %>%
    dplyr::rename(auroc=value) %>%
    # add nb of cells for var1 and var2
    left_join(var1_nbcells, by="Var1") %>%
    left_join(var2_nbcells, by="Var2") %>%
    # transform var1 and var2 into factors for ordering
    mutate(Var1=factor(Var1, levels=var1_order),
           Var2=factor(Var2, levels=var2_order))
    
    # PROPORTION OF VAR1 THYMOCYTES IN EACH CLUSTER
    bp.x <-
      ggplot(
        data = mtn.df %>% select(Var1, propcells_var1) %>% distinct(),
        aes(x = Var1, y = propcells_var1)
      ) +
      geom_bar(stat = "identity", fill = "#bdbdbd") + theme_cowplot() +
      scale_x_discrete(position = "top") +
      scale_y_continuous(limits = c(0, bp_ymax), breaks = c(0, 10, 20)) +
      labs(y = "%cells") +
      theme(
        axis.text.x = element_text(angle = 45, hjust = 0),
        axis.ticks.y = element_blank(),
        axis.title.x = element_blank(),
        axis.line.x = element_blank(),
        legend.position = "none"
      )
    
    # PROPORTION OF VAR2 THYMOCYTES IN EACH CLUSTER
    bp.y <-
      ggplot(
        data = mtn.df %>% select(Var2, propcells_var2) %>% distinct(),
        aes(x = Var2, y = propcells_var2)
      ) +
      geom_bar(stat = "identity", fill = "#bdbdbd") +
      scale_x_discrete(position = "top") +
      scale_y_continuous(limits = c(0, bp_ymax), breaks = c(0, 10, 20)) +
      labs(y = "%cells") + coord_flip() + theme_cowplot() +
      theme(
        axis.title.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.line.y = element_blank(),
        legend.position = "none"
      )
    
    # BUBBLEPLOT
    hm.clean <-
      ggplot(mtn.df, aes(x = Var1, y = Var2)) +
      geom_point(aes(size = abs(auroc-0.5), color = auroc)) +
      geom_text(
        data = mtn.df %>% filter(auroc > 0.8) %>% mutate(across("auroc", \(x) round(x, 2))),
        aes(label = auroc),
        color = "white"
      ) +
      scale_size_continuous(
        limits = c(0, 0.5),
        breaks = seq(0, 0.5, by = 0.1),
        range = c(1, 15)
      ) +
      scale_color_gradient2(
        low = "#2166ac",
        mid = "white",
        high = "#a50f15",
        midpoint = 0.5,
        limits = c(0, 1),
        name = "AUROC",
        breaks = seq(0, 1, by = 0.2)
      ) +
      labs(x = paste(var1, "clusters"), y = paste(var2, "clusters"), size = "AUROC") +
      theme_cowplot() +
      theme(
        legend.position = "bottom",
        legend.key.width = unit(0.8, 'cm'),
        axis.text.x = element_text(angle = 45, hjust = 1)
      )
    
    # COMBINE
    p <- (bp.x+plot_spacer() + plot_layout(widths = c(5, 1))) / (hm.clean + bp.y + plot_layout(widths = c(5, 1))) + plot_layout(heights = c(1, 5))
    if(file_name!="no"){ggsave(paste0("./data/figures/chapter_04/figs_unfinished/section1/", file_name), plot=p, width=figsize, height=figsize)}
    return(p)
}

# Get nb of cells in human and mouse clusters
nbcells_hu <- as.data.frame(table(seur_chopp_human$cluster_figure)) %>%
    dplyr::rename(ncells_var1=Freq) %>%
    mutate(propcells_var1 = ncells_var1*100/dim(seur_chopp_human)[2]) 
nbcells_ms <- as.data.frame(table(seur_chopp_mouse$cluster)) %>%
    dplyr::rename(Var2=Var1, ncells_var2=Freq) %>%
    mutate(propcells_var2 = ncells_var2*100/dim(seur_chopp_mouse)[2])

# Define order of mouse and human clusters
# hu_levels <- levels(seur_chopp_human@meta.data$cluster_figure)[levels(seur_chopp_human@meta.data$cluster_figure)!="hs-Bcell"]
hu_levels <- c(
  "hs-ISP",
  "hs-DP1",
  "hs-DP2",
  "hs-DP3",
  "hs-Sig1",
  "hs-Sig4",
  "hs-Sig2",
  "hs-Sig3",
  "hs-PreTreg",
  "hs-Treg",
  "hs-Treg2",
  "hs-ImCD4",
  "hs-MatCD4",
  "hs-ISC",
  "hs-ImCD8",
  "hs-CD8",
  "hs-MatCD8",
  "hs-Mix",
  "hs-NC1",
  "hs-Unk"
)
# ms_levels <- names(cols_chopp_ms)[names(cols_chopp_ms)!="Bcell"]
ms_levels <- c(
  "DP-1",
  "DP-2 ",
  "Sig-1",
  "Sig-2",
  "Sig-3",
  "Sig-4",
  "Sig-5a",
  "Sig-5b",
  "Treg",
  "ImCD4",
  "MatCD4",
  "ISC CD4",
  "ImCD8",
  "MatCD8",
  "NC-1",
  "NC-2"
)
```


## Mouse x Human
```{r mtn-plot-ms-hu, fig.height=10, fig.width=10}
p_ms_hu <- plot_mtn(
  mtn_output = mtn,
  var1="human",
  var2="mouse",
  var1_nbcells = nbcells_hu,
  var2_nbcells = nbcells_ms,
  var1_order = hu_levels,
  var2_order= rev(ms_levels),
  # file_name="no"
  file_name="ch4_fig1_tconv_mtn_5000hvg.pdf"
)
p_ms_hu
```


## Mouse x Mouse
```{r mtn-plot-ms-ms, fig.height=10, fig.width=10}
p_ms_ms <- plot_mtn(
  mtn_output = mtn,
  var1="mouse",
  var2="mouse",
  var1_nbcells=nbcells_ms %>% dplyr::rename(Var1=Var2, ncells_var1=ncells_var2, propcells_var1=propcells_var2),
  var2_nbcells = nbcells_ms,
  var1_order = ms_levels,
  var2_order= rev(ms_levels),
  file_name="ch4_fig1_tconv_mtn_5000hvg_ms_ms.pdf"
)
p_ms_ms
```


## Human x Human
```{r mtn-plot-hu-hu, fig.height=10, fig.width=10}
p_hu_hu <- plot_mtn(
  mtn_output = mtn,
  var1="human",
  var2="human",
  var1_nbcells=nbcells_hu,
  var2_nbcells = nbcells_hu %>% dplyr::rename(Var2=Var1, ncells_var2=ncells_var1, propcells_var2=propcells_var1),
  var1_order = hu_levels,
  var2_order= rev(hu_levels),
  file_name="ch4_fig1_tconv_mtn_5000hvg_hu_hu.pdf",
  figsize=14
)
p_hu_hu
```




# PLOT MARKER GENES

## Marker genes
```{r}
# Mouse
marker_genes_ms <- c(
  # DP
  "Rag1",
  "Dntt",
  "Cd4",
  "Cd8a",
  "Cd8b1",
  # SP entry
  "Ccr9",
  "Tox",
  "Lef1",
  # TCR induced
  "Cd69",
  "Cd5",
  "Ikzf2",
  "Id3",
  # CD8aa1 (IELpA)
  "Gng4",
  "Nr4a3",
  "Tnfrsf9",
  "Pdcd1",
  "Hivep3",
  # CD8aa2 (IELpB?)
  # "Znf683",
  "Xcl1",
  "Il2rb",
  # Tagonist
  "Bcl2l11",
  "Nr4a1",
  # Treg
  "Foxp3",
  "Ctla4",
  "Il2ra",
  "Tnfrsf18",
  "Tnfrsf4",
  # CD4 SP
  "Zbtb7b",
  "Cd40lg",
  # "Bcl11b",
  # "Satb1",
  # "Tcf7",
  # CD8 SP
  "Runx3",
  "Plac8",
  # Maturation & exit
  "Ccr7",
  "Klf2",
  "S1pr1",
  # "Sell",
  # "Il7r",
  # NC type 1
  "Zbtb16",
  "Tbx21",
  "Ifng",
  "Gzmk",
  # NC type 17
  "Rorc",
  "Rora",
  "Il17rb"
)

# Human
marker_genes_hu <- c(
  # DP
  "RAG1",
  "DNTT",
  "CD4",
  "CD8A",
  "CD8B",
  # SP entry
  "CCR9",
  "TOX",
  "LEF1",
  # TCR induced
  "CD69",
  "CD5",
  "IKZF2",
  "ID3",
  # CD8aa1
  "GNG4",
  "NR4A3",
  "TNFRSF9",
  "PDCD1",
  "HIVEP3",
  # CD8aa2
  "ZNF683",
  "XCL1",
  "IL2RB",
  # T agonist
  "BCL2L11",
  "NR4A1",
  # Treg
  "FOXP3",
  "CTLA4",
  "IL2RA",
  "TNFRSF18",
  "TNFRSF4",
  # CD4 SP
  "ZBTB7B",
  "CD40LG",
  # "BCL11B",
  # "SATB1",
  # "TCF7",
  # CD8 SP
  "RUNX3",
  "PLAC8",
  "LINC02446",
  # Maturation & exit
  "CCR7",
  "KLF2",
  "S1PR1",
  # "SELL",
  # "IL7R",
  # NC
  "ZBTB16",
  "TBX21",
  "IFNG",
  "GZMK",
  "RORC",
  "RORA",
  "IL17RB"
)


length(marker_genes_ms)==length(marker_genes_hu)
```


## Mouse
```{r, fig.height=8, fig.width=5}
seur_chopp_mouse$cluster <- factor(seur_chopp_mouse$cluster, levels=ms_levels)

do_DotPlot(
  seur_chopp_mouse,
  group.by="cluster",
  features=marker_genes_ms,
  flip=T,
  use_viridis = T,
  viridis.palette="B"
)+
  theme(axis.text.y=element_text(face="italic"))
# ggsave("./data/figures/chapter_04/figs_unfinished/section1/ch4_fig1_dotplot_mouse.pdf", width=5, height=9)
```

## Human
```{r, fig.height=7, fig.width=5}
seur_chopp_human$cluster_figure <- factor(seur_chopp_human$cluster_figure, levels=hu_levels)

do_DotPlot(
  seur_chopp_human,
  group.by="cluster_figure",
  features=marker_genes_hu,
  flip=T,
  use_viridis = T,
  viridis.palette="B"
)+
  theme(axis.text.y=element_text(face="italic"))
# ggsave("./data/figures/chapter_04/figs_unfinished/section1/ch4_fig1_dotplot_human.pdf", width=6, height=9)
```




# SESSION INFO
```{r}
sessionInfo()
```


