---
title: "Chapter 4 - Cross-species gdT"
author: "Salomé Carcy"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 5, fig.height = 5,
                      warning=FALSE, message=FALSE,
                      root.dir = "~/Projects/phd/")
```

# IMPORT

## Import librairies

```{r import-librairies}
library(ggplot2)
library(RColorBrewer)
library(cowplot)
library(tidyverse)
library(dplyr)
library(Seurat)
library(SCpubr)
library(patchwork)
library(SummarizedExperiment)
library(MetaNeighbor)
library(gplots)
library(reshape2)
# library(ggrepel)

source("~/Projects/phd/scripts/colors_universal.R")
```

## Import data

```{r import-data}
# Seurat objects
seur_hu_thym <- list(
  "nkt"=readRDS("./data/seurat_objects/thymus_objects/seurat_thymus_nkt_with_gene_signatures.rds"),
  "mait"=readRDS("./data/seurat_objects/thymus_objects/seurat_thymus_mait_with_gene_signatures.rds"),
  "gdt"=readRDS("./data/seurat_objects/thymus_objects/seurat_thymus_gdt_with_gene_signatures.rds")
)

seur_ms_thym <- list(
  "nkt"=readRDS("./data/seurat_objects/harshakrovi2020/ms_nkt_seurobj.rds"),
  "mait"=readRDS("./data/seurat_objects/legoux2019/ms_mait_seurobj.rds"),
  "gdt"=readRDS("./data/seurat_objects/lee2020/ms_gdt_seurobj_lee.rds")
)

seur_ms_thym$gdt$gd_clusters_final <- factor(seur_ms_thym$gdt$gd_clusters_final, levels=c(
  "Tγδp",
  "immature Tγδ1/17",
  "immature Tγδ17",
  "Tγδ17",
  "immature Tγδ1",
  "Tγδ1"
))

# Orthologs table
ortholog.df <- read.csv("~/Projects/phd/data/cross_species/ortholog_table_ms_hu_full.csv")
```




# FUNCTIONS

```{r define-functions}
lineage_genesignature <- function(seuratobj, genesignature, file_name="no"){
  p <- ggrastr::rasterise(
    do_FeaturePlot(
    seuratobj,
    features = genesignature,
    min.cutoff=0,
    use_viridis = T,
    viridis.palette = "E",
    legend.position="right",
    order=T,
    pt.size = 1.5,
    raster=F
    # raster.dpi=2048,
    # pt.size=10
  ),
  layers="Point",
  dpi=300
  )
  print(p)
  if(file_name != "no"){
    ggsave(paste0("~/Projects/phd/data/figures/chapter_04/figs_unfinished/section1/", file_name),
           plot=p,
           width=9, height=6)
  }
}

plot_DotPlot <- function(seurobj, group, features, scaling=T){
  # get plot data
  p <- Seurat::DotPlot(
    seurobj,
    group.by=group,
    features=features,
    scale=scaling
  )
  # plot in personalized way
  p <- ggplot(p$data, aes(x=id, y=features.plot, fill=avg.exp.scaled, size=pct.exp))+
    geom_point(color="black", shape=21)+
    # scale_fill_gradient2(low=scales::muted("blue"), high=scales::muted("red"), name="z-score\nnormalized\navg expression")+
    scale_size_continuous(range=c(0,6), limits=c(0,100), name="%cells\nexpressing\ngene")+
    theme_bw()+
    theme(axis.text.y=element_text(face="italic"),
          axis.text.x=element_text(angle=45, hjust=1))+
    labs(y="", x="Clusters")
  # different color scale if scaled or not
  if(scaling==T){
    p <- p + scale_fill_gradient2(low=scales::muted("blue"), high=scales::muted("red"), name="z-score\nnormalized\navg expression")
  } else{
    p <- p + viridis::scale_fill_viridis(option="B", direction=-1, name="normalized\navg expression")
  }
  return(p)
}

plot_mtn <-
  function(mtn_output,
           var1 = "human",
           var2 = "mouse",
           var1_nbcells,
           var2_nbcells,
           var1_order,
           var2_order,
           bp_ymax = 20,
           file_name="no",
           figsize=12
           ) {
    
  # REMODEL DATAFRAME
  mtn.df <- melt(mtn_output) %>%
    filter(str_detect(Var1,var1)) %>%
    mutate(Var1 = gsub(paste0(var1, "\\|"), "", Var1)) %>%
    filter(str_detect(Var2, var2)) %>%
    mutate(Var2 = gsub(paste0(var2, "\\|"), "", Var2)) %>%
    as_tibble() %>%
    dplyr::rename(auroc=value) %>%
    # add nb of cells for var1 and var2
    left_join(var1_nbcells, by="Var1") %>%
    left_join(var2_nbcells, by="Var2") %>%
    # transform var1 and var2 into factors for ordering
    mutate(Var1=factor(Var1, levels=var1_order),
           Var2=factor(Var2, levels=var2_order))
  
    
    # PROPORTION OF VAR1 THYMOCYTES IN EACH CLUSTER
    bp.x <-
      ggplot(
        data = mtn.df %>% select(Var1, propcells_var1) %>% distinct(),
        aes(x = Var1, y = propcells_var1)
      ) +
      geom_bar(stat = "identity", fill = "#bdbdbd") + theme_cowplot() +
      scale_x_discrete(position = "top") +
      scale_y_continuous(limits = c(0, bp_ymax), breaks = seq(0,bp_ymax, length.out=3)) +
      labs(y = "%cells") +
      theme(
        axis.text.x = element_text(angle = 45, hjust = 0),
        axis.ticks.y = element_blank(),
        axis.title.x = element_blank(),
        axis.line.x = element_blank(),
        legend.position = "none"
      )
    
    # PROPORTION OF VAR2 THYMOCYTES IN EACH CLUSTER
    bp.y <-
      ggplot(
        data = mtn.df %>% select(Var2, propcells_var2) %>% distinct(),
        aes(x = Var2, y = propcells_var2)
      ) +
      geom_bar(stat = "identity", fill = "#bdbdbd") +
      scale_x_discrete(position = "top") +
      scale_y_continuous(limits = c(0, bp_ymax), breaks = seq(0,bp_ymax, length.out=3)) +
      labs(y = "%cells") + coord_flip() + theme_cowplot() +
      theme(
        axis.title.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.line.y = element_blank(),
        legend.position = "none"
      )
    
    # BUBBLEPLOT
    hm.clean <-
      ggplot(mtn.df, aes(x = Var1, y = Var2)) +
      geom_point(aes(size = abs(auroc-0.5), fill = auroc), color="black", shape=21) +
      geom_text(
        data = mtn.df %>% filter(auroc > 0.8) %>% mutate(across("auroc", \(x) round(x, 2))),
        aes(label = auroc),
        color = "white"
      ) +
      scale_size_continuous(
        limits = c(0, 0.5),
        breaks = seq(0, 0.5, by = 0.1),
        range = c(1, 15)
      ) +
      scale_fill_gradient2(
        low = "#2166ac",
        mid = "white",
        high = "#a50f15",
        midpoint = 0.5,
        limits = c(0, 1),
        name = "AUROC",
        breaks = seq(0, 1, by = 0.2)
      ) +
      labs(x = paste(var1, "clusters"), y = paste(var2, "clusters"), size = "AUROC") +
      theme_cowplot() +
      theme(
        legend.position = "bottom",
        legend.key.width = unit(0.8, 'cm'),
        axis.text.x = element_text(angle = 45, hjust = 1)
      )
    
    # COMBINE
    p <- (bp.x+plot_spacer() + plot_layout(widths = c(5, 1))) / (hm.clean + bp.y + plot_layout(widths = c(5, 1))) + plot_layout(heights = c(1, 5))
    if(file_name!="no"){ggsave(paste0("./data/figures/chapter_04/figs_unfinished/section1/", file_name), plot=p, width=figsize, height=figsize, device=cairo_pdf)}
    return(p)
  }

# Get nb of cells in human and mouse clusters
nbcells_hu <- as.data.frame(table(seur_hu_thym$gdt$clusters_per_lineage)) %>%
    dplyr::rename(ncells_var1=Freq) %>%
    mutate(propcells_var1 = ncells_var1*100/dim(seur_hu_thym$gdt)[2]) 
nbcells_ms <- as.data.frame(table(seur_ms_thym$gdt$gd_clusters_final)) %>%
    dplyr::rename(Var2=Var1, ncells_var2=Freq) %>%
    mutate(propcells_var2 = ncells_var2*100/dim(seur_ms_thym$gdt)[2])

# Define order of mouse and human clusters
hu_levels <- paste0("GDT_thymus_c", 0:7)
ms_levels <- rev(levels(seur_ms_thym$gdt$gd_clusters_final))
```

Let's do a quick cleanup of the γδT object & save.
```{r gdt-cleanup, echo=T, eval=F}
seur_ms_thym$gdt@meta.data$gd_clusters_final <- case_when(
  seur_ms_thym$gdt@meta.data$gd_subclusters== "G7-1 (Tγδ1)" ~ "immature Tγδ1",
  seur_ms_thym$gdt@meta.data$gd_subclusters== "G7-2 (Tγδ1)" ~ "Tγδ1",
  .default=seur_ms_thym$gdt@meta.data$gd_clusters
)

# sanity check
table(seur_ms_thym$gdt@meta.data[,c("gd_subclusters", "gd_clusters_final")], useNA="ifany")
# saveRDS(seur_ms_thym$gdt, "./data/seurat_objects/lee2020/ms_gdt_seurobj_lee.rds")
```

We also need to translate ENSEMBL genes to gene symbols.
```{r gdt-cleanup2}
# ---
# TRANSLATE GENE ENSEMBL TO MOUSE SYMBOLS
ms_ensembl <- read.csv("./data/cross_species/biomart_export_ms.txt")[,1:3]
colnames(ms_ensembl) <- c("gene_ensembl", "gene_version", "gene_symbol")
ms_ensembl <- ms_ensembl %>% distinct() %>% filter(gene_symbol != "")
table(rownames(seur_ms_thym$gdt) %in% ms_ensembl$gene_ensembl, useNA="ifany") # 515 genes out of 25614 don't have gene symbol
table(VariableFeatures(seur_ms_thym$gdt) %in% ms_ensembl$gene_ensembl, useNA="ifany") # only 5 HVGs don't have gene symbol
# check if some gene symbols are assigned to multiple ENSEMBL
ms_ensembl %>%
  group_by(gene_symbol) %>%
  filter(n_distinct(gene_ensembl)>1) %>%
  dplyr::arrange(gene_symbol) %>%
  filter(gene_ensembl %in% rownames(seur_ms_thym$gdt)) # mostly some Rik and Gm genes
# let's keep only genes that have 1:1 ENSEMBL to gene symbol matching (for now for simplicity)
genes_to_keep <- ms_ensembl %>%
  group_by(gene_symbol) %>%
  filter(n_distinct(gene_ensembl)==1) %>%
  ungroup() %>%
  group_by(gene_ensembl) %>%
  filter(n_distinct(gene_symbol)==1) %>%
  ungroup() %>%
  pull(gene_ensembl)
# subset seurat object to genes with known symbols
seur_ms_thym$gdt <- seur_ms_thym$gdt[rownames(seur_ms_thym$gdt) %in% genes_to_keep]
# subset gene dataframe to seurat genes
ms_ensembl <- ms_ensembl %>% filter(gene_ensembl %in% rownames(seur_ms_thym$gdt))
ms_ensembl <- ms_ensembl[match(rownames(seur_ms_thym$gdt), ms_ensembl$gene_ensembl),] # reorder in same order as genes in seurat object
table(ms_ensembl$gene_ensembl==rownames(seur_ms_thym$gdt), useNA="ifany") # sanity check

# DISASSEMBLE SEURAT AND TRANSLATE TO GENE SYMBOLS
count_matrix <- seur_ms_thym$gdt@assays$RNA@counts
# count_matrix <- seur_ms_thym$gdt@assays[["RNA"]]
data_matrix <- seur_ms_thym$gdt@assays$RNA@data
hvg <- VariableFeatures(seur_ms_thym$gdt)
# table(rownames(count_matrix)==ms_ensembl$gene_ensembl, useNA="ifany")
# table(rownames(data_matrix)==ms_ensembl$gene_ensembl, useNA="ifany")
# table(rownames(seur_ms_thym$gdt@assays$RNA@meta.features)==ms_ensembl$gene_ensembl, useNA="ifany")
rownames(count_matrix) <- ms_ensembl$gene_symbol
rownames(data_matrix) <- ms_ensembl$gene_symbol
hvg <- ms_ensembl %>% filter(gene_ensembl %in% hvg) %>% pull(gene_symbol)

# REASSEMBLE SEURAT
rownames(seur_ms_thym$gdt@assays$RNA@meta.features) <- ms_ensembl$gene_symbol
seur_ms_thym$gdt@assays$RNA@counts <- count_matrix
seur_ms_thym$gdt@assays$RNA@data <- data_matrix
VariableFeatures(seur_ms_thym$gdt) <- hvg
# saveRDS(seur_ms_thym$gdt, "./data/seurat_objects/lee2020/ms_gdt_seurobj_lee_genesymbols.rds")
```




# γδT FIGURE

## γδT UMAPs
```{r gdt-umap-mouse}
ggrastr::rasterise(
  do_DimPlot(
    seur_ms_thym$gdt,
    group.by = "gd_clusters_final",
    colors.use = cols_lee_ms,
    legend.position = "right"
  ),
  layers = "Point",
  dpi = 300
) +
  labs(title = "Mouse") +
  theme(
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    axis.line = element_blank(),
    panel.border = element_rect(
      colour = "black",
      fill = NA,
      size = 1
    ),
    axis.title = element_text(hjust = 0)
  )
# ggsave("./data/figures/chapter_04/figs_unfinished/section1/ch4_fig5_gdt_umap_mouse.pdf", width=7, height=5, device=cairo_pdf)
```

```{r gdt-umap-human}
ggrastr::rasterise(
  do_DimPlot(seur_hu_thym$gdt, group.by="clusters_per_lineage", colors.use=cols_thym_gdt, legend.position="right"),
  layers="Point",
  dpi=300
  ) +
  labs(title="Human")+
  theme(
    axis.ticks=element_blank(),
    axis.text=element_blank(),
    axis.line = element_blank(),
    panel.border = element_rect(colour = "black", fill=NA, size=1),
    axis.title=element_text(hjust=0)
    )
# ggsave("./data/figures/chapter_04/figs_unfinished/section1/ch4_fig5_gdt_umap_human.pdf", width=7, height=5)
```


## γδT DotPlots
Here we select genes that define each mouse cluster based on [Lee et al. 2020](https://www.nature.com/articles/s41467-020-18155-8), [Sumaria et al. 2017](https://www.sciencedirect.com/science/article/pii/S2211124717307283?via%3Dihub), [Sumaria et al. 2018](https://onlinelibrary.wiley.com/doi/10.1111/imm.13032) and . These are not "marker genes" *per se*, as we didn't obtain them by computing the most DE genes in each cluster. Instead we selected the genes highlighted in the literature as defining each $\gamma\delta$ developmental stage.
```{r gdt-marker-genes1, fig.height=8, fig.width=4}
marker_genes_gdt_ms <- c(
  # GDp (whatever this is)
  "Cdk1",
  "Mki67",
  # "Cd27",
  # "Egr1",
  # "Egr3",
  # "Id3",
  # "Tcf3",
  # Maturation markers (loss CD24, gain CD44)
  "Il2ra", # CD25 not well detected
  "Cd24a",
  "Clec12a",
  "Cd44",
  # ----
  # GD 2
  # "Il4",
  # "Cd27", # also GD1
  # "Zbtb16", # also GD NKT
  # "Cd4",
  # "Il17rb",
  # "Slamf6",
  # "Il21r",
  # "Satb1",
  # "Egr2",
  # ----
  # GD 17
  # "Sox4",
  "Sox13",
  "Notch1",
  # "Smad3",
  "Rorc",
  "Rora",
  # "Cd44",
  "Il7r",
  "Ccr6",
  "Il23r",
  "Tmem176a",
  # ----
  # GD 1
  "Cd27",
  "Cd200",
  "Nr4a1",
  "Tbx21",
  "Eomes",
  "Ifng",
  "Cxcr3",
  "Nt5e",
  "Xcl1",
  "Il2rb", # GD NKT
  # "Klrb1c", # GD NKT (barely expressed there)
  "Slamf7",
  # Exit
  "Klf2",
  "S1pr1"
)


marker_genes_gdt_hu <- c(
  # GDp (whatever this is)
  "CDK1",
  "MKI67",
  # "Cd27",
  # "Egr1",
  # "Egr3",
  # "Id3",
  # "Tcf3",
  # Maturation markers (loss CD24, gain CD44)
  "IL2RA", # CD25 not well detected
  "CD24",
  "CLEC12A",
  "CD44",
  # ----
  # GD 2
  # "Il4",
  # "Cd27", # also GD1
  # "Zbtb16", # also GD NKT
  # "Cd4",
  # "Il17rb",
  # "Slamf6",
  # "Il21r",
  # "Satb1",
  # "Egr2",
  # ----
  # GD 17
  # "Sox4",
  "SOX13",
  "NOTCH1",
  # "Smad3",
  "RORC",
  "RORA",
  # "Cd44",
  "IL7R",
  "CCR6",
  "IL23R",
  "IL12RB1", # TMEM176A was not found, so putting a placeholder gene instead
  # ----
  # GD 1
  "CD27",
  "CD200",
  "NR4A1",
  "TBX21",
  "EOMES",
  "IFNG",
  "CXCR3",
  "NT5E",
  "XCL1",
  "IL2RB", # GD NKT
  # "Klrb1c", # GD NKT (barely expressed there)
  "SLAMF7",
  # Exit
  "KLF2",
  "S1PR1"
)

```

```{r gdt-marker-genes2, fig.height=8, fig.width=8}
plot_DotPlot(
  seurobj = seur_ms_thym$gdt,
  group = "gd_clusters_final",
  features = marker_genes_gdt_ms,
  scaling = T
) +
plot_DotPlot(
  seurobj = seur_hu_thym$gdt,
  group = "clusters_per_lineage",
  features = marker_genes_gdt_hu,
  scaling = T
)
# ggsave("./data/figures/chapter_04/figs_unfinished/section1/ch4_fig5_gdt_stages_dotplot.pdf", width=8, height=7, device=cairo_pdf)
```


## γδT gene signature
Let's first prepare the mouse and human gene signatures.
```{r gdt-gene-signatures-prep}
# Import mouse gene signatures
lee_gene_signature <- read.csv("./data/litterature_gene_signatures/mouse_tinn_signatures_combined.csv")[,6:7]
lee_gene_signature_list <- as.list(lee_gene_signature)
lee_gene_signature_list <- lapply(lee_gene_signature_list, function(x) x[x!=""]) # remove empty genes
lee_gene_signature_list <- lapply(lee_gene_signature_list, function(x) unique(x)) # keep unique genes

# # translate into ENSEMBL IDs
# library(biomaRt)
# mart.ms <- useMart(biomart="ENSEMBL_MART_ENSEMBL", dataset="mmusculus_gene_ensembl")
# mart.ms.df <- getBM(attributes = c('ensembl_gene_id', 'external_gene_name'),
#                     filters = 'external_gene_name',
#                     values = c(lee_gene_signature_list$lee_gdt17_thymus, lee_gene_signature_list$lee_gdt1_thymus),
#                     mart = mart.ms)
# # table(lee_gene_signature_list$lee_gdt17_thymus %in% mart.ms.df$external_gene_name, useNA="ifany") # only 10 out of 150 not found
# # table(lee_gene_signature_list$lee_gdt1_thymus %in% mart.ms.df$external_gene_name, useNA="ifany") # only 6 out of 180 not found
# lee_gene_signature_list$lee_gdt17_thymus <- mart.ms.df %>% filter(external_gene_name %in% lee_gene_signature_list$lee_gdt17_thymus) %>% pull(ensembl_gene_id)
# lee_gene_signature_list$lee_gdt1_thymus <- mart.ms.df %>% filter(external_gene_name %in% lee_gene_signature_list$lee_gdt1_thymus) %>% pull(ensembl_gene_id)

 # remove ms genes not found in our ms data
lee_gene_signature_list <- lapply(lee_gene_signature_list, function(x) x[x %in% rownames(seur_ms_thym$gdt)])
lapply(lee_gene_signature_list, function(x) length(x)) # sanity check (between 7-10 genes per signature lost)

# the Tγδ1 signature didn't seem very specific to Tγδ1 but was also high in Tγδ17 (not shown)
# so let's remove Tγδ17 signature genes we find in the Tγδ1 signature
table(lee_gene_signature_list$lee_gdt1_thymus %in% lee_gene_signature_list$lee_gdt17_thymus)
lee_gene_signature_list$lee_gdt1_thymus <- lee_gene_signature_list$lee_gdt1_thymus[!lee_gene_signature_list$lee_gdt1_thymus %in% lee_gene_signature_list$lee_gdt17_thymus]

# check how many genes in gene signature we can find in ortholog table
lapply(lee_gene_signature_list, function(x) table(x %in% ortholog.df$ms_symbol, useNA="ifany"))

# More than 80% of genes in gene signatures can be found in ortholog table, so let's translate them
lee_gene_signature_list_humangenes <- lapply(lee_gene_signature_list, function(x) ortholog.df %>% filter(ms_symbol %in% x) %>% pull(hu_symbol))
lee_gene_signature_list_humangenes <- lapply(lee_gene_signature_list_humangenes, function(x) unique(x[x %in% rownames(seur_hu_thym$gdt)])) # remove genes not found in hu data, and make sure to have unique genes
lapply(lee_gene_signature_list_humangenes, function(x) length(x)) # sanity check (between 10-13 genes per signature lost, but still >100 genes per signature left)

# Let's remove any mouse genes in the mouse signatures that didn't have human orthologs, or the human orthologs were not found in human data
lee_gene_signature_list_mousegenes <- lapply(lee_gene_signature_list_humangenes, function(x) ortholog.df %>% filter(hu_symbol %in% x) %>% pull(ms_symbol))
lee_gene_signature_list_mousegenes <- lapply(lee_gene_signature_list_mousegenes, function(x) unique(x[x %in% rownames(seur_ms_thym$gdt)])) # remove genes not found in ms data, and make sure to have unique genes
lapply(lee_gene_signature_list_mousegenes, function(x) length(x)) # sanity check (there may be a slight different number of genes compared to human because of the 1:many and many:many orthologs)
```

Now let's score and plot!
```{r gdt-gene-signatures-mouse, fig.height=8, fig.width=8}
# Score
seur_ms_thym$gdt <- AddModuleScore(
  object = seur_ms_thym$gdt,
  assay = "RNA",
  features = lee_gene_signature_list_mousegenes,
  name=names(lee_gene_signature_list_mousegenes),
  seed=1
  )
colnames(seur_ms_thym$gdt@meta.data)[17:18] <- names(lee_gene_signature_list_mousegenes)

# Plot and save
lineage_genesignature(
  seuratobj = seur_ms_thym$gdt,
  # file_name = "ch4_fig5_gdt_signatureGDT1_mouse.pdf",
  genesignature = "lee_gdt1_thymus"
  )
lineage_genesignature(
  seuratobj = seur_ms_thym$gdt,
  # file_name = "ch4_fig5_gdt_signatureGDT17_mouse.pdf",
  genesignature = "lee_gdt17_thymus"
)
```

```{r gdt-gene-signatures-human, fig.height=8, fig.width=8}
seur_hu_thym$gdt <- AddModuleScore(
  object = seur_hu_thym$gdt,
  assay = "RNA",
  features = lee_gene_signature_list_humangenes,
  name=names(lee_gene_signature_list_humangenes),
  seed=1
  )
colnames(seur_hu_thym$gdt@meta.data)[49:50] <- names(lee_gene_signature_list_humangenes)

# Plot and save
lineage_genesignature(
  seuratobj = seur_hu_thym$gdt,
  # file_name = "ch4_fig5_gdt_signatureGDT1_human.pdf",
  genesignature = "lee_gdt1_thymus"
)
lineage_genesignature(
  seuratobj = seur_hu_thym$gdt,
  # file_name = "ch4_fig5_gdt_signatureGDT17_human.pdf",
  genesignature = "lee_gdt17_thymus"
)
```




# γδT METANEIGHBOR

## Prepare data for metaneighbor
```{r gdt-prep-metaneighbor}
# Get the ms seurat object with ENSEMBL IDs again (it's better for identifying orthologs)
seur_ms_thym$gdt <- readRDS("./data/seurat_objects/lee2020/ms_gdt_seurobj_lee.rds")
# rownames(seur_ms_thym$gdt)[1:5]

# Get the ortholog table with 1:1 orthologs only and orthology confidence of 1
ortholog_df_1to1 <- read.csv("~/Projects/phd/data/cross_species/ortholog_table_ms_hu_one2one.csv")

# Get counts, HVGs and metadata
ms.hvg <- VariableFeatures(FindVariableFeatures(seur_ms_thym$gdt, nfeatures=5000))
length(ms.hvg)
ms.counts <- seur_ms_thym$gdt[["RNA"]]@counts
ms.metadata <- seur_ms_thym$gdt@meta.data

hu.hvg <- VariableFeatures(FindVariableFeatures(seur_hu_thym$gdt, nfeatures=5000))
length(hu.hvg)
hu.counts <- seur_hu_thym$gdt[["RNA"]]@counts
hu.metadata <- seur_hu_thym$gdt@meta.data


# First, check whether genes can all be found in the ortholog table (many can't be found because I removed genes with orthology confidence=0)
table(unique(rownames(ms.counts)) %in% unique(ortholog_df_1to1$ms_ensemblID)) # 13,117 not
table(unique(rownames(hu.counts)) %in% ortholog_df_1to1$hu_symbol) # 5,238 not


# Subset the ortholog table to only genes that we can "translate"
dictionary <- ortholog_df_1to1 %>%
  as_tibble() %>%
  ## Intersection
  filter(ms_ensemblID %in% unique(rownames(ms.counts)) & hu_symbol %in% unique(rownames(hu.counts))) %>%
  ## Remove any symbols that are NAs
  filter(!is.na(ms_ensemblID)) %>%
  filter(!is.na(hu_symbol)) #%>%
  ## Keep only 1:1 orthologs
  # group_by(ms_ensemblID) %>% filter(n_distinct(hu_symbol) == 1) %>% ungroup() %>%
  # group_by(hu_symbol) %>% filter(n_distinct(ms_ensemblID) == 1) %>% ungroup()
dim(dictionary) # 11,098 genes


# Translate the mouse HVGs into "human gene" language
ms.hvg.translated <- pull(dictionary %>% filter(ms_ensemblID %in% ms.hvg), # not all ms HVGs are found in ortholog.df
                          hu_symbol)
hu.hvg.translated <- pull(dictionary %>% filter(hu_symbol %in% hu.hvg), # not all hu HVGs are found in ortholog.df
                          hu_symbol)
total.hvg <- unique(union(ms.hvg.translated, hu.hvg.translated))
length(total.hvg) # 4,771 (if 5000 HVGs each)
table(total.hvg %in% unique(ms.hvg.translated), useNA="ifany") # 2,864 come from mouse
table(total.hvg %in% unique(hu.hvg.translated), useNA="ifany") # 3,075 come from human => overlap of 1,168 common HVGs


# Keep only ms and hu genes that have 1:1 orthologs
table(unique(rownames(ms.counts)) %in% dictionary$ms_ensemblID) # 11,098 genes should have a translation
table(unique(rownames(hu.counts)) %in% dictionary$hu_symbol) # 11,098 genes should have a translation
ms.counts <- ms.counts[rownames(ms.counts) %in% dictionary$ms_ensemblID,]
hu.counts <- hu.counts[rownames(hu.counts) %in% dictionary$hu_symbol,]
nrow(ms.counts)==nrow(hu.counts) # should be TRUE


# Translate the mouse genes in count table into "human gene"
ms.dict <- dictionary %>%
  filter(ms_ensemblID %in% rownames(ms.counts)) %>%
  select(ms_ensemblID, hu_symbol) %>%
  # distinct() %>%
  # group_by(ms_ensemblID) %>% filter(n_distinct(hu_symbol)>1)
  distinct(ms_ensemblID, .keep_all=T)
ms.dict <- ms.dict[match(rownames(ms.counts), ms.dict$ms_ensemblID),]
table(ms.dict$ms_ensemblID == rownames(ms.counts)) # should be all TRUE
table(is.na(ms.dict$hu_symbol)) # should have no NAs
# Translate
rownames(ms.counts) <- ms.dict$hu_symbol


# Merge everything into one
ms.metadata$study <- "Mouse"
ms.metadata <- ms.metadata[,c("gd_clusters_final", "study")]
colnames(ms.metadata)[1] <- "clusters_GDT"
head(ms.metadata)

hu.metadata$study <- "Human"
hu.metadata <- hu.metadata[,c("clusters_per_lineage", "study")]
colnames(hu.metadata)[1] <- "clusters_GDT"
head(hu.metadata)

ms.counts <- ms.counts[rownames(hu.counts),]
table(rownames(ms.counts)==rownames(hu.counts), useNA="ifany")
table(colnames(ms.metadata)==colnames(hu.metadata), useNA="ifany")
tot.counts   <- cbind(ms.counts, hu.counts)
tot.metadata <- rbind(ms.metadata, hu.metadata)
table(colnames(tot.counts)==rownames(tot.metadata), useNA="ifany")

# Convert seurat count matrix to SummarizedExperiment object
count <- SummarizedExperiment(assays = tot.counts,
                              colData = tot.metadata)
```


## Run metaneighbor
```{r gdt-run-metaneighbor}
# Run metaneighbor
mtn <- MetaNeighborUS(var_genes=total.hvg,
                      dat=count,
                      study_id=count$study,
                      cell_type=count$clusters_GDT,
                      fast_version=T)

# Plot heatmap
# pdf(
#   file = "./data/figures/chapter_04/figs_unfinished/suppfigures/gdt_crosspecies_metaneighbor_heatmap.pdf",
#   width = 5.5,
#   height = 5.5
#   )
heatmap.2(
  mtn,
  trace = "none",
  # superimpose a density histogram on color key
  density.info = "none",
  # color scale
  col = rev(colorRampPalette(brewer.pal(11, "RdYlBu"))(100)),
  breaks = seq(0, 1, length = 101),
  key.xlab = "AUROC",
  key.title = "",
  # text labels
  main = "Mouse vs Human GDT",
  cexRow = 0.6,
  cexCol = 0.6,
  # margins
  margins = c(7, 7)
)
# dev.off()
```


## Make nice figure out of metaneighbor
```{r gdt-plot-metaneighbor, fig.height=7, fig.width=7}
plot_mtn(
  mtn_output = mtn,
  var1 = "Human",
  var2 = "Mouse",
  var1_nbcells = nbcells_hu,
  var2_nbcells = nbcells_ms,
  var1_order = hu_levels,
  var2_order = ms_levels,
  bp_ymax = 50,
  # file_name = "chap4_fig5_gdt_metaneighbor.pdf",
  figsize = 8
)
```




# Sanchez-Sanchez human $\gamma\delta$ data

```{r}
seur_sanchez <- list(
  "gdt_fetal"=readRDS("./data/seurat_objects/sanchez2023/GSE180059_FT_Seurat.rds"),
  "gdt_pnt"=readRDS("./data/seurat_objects/sanchez2023/GSE180059_PNT_Seurat.rds")
)

# Plot
DimPlot(seur_sanchez$gdt_fetal, group.by="seurat_clusters", label=T)
DimPlot(seur_sanchez$gdt_pnt)

# metadata
seur_sanchez$gdt_fetal@meta.data
```





# SESSION INFO

```{r}
sessionInfo()
```
