---
title: "Chapter 2 - Tinn cells and agonist signaling"
author: "Salomé Carcy"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true

---


# IMPORT

## Import librairies
```{r import-librairies}
library(ggplot2)
library(RColorBrewer)
library(cowplot)
library(tidyverse)
library(dplyr)
library(Seurat)
library(SCpubr)
library(patchwork)

source("~/Projects/phd/scripts/colors_universal.R")
```

## Import data
```{r import-data}
seur_integrated <- readRDS("./data/seurat_objects/seurat_human_integrated_object_23_12_01.rds")

seur_thym <- list(
  "cd4"=readRDS("./data/seurat_objects/thymus_objects/seurat_filtered_harmony_02_15_23_thymus.cd4.RDS"),
  "cd8"=readRDS("./data/seurat_objects/thymus_objects/seurat_filtered_harmony_02_15_23_thymus.cd8.RDS"),
  "nkt"=readRDS("./data/seurat_objects/thymus_objects/seurat_filtered_harmony_02_15_23_thymus.nkt.RDS"),
  "mait"=readRDS("./data/seurat_objects/thymus_objects/seurat_filtered_harmony_02_15_23_thymus.mait.RDS"),
  "gdt"=readRDS("./data/seurat_objects/thymus_objects/seurat_filtered_harmony_02_15_23_thymus.gd.RDS")
)
```




# PREPARE DATA

## Clean up data
Do a bit of cleanup
```{r}
# sanity check
# table(rownames(seur_integrated@meta.data[rownames(seur_integrated@meta.data)%in%rownames(seur_thym$mait@meta.data),]) == rownames(seur_thym$mait@meta.data), useNA="ifany"
# )

# clean up metadata columns
# colnames(seur_thym$cd8@meta.data)
seur_thym <- lapply(
  seur_thym,
  function(x){
    x@meta.data[,13:77] <- NULL
    return(x)
    }
)

# add column "clusters per lineage"
seur_thym <- lapply(
  seur_thym,
  function(x){
    x@meta.data$clusters_per_lineage <- seur_integrated@meta.data[rownames(seur_integrated@meta.data)%in%rownames(x@meta.data), "clusters_per_lineage"]
    return(x)
    }
  )
```


## Gene signatures
```{r gene-signatures}
# Get Park et al. gene signatures
park_markergenes <- read.csv("~/Projects/phd/data/litterature_gene_signatures/park_genesignatures.csv", row.names=1)
park_gene_signatures <- c("DPp", "DPq", "CD8aa1", "CD8aa2", "T_agonist", "abT_entry", "NKT", "CD4T", "CD8T")
park_gene_signatures_list <- as.list(park_markergenes[,park_gene_signatures])
names(park_gene_signatures_list) <- paste0("park_", names(park_gene_signatures_list))

# Get Chopp et al. gene signatures
chopp_signatures <- read.csv("~/Projects/phd/data/litterature_gene_signatures/chopp_gene_signatures.csv")
chopp_gene_signatures_list <- as.list(chopp_signatures)
names(chopp_gene_signatures_list) <- paste0("chopp_", names(chopp_gene_signatures_list))

# Personally curated gene signatures from litterature
perso_gene_signatures <- list(
  "effector"=c("HOPX", "GZMB", "NKG7", "TBX21", "PRF1", "GZMA", "KLRD1", "EOMES", "CCR6", "RORC", "JUNB", "FOS", "RORA", "FOSB"),
  "naive"   =c("SATB1", "TCF7", "LEF1", "CCR7", "SELL", "FOXP1", "KLF2", "SOX4", "ID3", "BACH2"),
  "egress"  =c("KLF2", "CORO1A", "CCR7", "CXCR4", "CXCR6", "FOXO1", "CXCR3", "S1PR1", "S1PR4", "S100A4", "S100A6", "EMP3")
  )


# Combined gene lists
gene_signatures_list <- c(park_gene_signatures_list, chopp_gene_signatures_list, perso_gene_signatures)
gene_signatures_list <- lapply(gene_signatures_list, function(x) x[x!=""]) # remove empty genes
gene_signatures_list <- lapply(gene_signatures_list, function(x) x[x %in% rownames(seur_integrated)]) # remove genes not found in our data
lapply(gene_signatures_list, function(x) length(x)) # sanity check (between 0-5 genes per signature lost)


# add module score to integrated object (better for choice of control genes)
seur_integrated <- AddModuleScore(
  object = seur_integrated,
  assay = "RNA",
  features = gene_signatures_list,
  name=names(gene_signatures_list),
  seed=1
  )
colnames(seur_integrated@meta.data)[32:46] <- names(gene_signatures_list)

# little sanity check
# FeaturePlot(seur_integrated, reduction="umap_integrated", order=T, features=c("score_naive", "naive"))
# FeaturePlot(seur_integrated, reduction="umap_integrated", order=T, features=c("score_effector", "effector"))
# FeaturePlot(seur_integrated, reduction="umap_integrated", order=T, features=c("score_egress", "egress"))
# FeaturePlot(seur_integrated, reduction="umap_integrated", order=T, features=c("score_cd8aa", "park_CD8aa1"))
# FeaturePlot(seur_integrated, reduction="umap_integrated", order=T, features=c("chopp_cd4_signature", "chopp_cd8_signature"), split.by="tissue")

# add scores to all individual thymic seurat objects
seur_thym <- lapply(
  seur_thym,
  function(x){
    x@meta.data[,names(gene_signatures_list)] <- seur_integrated@meta.data[rownames(seur_integrated@meta.data)%in%rownames(x@meta.data), names(gene_signatures_list)]
    return(x)
    }
  )
```




# FUNCTIONS
```{r define-functions}
lineage_dimplot <- function(seurobj, col_vector, file_name="no"){
  p <- do_DimPlot(
    seurobj,
    group.by="clusters_per_lineage",
    colors.use = col_vector,
    legend.position = "right",
    # raster = T,
    # raster.dpi=2048,
    pt.size = 0.5#,
    # border.size=1.5
  )+
    theme(
      # axis.title=element_text(size=8),
      # axis.text=element_text(size=6),
      legend.key.size = unit(0.4, "cm"),
      legend.title=element_text(size=8),
      legend.text = element_text(size=6)
    )
  if(file_name != "no"){
    ggsave(paste0("~/Projects/phd/data/figures/chapter_02/figs_unfinished/section2/", file_name),
           plot=p,
           width=4, height=2.5, units="in")
  }
}

lineage_genesignature <- function(seuratobj, genesignature, file_name="no"){
  p <- do_FeaturePlot(
    seuratobj,
    features = genesignature,
    min.cutoff=0,
    use_viridis = T,
    viridis.palette = "F",
    legend.position="right",
    order=T,
    pt.size = 1.5,
    raster=F
    # raster.dpi=2048,
    # pt.size=10
  )
  if(file_name != "no"){
    ggsave(paste0("~/Projects/phd/data/figures/chapter_02/figs_unfinished/section2/", file_name),
           plot=p,
           width=7, height=5)
  }
}

lineage_nebulosagenes <- function(seuratobj, genes_vector, pgrid_ncol=2, pgrid_size=c(10,30), file_name="no"){
  
  pnebulosa <- list()
  for(gene in genes_vector) {
    p <- SCpubr::do_NebulosaPlot(seuratobj, features = gene)
    pnebulosa[[gene]] <- p
  }
  pgrid <- plot_grid(plotlist=pnebulosa, ncol=pgrid_ncol)
  
  if(file_name != "no"){
    ggsave(paste0("~/Projects/phd/data/figures/chapter_02/figs_unfinished/section2/", file_name),
           plot=pgrid,
           width=pgrid_size[1], height=pgrid_size[2])
  }
}

lineage_dotplot <- function(seuratobj, genes_vector, lines_coord, p_size=c(5,7.5), file_name="no"){
  p <- do_DotPlot(
    sample=seuratobj,
    features=genes_vector,
    group.by = "clusters_per_lineage",
    legend.position = "right",
    use_viridis = T,
    viridis.palette = "B",
    legend.title="z-score avg expression",
    flip=T,
    scale=T,
  )+
    geom_hline(yintercept=lines_coord, linetype="dashed", color="grey")+
    theme(
      legend.title=element_text(size=8),
      legend.text = element_text(size=6),
      axis.text.x=element_text(size=8, angle=45, hjust=1),
      axis.text.y=element_text(size=8, face="italic")
    )
  
  if(file_name != "no"){
    ggsave(paste0("~/Projects/phd/data/figures/chapter_02/figs_unfinished/section2/", file_name),
           plot=p,
           width=p_size[1], height=p_size[2])
  }
}
```




# iNKT FIGURE

## DimPlot
```{r inkt-umap}
lineage_dimplot(seurobj = seur_thym$nkt, col_vector = cols_thym_nkt, file_name = "ch2_fig10_inkt_umap.pdf")
```


## DotPlot with marker genes
```{r inkt-dotplot, fig.width=10, fig.height=7}
Idents(seur_thym$nkt) <- "clusters_per_lineage"
inkt_markergenes <- FindAllMarkers(
  seur_thym$nkt,
  test.use = 'wilcox',
  logfc.threshold = 0.4,
  min.pct = 0.3,
  only.pos = TRUE
  )
inkt_markergenes_list <- inkt_markergenes %>%
  mutate(cluster=as.character(cluster)) %>%
  arrange(cluster) %>%
  filter(p_val_adj<0.05) %>%
  group_by(cluster) %>%
  top_n(-5, p_val_adj) %>%
  # top_n(5, avg_log2FC) %>%
  ungroup() %>%
  # filter(avg_log2FC>1.5) %>%
  pull(gene) %>% unique()

# dotplot
do_DotPlot(
  sample=seur_thym$nkt,
  features=unique(inkt_markergenes_list),
  group.by = "clusters_per_lineage",
  legend.position = "right",
  use_viridis = T,
  viridis.palette = "B",
  legend.title="z-score avg expression",
  # flip=T,
  scale=T,
  )+
  geom_vline(xintercept=c(5.5, 10.5, 15.5, 20.5, 23.5, 29.5), linetype="dashed", color="grey")+
  theme(
    legend.title=element_text(size=8),
    legend.text = element_text(size=6),
    axis.text.x=element_text(size=8, angle=45, hjust=1, face="italic"),
    axis.text.y=element_text(size=8)
    )
# ggsave("~/Projects/phd/data/figures/chapter_02/figs_unfinished/section2/ch2_fig10_inkt_markergenes.pdf", width=9, height=5)
```


## Gene signatures
```{r inkt-cd8aa, fig.height=5, fig.width=10}
lineage_genesignature(seuratobj = seur_thym$nkt, genesignature = "park_CD8aa1", file_name = "ch2_fig10_inkt_cd8aa_sig.pdf")
lineage_genesignature(seuratobj = seur_thym$nkt, genesignature = "park_abT_entry", file_name = "ch2_fig10_inkt_SPentry_sig.pdf")
lineage_genesignature(seuratobj = seur_thym$nkt, genesignature = "naive", file_name = "ch2_fig10_inkt_naive.pdf")
lineage_genesignature(seuratobj = seur_thym$nkt, genesignature = "effector", file_name = "ch2_fig10_inkt_effector.pdf")
lineage_genesignature(seuratobj = seur_thym$nkt, genesignature = "egress", file_name = "ch2_fig10_inkt_egress.pdf")
```


## Nebulosa plot highlight CCR9 and CCR7
```{r inkt-umap-densityplots, fig.height=10, fig.width=10}
inkt_genes_nebulosa <- c(
    # "CD4", "CD8A", "CD8B",
    "EGR2", "ZBTB16",
    "CCR9", "CCR7",
    "TBX21", "EOMES",
    "RORC", "RORA",
    "GZMK", "PRF1",
    "KLRB1", "NKG7"
    )

lineage_nebulosagenes(seuratobj = seur_thym$nkt, genes_vector = inkt_genes_nebulosa, pgrid_ncol = 2, pgrid_size = c(10,30), file_name = "ch2_fig10_inkt_densityplots.pdf")
```


## iNKT DotPlot genes of interest
```{r inkt-dotplot-genes-relevant}
inkt_genes_dotplot <- c(
  "CD4", "CD8A", "CD8B", 
  "GNG4", "PDCD1", "TNFRSF9", # CD8aa
  "HIVEP3", "EGR2", "ZBTB16", # TFs induced in iNKT cells post-selection
  "CCR9", "LEF1", "ID3", # naive CCR9
  "CCR7", "SELL", "TCF7", # naive CCR7
  "S1PR1", "KLF2", # egress
  "FOS", "JUN", "JUNB", # AP1 signaling
  "TBX21", "EOMES", "IFNG", "GZMK", # type 1 effector
  "RORC", "RORA", "CCR6", "IL23R", # type 3 effector
  "KLRB1"
)

lineage_dotplot(
  seuratobj = seur_thym$nkt,
  genes_vector = inkt_genes_dotplot,
  lines_coord = c(3.5, 6.5, 9.5, 15.5, 17.5, 20.5),
  p_size = c(5, 7.5),
  file_name = "ch2_fig10_inkt_dotplot2.pdf"
)
```


## test TCR shannon entropy
```{r}
tcr_df <- seur_thym$nkt@meta.data %>%
  


TCR_data_NKT_cells_TCRb_diversity_thymus.df <- TCR_data_NKT_cells.df %>%
  dplyr::filter(TRAV == "TRAV10*01") %>%
  mutate(TRBV = str_remove(TRBV, pattern = "\\-[^.]*$"),
         TRBV = str_remove(TRBV, pattern = "\\*[^.]*$"),
         TRBJ = str_remove(TRBJ, pattern = "\\*[^.]*$")) %>%
  dplyr::select(-TRAV, -TRAJ, -TRBD, -CDR3a)

TCR_data_NKT_cells_TCRb_diversity_thymus_naive.df <- TCR_data_NKT_cells_TCRb_diversity_thymus.df %>%
  dplyr::filter(new_clusters_NKT %in% c("1", "2", "3")) %>%
  mutate(TCRb_d_clonotype = paste(TRBV, TRBJ, CDR3b, sep = "_")) %>%
  dplyr::select(-new_clusters_NKT) %>% group_by(TCRb_d_clonotype) %>%
  summarise(n = n()) %>% arrange(desc(n)) %>% pivot_wider(names_from = TCRb_d_clonotype, values_from = n)

shannon_iteration <- function(){
  TCR_data_NKT_cells_TCRb_diversity_thymus_naive.df %>% 
    rrarefy(sample = 23) %>% 
    diversity(index = "shannon")
}

mean_shannon_thymus_naive <- mean(replicate(100, shannon_iteration()))

TCR_data_NKT_cells_TCRb_diversity_thymus_C5.df <- TCR_data_NKT_cells_TCRb_diversity_thymus.df %>%
  dplyr::filter(new_clusters_NKT %in% c("5")) %>%
  mutate(TCRb_d_clonotype = paste(TRBV, TRBJ, CDR3b, sep = "_")) %>%
  dplyr::select(-new_clusters_NKT) %>% group_by(TCRb_d_clonotype) %>%
  summarise(n = n()) %>% arrange(desc(n)) %>% pivot_wider(names_from = TCRb_d_clonotype, values_from = n)

shannon_iteration <- function(){
  TCR_data_NKT_cells_TCRb_diversity_thymus_C5.df %>% 
    rrarefy(sample = 23) %>% 
    diversity(index = "shannon")
}

mean_shannon_thymus_C5 <- mean(replicate(100, shannon_iteration()))

TCR_data_NKT_cells_TCRb_diversity_thymus_C6.df <- TCR_data_NKT_cells_TCRb_diversity_thymus.df %>%
  dplyr::filter(new_clusters_NKT %in% c("6")) %>%
  mutate(TCRb_d_clonotype = paste(TRBV, TRBJ, CDR3b, sep = "_")) %>%
  dplyr::select(-new_clusters_NKT) %>% group_by(TCRb_d_clonotype) %>%
  summarise(n = n()) %>% arrange(desc(n)) %>% pivot_wider(names_from = TCRb_d_clonotype, values_from = n)

shannon_iteration <- function(){
  TCR_data_NKT_cells_TCRb_diversity_thymus_C6.df %>% 
    rrarefy(sample = 23) %>% 
    diversity(index = "shannon")
}

mean_shannon_thymus_C6 <- mean(replicate(100, shannon_iteration()))
```




# MAIT FIGURE

## DimPlot
```{r mait-umap}
lineage_dimplot(seurobj = seur_thym$mait, col_vector = cols_thym_mait, file_name = "ch2_fig10_mait_umap.pdf")
```

## DotPlot with marker genes
```{r mait-dotplot, fig.width=10, fig.height=7}
Idents(seur_thym$mait) <- "clusters_per_lineage"
mait_markergenes <- FindAllMarkers(
  seur_thym$mait,
  test.use = 'wilcox',
  logfc.threshold = 0.4,
  min.pct = 0.3,
  only.pos = TRUE
  )
mait_markergenes_list <- mait_markergenes %>%
  mutate(cluster=as.character(cluster)) %>%
  arrange(cluster) %>%
  filter(p_val_adj<0.05) %>%
  filter(!gene %in% grep("RPL|RPS", gene, value=T)) %>%
  group_by(cluster) %>%
  top_n(-5, p_val_adj) %>%
  # top_n(5, avg_log2FC) %>%
  ungroup() %>%
  # filter(avg_log2FC>1.5) %>%
  pull(gene) %>% unique()

# dotplot
do_DotPlot(
  sample=seur_thym$mait,
  features=unique(mait_markergenes_list),
  group.by = "clusters_per_lineage",
  legend.position = "right",
  use_viridis = T,
  viridis.palette = "B",
  legend.title="z-score avg expression",
  # flip=T,
  scale=T,
  )+
  geom_vline(xintercept=c(5.5, 10.5, 15.5, 20.5, 25.5, 30.5), linetype="dashed", color="grey")+
  theme(
    legend.title=element_text(size=8),
    legend.text = element_text(size=6),
    axis.text.x=element_text(size=8, angle=45, hjust=1, face="italic"),
    axis.text.y=element_text(size=8)
    )
# ggsave("~/Projects/phd/data/figures/chapter_02/figs_unfinished/section2/ch2_fig10_mait_markergenes.pdf", width=10, height=5)
```


## Gene signatures
```{r mait-gene-signatures, fig.height=5, fig.width=10}
lineage_genesignature(seuratobj = seur_thym$mait, genesignature = "park_CD8aa1", file_name = "ch2_fig10_mait_cd8aa_sig.pdf")
lineage_genesignature(seuratobj = seur_thym$mait, genesignature = "park_abT_entry", file_name = "ch2_fig10_mait_SPentry_sig.pdf")
lineage_genesignature(seuratobj = seur_thym$mait, genesignature = "naive", file_name = "ch2_fig10_mait_naive.pdf")
lineage_genesignature(seuratobj = seur_thym$mait, genesignature = "effector", file_name = "ch2_fig10_mait_effector.pdf")
lineage_genesignature(seuratobj = seur_thym$mait, genesignature = "egress", file_name = "ch2_fig10_mait_egress.pdf")
```


## Nebulosa plot highlight CCR9 and CCR7
```{r mait-umap-densityplots, fig.height=10, fig.width=10}
lineage_nebulosagenes(seuratobj = seur_thym$mait, genes_vector = inkt_genes_nebulosa, pgrid_ncol = 2, pgrid_size = c(10,30), file_name = "ch2_fig10_mait_densityplots.pdf")
```


## iNKT DotPlot genes of interest
```{r inkt-dotplot-genes-relevant}
# inkt_genes_dotplot <- c(
#   "CD4", "CD8A", "CD8B", 
#   "GNG4", "PDCD1", "TNFRSF9", # CD8aa
#   "HIVEP3", "EGR2", "ZBTB16", # TFs induced in iNKT cells post-selection
#   "CCR9", "LEF1", "ID3", # naive CCR9
#   "CCR7", "SELL", "TCF7", # naive CCR7
#   "S1PR1", "KLF2", # egress
#   "FOS", "JUN", "JUNB", # AP1 signaling
#   "TBX21", "EOMES", "IFNG", "GZMK", # type 1 effector
#   "RORC", "RORA", "CCR6", "IL23R", # type 3 effector
#   "KLRB1"
# )

lineage_dotplot(
  seuratobj = seur_thym$mait,
  genes_vector = c(inkt_genes_dotplot),
  lines_coord = c(3.5, 6.5, 9.5, 15.5, 17.5, 20.5),
  p_size = c(5, 7.5),
  file_name = "ch2_fig10_mait_dotplot.pdf"
)
```






## Analysis 2
```{r analysis-2, fig.height=6, fig.width=10}
Idents(seur_thym$cd8) <- "clusters_per_lineage"
cd8_markergenes <- FindAllMarkers(
  seur_thym$cd8,
  test.use = 'wilcox',
  logfc.threshold = 0.4,
  min.pct = 0.3,
  only.pos = TRUE
  )
cd8_markergenes_list <- cd8_markergenes %>%
  mutate(cluster=as.character(cluster)) %>%
  arrange(cluster) %>%
  filter(p_val_adj<0.05) %>%
  filter(!gene %in% grep("RPL|RPS", gene, value=T)) %>%
  group_by(cluster) %>%
  top_n(-5, p_val_adj) %>%
  # top_n(5, avg_log2FC) %>%
  ungroup() %>%
  # filter(avg_log2FC>1.5) %>%
  pull(gene) %>% unique()

# dotplot
do_DotPlot(
  sample=seur_thym$cd8,
  features=c(unique(cd8_markergenes_list), "CD4"),
  group.by = "clusters_per_lineage",
  legend.position = "right",
  use_viridis = T,
  viridis.palette = "B",
  legend.title="z-score avg expression",
  # flip=T,
  scale=T,
  )+
  geom_vline(xintercept=c(5.5, 10.5, 15.5, 20.5, 25.5, 30.5), linetype="dashed", color="grey")+
  theme(
    legend.title=element_text(size=8),
    legend.text = element_text(size=6),
    axis.text.x=element_text(size=8, angle=45, hjust=1, face="italic"),
    axis.text.y=element_text(size=8)
    )
```

## Analysis 3
```{r analysis-3}

```


## Analysis 4
```{r analysis-4}

```




# SESSION INFO
```{r}
sessionInfo()
```


