---
title: "Chapter 2 - Figure 8 iNKT thymocytes"
author: "Salomé Carcy"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true

---

This notebook is for figures 8 and 10 in chapter 2.

```{r, setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 5, fig.height = 5,
                      warning=FALSE, message=FALSE,
                      root.dir = "~/Projects/phd/")
```


# IMPORT

## Import librairies
```{r import-librairies}
library(ggplot2)
library(RColorBrewer)
library(cowplot)
library(tidyverse)
library(dplyr)
library(Seurat)
library(SCpubr)
library(patchwork)

source("~/Projects/phd/scripts/colors_universal.R")
```

## Import data
```{r import-data}
# seur_integrated <- readRDS("./data/seurat_objects/seurat_human_integrated_object_23_12_01.rds")

seur_thym <- list(
  "cd4"=readRDS("./data/seurat_objects/thymus_objects/seurat_thymus_cd4_with_gene_signatures.rds"),
  "cd8"=readRDS("./data/seurat_objects/thymus_objects/seurat_thymus_cd8_with_gene_signatures.rds"),
  "nkt"=readRDS("./data/seurat_objects/thymus_objects/seurat_thymus_nkt_with_gene_signatures.rds"),
  "mait"=readRDS("./data/seurat_objects/thymus_objects/seurat_thymus_mait_with_gene_signatures.rds"),
  "gdt"=readRDS("./data/seurat_objects/thymus_objects/seurat_thymus_gdt_with_gene_signatures.rds")
)
```




# FUNCTIONS
```{r define-functions}
lineage_dimplot <- function(seurobj, col_vector, file_name="no"){
  p <- do_DimPlot(
    seurobj,
    group.by="clusters_per_lineage",
    colors.use = col_vector,
    legend.position = "right",
    # raster = T,
    # raster.dpi=2048,
    pt.size = 0.5#,
    # border.size=1.5
  )+
    theme(
      # axis.title=element_text(size=8),
      # axis.text=element_text(size=6),
      legend.key.size = unit(0.4, "cm"),
      legend.title=element_text(size=8),
      legend.text = element_text(size=6)
    )
  if(file_name != "no"){
    ggsave(paste0("~/Projects/phd/data/figures/chapter_02/figs_unfinished/section2/", file_name),
           plot=p,
           width=4, height=2.5, units="in")
  }
}


lineage_clustersperdonor <- function(seurobj, col_vector, file_name="no", figwidth=6){
  p <- seurobj@meta.data %>%
    group_by(Donor, clusters_per_lineage) %>%
    count() %>%
    ggplot(aes(x=Donor, y=n, fill=clusters_per_lineage))+
    geom_bar(stat="identity", position="fill")+
    scale_fill_manual(values=col_vector)+
    labs(x="Donor", y="% cells", fill="")+
    theme_cowplot()#+
  # theme(axis.text=element_text(size=5/.pt),
  #       axis.title=element_text(size=7/.pt),
  #       legend.text = element_text(size=5/.pt))
  if(file_name != "no"){
    ggsave(paste0("~/Projects/phd/data/figures/chapter_02/figs_unfinished/section2/", file_name),
           plot=p,
           width=figwidth, height=5, units="in")
  }
}


lineage_genesignature <- function(seuratobj, genesignature, file_name="no"){
  p <- do_FeaturePlot(
    seuratobj,
    features = genesignature,
    min.cutoff=0,
    use_viridis = T,
    viridis.palette = "E",
    legend.position="right",
    order=T,
    pt.size = 1.5,
    raster=F
    # raster.dpi=2048,
    # pt.size=10
  )
  if(file_name != "no"){
    ggsave(paste0("~/Projects/phd/data/figures/chapter_02/figs_unfinished/section2/", file_name),
           plot=p,
           width=7, height=5)
  }
  return(p)
}


lineage_nebulosagenes <- function(seuratobj, genes_vector, pgrid_ncol=2, pgrid_size=c(10,30), file_name="no"){
  pnebulosa <- list()
  for(gene in genes_vector) {
    p <- SCpubr::do_NebulosaPlot(seuratobj, features = gene, sequential.palette="YlOrRd")
    p <- ggrastr::rasterise(p, layers="Point", dpi=200)
    pnebulosa[[gene]] <- p
  }
  pgrid <- plot_grid(plotlist=pnebulosa, ncol=pgrid_ncol)
  print(pgrid)
  
  if(file_name != "no"){
    ggsave(paste0("~/Projects/phd/data/figures/chapter_02/figs_unfinished/section2/", file_name),
           plot=pgrid,
           width=pgrid_size[1], height=pgrid_size[2])
  }
}


plot_DotPlot <- function(seurobj, group, features, scaling=T){
  # get plot data
  p <- Seurat::DotPlot(
    seurobj,
    group.by=group,
    features=features,
    scale=scaling
  )
  # plot in personalized way
  p <- ggplot(p$data, aes(x=id, y=features.plot, fill=avg.exp.scaled, size=pct.exp))+
    geom_point(color="black", shape=21)+
    # scale_fill_gradient2(low=scales::muted("blue"), high=scales::muted("red"), name="z-score\nnormalized\navg expression")+
    scale_size_continuous(range=c(0,6), limits=c(0,100), name="%cells\nexpressing\ngene")+
    theme_bw()+
    theme(
      axis.text.y=element_text(face="italic"),
      axis.text.x=element_text(angle=45, hjust=1)
      )+
    labs(y="", x="")
  # different color scale if scaled or not
  if(scaling==T){
    p <- p + scale_fill_gradient2(low=scales::muted("blue"), high=scales::muted("red"), name="z-score\naverage\nnormalized\nexpression")
  } else{
    p <- p + viridis::scale_fill_viridis(option="B", direction=-1, name="average\nnormalized\nexpression")
  }
  return(p)
}
```




# iNKT FIGURE

## DimPlot
```{r inkt-umap}
lineage_dimplot(seurobj = seur_thym$nkt, col_vector = cols_thym_nkt, file_name = "ch2_fig8_inkt_umap.pdf")

# cluster comp per donor
lineage_clustersperdonor(seurobj = seur_thym$nkt, col_vector = cols_thym_nkt, file_name = "ch2_fig8_inkt_clustersperdonor.pdf")
```


## Gene signatures
```{r inkt-cd8aa, fig.height=5, fig.width=10}
lineage_genesignature(seuratobj = seur_thym$nkt, genesignature = "park_CD8aa1", file_name = "ch2_fig8_inkt_cd8aa_sig.pdf")
lineage_genesignature(seuratobj = seur_thym$nkt, genesignature = "park_abT_entry", file_name = "ch2_fig8_inkt_SPentry_sig.pdf")
lineage_genesignature(seuratobj = seur_thym$nkt, genesignature = "naive", file_name = "ch2_fig8_inkt_naive.pdf")
lineage_genesignature(seuratobj = seur_thym$nkt, genesignature = "effector", file_name = "ch2_fig8_inkt_effector.pdf")
lineage_genesignature(seuratobj = seur_thym$nkt, genesignature = "egress", file_name = "ch2_fig8_inkt_egress.pdf")
```


## Nebulosa plot highlight genes of interest
```{r inkt-umap-densityplots, fig.height=10, fig.width=5}
inkt_genes_nebulosa <- c(
    # "CD4", "CD8A", "CD8B",
    "EGR2", "ZBTB16",
    "CCR9", "CCR7",
    "TBX21", "EOMES",
    "RORC", "RORA",
    "GZMK", "PRF1",
    "KLRB1", "NKG7"
    )

lineage_nebulosagenes(seuratobj = seur_thym$nkt, genes_vector = inkt_genes_nebulosa, pgrid_ncol = 2, pgrid_size = c(10,30), file_name = "ch2_fig8_inkt_densityplots.pdf")
```


## iNKT DotPlot genes of interest
```{r inkt-dotplot-genes-relevant, fig.height=8, fig.width=5}
inkt_genes_dotplot <- c(
  "CD4", "CD8A", "CD8B", 
  "GNG4", "PDCD1", "TNFRSF9", # CD8aa
  "HIVEP3", "EGR2", "ZBTB16", # TFs induced in iNKT cells post-selection
  "CCR9", "LEF1", "ID3", # naive CCR9
  "CCR7", "SELL", "TCF7", # naive CCR7
  "S1PR1", "KLF2", # egress
  "FOS", "JUN", "JUNB", # AP1 signaling
  "TBX21", "EOMES", "IFNG", "GZMK", "GZMB", # type 1 effector
  "RORC", "RORA", "CCR6", "IL23R", # type 3 effector
  "KLRB1"
)

plot_DotPlot(
  seurobj = seur_thym$nkt,
  group="clusters_per_lineage",
  features = inkt_genes_dotplot,
  scaling=T
  # file_name = "ch2_fig8_inkt_dotplot2.pdf"
)
# ggsave("~/Projects/phd/data/figures/chapter_02/figs_unfinished/section2/ch2_fig8_inkt_dotplot.pdf", width=4, height=8)
```


iNKT dotplot genes of interest for presentation:
```{r inkt-dotplot-genes-thesis-defense}
plot_DotPlot(
  seurobj = seur_thym_hu$nkt,
  group = "clusters_per_lineage",
  features = c(
  "CD4", "CD8A", "CD8B", 
  "ZBTB16",
  # "GNG4", "PDCD1", "TNFRSF9", # CD8aa
  # "HIVEP3", "EGR2", "ZBTB16", # TFs induced in iNKT cells post-selection
  "CCR9", "CCR7",
  "SELL", "TCF7","LEF1", "SATB1", "FOXP1","SOX4", "BACH2",
  "S1PR1", "KLF2", # egress
  "KLRB1"
  # "FOS", "JUN", "JUNB", # AP1 signaling
  # "TBX21", "EOMES", "IFNG", "GZMK", # type 1 effector
  # "RORC", "RORA", "CCR6", "IL23R", # type 3 effector
  # "KLRB1"
),
  scaling=F
  )
# ggsave("./data/presentation/figs_unfinished/nkt_thymus_dotplot_naive.pdf", width=5, height=6)


plot_DotPlot(
  seurobj = seur_thym_hu$nkt,
  group = "clusters_per_lineage",
  features = c(
  "CD4", "CD8A", "CD8B", 
  "ZBTB16",
  # "GNG4", "PDCD1", "TNFRSF9", # CD8aa
  # "HIVEP3", "EGR2", "ZBTB16", # TFs induced in iNKT cells post-selection
  # "CCR9", "CCR7",
  # "SELL", "TCF7","LEF1", "SATB1", "FOXP1","SOX4", "BACH2",
  # "S1PR1", "KLF2", # egress
  "KLRB1",
  "FOS", "JUN", "JUNB", # AP1 signaling
  "TBX21", "EOMES", "IFNG", "GZMK", # type 1 effector
  "RORC", "RORA", "CCR6", "IL23R" # type 3 effector
  # "KLRB1"
),
  scaling=F
  )
# ggsave("./data/presentation/figs_unfinished/nkt_thymus_dotplot_memory.pdf", width=5, height=6)
```


## Co-expression type 1 and 17 gene signatures
```{r inkt-type1-type17}
df <- rbind(
  as.data.frame(seur_thym$nkt@meta.data) %>%
    select(type1, type17, clusters_per_lineage) %>%
    mutate(tcell_lineage="iNKT"),
  as.data.frame(seur_thym$mait@meta.data) %>%
    select(type1, type17, clusters_per_lineage) %>%
    mutate(tcell_lineage="MAIT"),
  as.data.frame(seur_thym$gdt@meta.data) %>%
    select(type1, type17, clusters_per_lineage) %>%
    mutate(tcell_lineage="GD")
)

ggplot(
  df %>% filter(tcell_lineage=="iNKT"),
  aes(x=type1, y=type17, color=clusters_per_lineage)
)+
  geom_point()+
  scale_color_manual(values=c(cols_thym_nkt, cols_thym_mait, cols_thym_gdt))+
  labs(x="Type 1 signature", y="Type 17 signature")+
  theme_cowplot()
# ggsave("~/Projects/phd/data/figures/chapter_02/figs_unfinished/section2/ch2_fig8_inkt_type1_17_coexpression.pdf", width=6, height=4)
```


## test TCR shannon entropy
```{r inkt-tcr-shannon-entropy, include=FALSE, eval=FALSE}
tcr_df <- seur_thym$nkt@meta.data %>%
  


TCR_data_NKT_cells_TCRb_diversity_thymus.df <- TCR_data_NKT_cells.df %>%
  dplyr::filter(TRAV == "TRAV10*01") %>%
  mutate(TRBV = str_remove(TRBV, pattern = "\\-[^.]*$"),
         TRBV = str_remove(TRBV, pattern = "\\*[^.]*$"),
         TRBJ = str_remove(TRBJ, pattern = "\\*[^.]*$")) %>%
  dplyr::select(-TRAV, -TRAJ, -TRBD, -CDR3a)

TCR_data_NKT_cells_TCRb_diversity_thymus_naive.df <- TCR_data_NKT_cells_TCRb_diversity_thymus.df %>%
  dplyr::filter(new_clusters_NKT %in% c("1", "2", "3")) %>%
  mutate(TCRb_d_clonotype = paste(TRBV, TRBJ, CDR3b, sep = "_")) %>%
  dplyr::select(-new_clusters_NKT) %>% group_by(TCRb_d_clonotype) %>%
  summarise(n = n()) %>% arrange(desc(n)) %>% pivot_wider(names_from = TCRb_d_clonotype, values_from = n)

shannon_iteration <- function(){
  TCR_data_NKT_cells_TCRb_diversity_thymus_naive.df %>% 
    rrarefy(sample = 23) %>% 
    diversity(index = "shannon")
}

mean_shannon_thymus_naive <- mean(replicate(100, shannon_iteration()))

TCR_data_NKT_cells_TCRb_diversity_thymus_C5.df <- TCR_data_NKT_cells_TCRb_diversity_thymus.df %>%
  dplyr::filter(new_clusters_NKT %in% c("5")) %>%
  mutate(TCRb_d_clonotype = paste(TRBV, TRBJ, CDR3b, sep = "_")) %>%
  dplyr::select(-new_clusters_NKT) %>% group_by(TCRb_d_clonotype) %>%
  summarise(n = n()) %>% arrange(desc(n)) %>% pivot_wider(names_from = TCRb_d_clonotype, values_from = n)

shannon_iteration <- function(){
  TCR_data_NKT_cells_TCRb_diversity_thymus_C5.df %>% 
    rrarefy(sample = 23) %>% 
    diversity(index = "shannon")
}

mean_shannon_thymus_C5 <- mean(replicate(100, shannon_iteration()))

TCR_data_NKT_cells_TCRb_diversity_thymus_C6.df <- TCR_data_NKT_cells_TCRb_diversity_thymus.df %>%
  dplyr::filter(new_clusters_NKT %in% c("6")) %>%
  mutate(TCRb_d_clonotype = paste(TRBV, TRBJ, CDR3b, sep = "_")) %>%
  dplyr::select(-new_clusters_NKT) %>% group_by(TCRb_d_clonotype) %>%
  summarise(n = n()) %>% arrange(desc(n)) %>% pivot_wider(names_from = TCRb_d_clonotype, values_from = n)

shannon_iteration <- function(){
  TCR_data_NKT_cells_TCRb_diversity_thymus_C6.df %>% 
    rrarefy(sample = 23) %>% 
    diversity(index = "shannon")
}

mean_shannon_thymus_C6 <- mean(replicate(100, shannon_iteration()))
```




# MAIT FIGURE

## DimPlot
```{r mait-umap}
lineage_dimplot(seurobj = seur_thym$mait, col_vector = cols_thym_mait, file_name = "ch2_fig11_mait_umap.pdf")

lineage_clustersperdonor(seurobj = seur_thym$mait, col_vector = cols_thym_mait, file_name = "ch2_fig11_mait_clustersperdonor.pdf")
```


## DotPlot with marker genes
```{r mait-dotplot, fig.width=10, fig.height=7}
Idents(seur_thym$mait) <- "clusters_per_lineage"
mait_markergenes <- FindAllMarkers(
  seur_thym$mait,
  test.use = 'wilcox',
  logfc.threshold = 0.4,
  min.pct = 0.3,
  only.pos = TRUE
  )
mait_markergenes_list <- mait_markergenes %>%
  mutate(cluster=as.character(cluster)) %>%
  arrange(cluster) %>%
  filter(p_val_adj<0.05) %>%
  filter(!gene %in% grep("RPL|RPS", gene, value=T)) %>%
  group_by(cluster) %>%
  top_n(-5, p_val_adj) %>%
  # top_n(5, avg_log2FC) %>%
  ungroup() %>%
  # filter(avg_log2FC>1.5) %>%
  pull(gene) %>% unique()

# dotplot
do_DotPlot(
  sample=seur_thym$mait,
  features=unique(mait_markergenes_list),
  group.by = "clusters_per_lineage",
  legend.position = "right",
  use_viridis = T,
  viridis.palette = "B",
  legend.title="z-score avg expression",
  # flip=T,
  scale=T,
  )+
  geom_vline(xintercept=c(5.5, 10.5, 15.5, 20.5, 25.5, 30.5), linetype="dashed", color="grey")+
  theme(
    legend.title=element_text(size=8),
    legend.text = element_text(size=6),
    axis.text.x=element_text(size=8, angle=45, hjust=1, face="italic"),
    axis.text.y=element_text(size=8)
    )
# ggsave("~/Projects/phd/data/figures/chapter_02/figs_unfinished/section2/ch2_fig11_mait_markergenes.pdf", width=10, height=5)
```


## Gene signatures
```{r mait-gene-signatures, fig.height=5, fig.width=10}
lineage_genesignature(seuratobj = seur_thym$mait, genesignature = "park_CD8aa1", file_name = "ch2_fig11_mait_cd8aa_sig.pdf")
lineage_genesignature(seuratobj = seur_thym$mait, genesignature = "park_abT_entry", file_name = "ch2_fig11_mait_SPentry_sig.pdf")
lineage_genesignature(seuratobj = seur_thym$mait, genesignature = "naive", file_name = "ch2_fig11_mait_naive.pdf")
lineage_genesignature(seuratobj = seur_thym$mait, genesignature = "effector", file_name = "ch2_fig11_mait_effector.pdf")
lineage_genesignature(seuratobj = seur_thym$mait, genesignature = "egress", file_name = "ch2_fig11_mait_egress.pdf")
```


## Nebulosa plot highlight CCR9 and CCR7
```{r mait-umap-densityplots, fig.height=10, fig.width=10}
lineage_nebulosagenes(
  seuratobj = seur_thym$mait,
  genes_vector = inkt_genes_nebulosa,
  pgrid_ncol = 2,
  pgrid_size = c(10, 30),
  file_name="no"
  # file_name = "ch2_fig11_mait_densityplots.pdf"
)
```


## MAIT DotPlot genes of interest
```{r mait-dotplot-genes-relevant, fig.height=10, fig.width=4}
mait_genes_dotplot <- c(
  "CD27", "KLRB1", "IL18R1", # human stage markers (Koay et al. 2016)
  "CD4", "CD8A", "CD8B", 
  "GNG4", "PDCD1", "TNFRSF9", # CD8aa
  "HIVEP3", "EGR2", "ZBTB16", # TFs induced in MAIT cells post-selection
  "CCR9", "LEF1", "ID3", # naive CCR9
  "CCR7", "SELL", "TCF7", # naive CCR7
  "S1PR1", "KLF2", # egress
  "FOS", "JUN", "JUNB", # AP1 signaling
  "TBX21", "EOMES", "IFNG", "GZMK", "GZMB", # type 1 effector
  "RORC", "RORA", "CCR6", "IL23R" # type 3 effector
)

plot_DotPlot(
  seurobj = seur_thym$mait,
  group="clusters_per_lineage",
  features = mait_genes_dotplot,
  scaling=T
)
# ggsave("~/Projects/phd/data/figures/chapter_02/figs_unfinished/section2/ch2_fig11_mait_dotplot.pdf", width=4, height=8)
```


## Co-expression type 1 and 17 gene signatures
```{r mait-type1-type17}
ggplot(
  df %>% filter(tcell_lineage=="MAIT"),
  aes(x=type1, y=type17, color=clusters_per_lineage)
)+
  geom_point()+
  scale_color_manual(values=c(cols_thym_nkt, cols_thym_mait, cols_thym_gdt))+
  labs(x="Type 1 signature", y="Type 17 signature")+
  theme_cowplot()
# ggsave("~/Projects/phd/data/figures/chapter_02/figs_unfinished/section2/ch2_fig11_mait_type1_17_coexpression.pdf", width=6, height=4)
```




# GD FIGURE

## DimPlot
```{r gdt-umap}
lineage_dimplot(seurobj = seur_thym$gdt, col_vector = cols_thym_gdt, file_name = "ch2_fig15_gdt_umap.pdf")

lineage_clustersperdonor(seurobj = seur_thym$gdt, col_vector = cols_thym_gdt, file_name = "ch2_fig15_gdt_clustersperdonor.pdf", figwidth=5)
```

## DotPlot with marker genes
```{r gdt-dotplot, fig.width=10, fig.height=7}
Idents(seur_thym$gdt) <- "clusters_per_lineage"
gdt_markergenes <- FindAllMarkers(
  seur_thym$gdt,
  test.use = 'wilcox',
  logfc.threshold = 0.4,
  min.pct = 0.3,
  only.pos = TRUE
  )
gdt_markergenes_list <- gdt_markergenes %>%
  mutate(cluster=as.character(cluster)) %>%
  arrange(cluster) %>%
  filter(p_val_adj<0.05) %>%
  filter(!gene %in% grep("RPL|RPS", gene, value=T)) %>%
  group_by(cluster) %>%
  top_n(-5, p_val_adj) %>%
  # top_n(5, avg_log2FC) %>%
  ungroup() %>%
  # filter(avg_log2FC>1.5) %>%
  pull(gene) %>% unique()

# dotplot
do_DotPlot(
  sample=seur_thym$gdt,
  features=unique(gdt_markergenes_list),
  group.by = "clusters_per_lineage",
  legend.position = "right",
  use_viridis = T,
  viridis.palette = "B",
  legend.title="z-score avg expression",
  # flip=T,
  scale=T,
  )+
  geom_vline(xintercept=c(5.5, 10.5, 15.5, 20.5, 25.5, 30.5, 35.5), linetype="dashed", color="grey")+
  theme(
    legend.title=element_text(size=8),
    legend.text = element_text(size=6),
    axis.text.x=element_text(size=8, angle=45, hjust=1, face="italic"),
    axis.text.y=element_text(size=8)
    )
# ggsave("~/Projects/phd/data/figures/chapter_02/figs_unfinished/section2/ch2_fig15_gdt_markergenes.pdf", width=11, height=5)
```


## Gene signatures
```{r gdt-gene-signatures, fig.height=5, fig.width=10}
lineage_genesignature(seuratobj = seur_thym$gdt, genesignature = "park_gdT", file_name = "ch2_fig15_gdt_parkGD_sig.pdf")
lineage_genesignature(seuratobj = seur_thym$gdt, genesignature = "park_CD8aa1", file_name = "ch2_fig15_gdt_cd8aa1_sig.pdf")
lineage_genesignature(seuratobj = seur_thym$gdt, genesignature = "park_CD8aa2", file_name = "ch2_fig15_gdt_cd8aa2_sig.pdf")
# lineage_genesignature(seuratobj = seur_thym$gdt, genesignature = "park_abT_entry", file_name = "ch2_fig15_gdt_SPentry_sig.pdf")
# lineage_genesignature(seuratobj = seur_thym$gdt, genesignature = "naive", file_name = "ch2_fig15_gdt_naive.pdf")
lineage_genesignature(seuratobj = seur_thym$gdt, genesignature = "effector", file_name = "ch2_fig15_gdt_effector.pdf")
lineage_genesignature(seuratobj = seur_thym$gdt, genesignature = "egress", file_name = "ch2_fig15_gdt_egress.pdf")
```


## Nebulosa plot highlight CCR9 and CCR7
```{r gdt-umap-densityplots, fig.height=10, fig.width=10}
gdt_genes_nebulosa <- c(
  "BCL11B", "GATA3",
  "EGR1", "ID3",
  "NT5E", "CD7",
  "ZBTB16",
  "TBX21", "RORC",
  "GZMK", "PRF1",
  "KLRB1"
)

lineage_nebulosagenes(seuratobj = seur_thym$gdt, genes_vector = gdt_genes_nebulosa, pgrid_ncol = 2, pgrid_size = c(10,30), file_name = "ch2_fig15_gdt_densityplots.pdf")
```


## GDT DotPlot genes of interest
```{r gdt-dotplot-genes-relevant}
gdt_genes_dotplot <- c(
  "TRDC", "TRGC2", "TRBC1", "TRAC", # TCR genes
  "CD4", "CD8A", "CD8B", # co-receptors
  "RAG1", "CD1C", "CD3E", # DP(q)
  "MKI67", "CD28", # ISP (p)
  "NOTCH3", "HES4", # Notch signaling
  "BCL11B", "GATA3", "TCF12", "IL7R", "RUNX1", "STAT5B", "TCF3", "MYB", # TCR rearrangement
  "EGR1", "ID3", "IKZF4", "NR4A1", "GNG4", "PDCD1", # TCR signaling and stop rearrangement
  "NT5E", "CD7", "CD27", # immature GD
  "IFI6", "MX1", "IRF1", # IFN signaling
  "S1PR1", "KLF2", # egress
  "TBX21", "EOMES", "IFNG", "GZMK", # type 1 effector
  "RORC", "RORA", "CCR6", "IL23R" # type 3 effector
)

plot_DotPlot(
  seurobj = seur_thym$gdt,
  group="clusters_per_lineage",
  features = gdt_genes_dotplot,
  scaling=T
)
# ggsave("~/Projects/phd/data/figures/chapter_02/figs_unfinished/section2/ch2_fig15_gdt_dotplot.pdf", width=4, height=8)
```


## Co-expression type 1 and 17 gene signatures
```{r gdt-type1-type17}
ggplot(
  df %>% filter(tcell_lineage=="GD"),
  aes(x=type1, y=type17, color=clusters_per_lineage)
)+
  geom_point()+
  scale_color_manual(values=c(cols_thym_nkt, cols_thym_mait, cols_thym_gdt))+
  labs(x="Type 1 signature", y="Type 17 signature")+
  theme_cowplot()
ggsave("~/Projects/phd/data/figures/chapter_02/figs_unfinished/section2/ch2_fig15_gdt_type1_17_coexpression.pdf", width=6, height=4)
```





# SESSION INFO
```{r}
sessionInfo()
```


