---
title: "Chapter 2 - Figure 7"
author: "Salomé Carcy"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true

---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 5, fig.height = 5,
                      warning=FALSE, message=FALSE,
                      root.dir = "~/Projects/phd/")
```


# IMPORT

## Import librairies
```{r import-librairies}
library(ggplot2)
library(cowplot)
library(RColorBrewer)
library(reshape2)
library(ggrepel)
library(ComplexHeatmap)

library(tidyverse)
library(dplyr)
library(Seurat)
# library(harmony)
# library(SCpubr)
library(MetaNeighbor)
library(SummarizedExperiment)
library(gplots)

source("~/Projects/phd/scripts/colors_universal.R")
```

## Import data
```{r import-data}
seur_integrated <- readRDS("./data/seurat_objects/seurat_human_integrated_object_23_12_01.rds")
DimPlot(seur_integrated, reduction="umap_integrated", cols = cols_integrated)
```




# FUNCTIONS
```{r define-functions}
plot_mtn_heatmap <- function(mtn_matrix, vector_groups_with_min_cells, df_annotation, df_nbcells){
  
  # keep only groups that have at least 100 cells
  cat("Metaneighbor matrix dimensions:", dim(mtn_matrix), "\n") # 88 rows & columns
  mtn_sub <- mtn_matrix[rownames(mtn_matrix) %in% vector_groups_with_min_cells, colnames(mtn_matrix) %in% vector_groups_with_min_cells]
  cat("Metaneighbor matrix without groups < 100 cells:", dim(mtn_sub), "\n") # 43 rows and 43 columns
  
  # reorder rows & columns by cluster
  groups_order_rows <- df_annotation[rownames(mtn_sub),] %>% arrange(clusters_integrated_data, tcell_lineage)
  groups_order_cols <- df_annotation[colnames(mtn_sub),] %>% arrange(clusters_integrated_data, tcell_lineage)
  mtn_sub <- mtn_sub[rownames(groups_order_rows), rownames(groups_order_cols)]
  cat("\nOrder matrix rows:", rownames(mtn_sub), "\n\n")
  cat("Order matrix columns:", colnames(mtn_sub), "\n\n")
  
  # row annotation
  row_annot <-
    rowAnnotation(
      nbcells=anno_barplot(
        df_nbcells %>% filter(group %in% rownames(mtn_sub)) %>% arrange(match(group, rownames(mtn_sub))) %>% pull(n),
        axis_param = list(direction = "reverse")
        ),
      lineage=df_annotation[rownames(mtn_sub),"tcell_lineage"],
      cluster=df_annotation[rownames(mtn_sub),"clusters_integrated_data"],
      col=list(
        cluster=deframe(df_annotation[rownames(mtn_sub),c("clusters_integrated_data", "col_clusters")]),
        lineage=deframe(df_annotation[rownames(mtn_sub),c("tcell_lineage", "col_lineages")])
        )
      )
  
  # column annotation
  col_annot <-
    HeatmapAnnotation(
      nbcells=anno_barplot(
        df_nbcells %>% filter(group %in% colnames(mtn_sub)) %>% arrange(match(group, colnames(mtn_sub))) %>% pull(n)
        ),
      lineage=df_annotation[colnames(mtn_sub),"tcell_lineage"],
      cluster=df_annotation[colnames(mtn_sub),"clusters_integrated_data"],
      col=list(
        cluster=deframe(df_annotation[colnames(mtn_sub),c("clusters_integrated_data", "col_clusters")]),
        lineage=deframe(df_annotation[colnames(mtn_sub),c("tcell_lineage", "col_lineages")])
        )
      )
  
  # heatmap
  Heatmap(
    mtn_sub,
    name="auroc",
    # cluster
    cluster_rows=F,
    cluster_columns = F,
    top_annotation=col_annot,
    left_annotation=row_annot,
    # add values
    layer_fun = function(j, i, x, y, width, height, fill) {
          v = pindex(mtn_sub, i, j)
          l = v > 0.9 | v < 0.1
          grid.text(sprintf("%.2f", v[l]), x[l], y[l], gp = gpar(fontsize = 10))
          }
    )
}
```




# METANEIGHBOR BY CLUSTER

## Prepare data
```{r mtn_prepare_data}
# subset to thymocytes only
seur_thym <- subset(seur_integrated, subset=tissue=="Thymus")
print(seur_thym)

# get count matrix
se_thym <- SummarizedExperiment(
  assays=seur_thym@assays[["RNA"]]@counts,
  colData=seur_thym@meta.data[,c("tcell_lineage", "donor_id", "clusters_integrated_data")]
  )
```


## Run metaneighbor
```{r run_mtn}
mtn <- MetaNeighborUS(
  var_genes=VariableFeatures(seur_thym),
  dat=se_thym,
  study_id=se_thym$tcell_lineage,
  cell_type=se_thym$clusters_integrated_data,
  fast_version=T
  )
```

```{r plot_mtn_fullheatmap, fig.width=14, fig.height=14}
heatmap.2(mtn,
          # trace
          trace="none",
          # dendrogram
          # Rowv=FALSE,
          # Colv=FALSE,
          # dendrogram="none",
          # superimpose a density histogram on color key
          density.info="none",
          # color scale
          col=rev(colorRampPalette(brewer.pal(11,"RdYlBu"))(100)),
          breaks=seq(0,1,length=101),
          key.xlab = "AUROC",
          key.title="",
          keysize = 1.2,
          # text labels
          main="",
          cexRow=0.6,
          cexCol=0.6,
          # margins
          margins=c(9,9))
```


## Plot metaneighbor full heatmap
First, let's obtain the number of cells per group (we are not interested in groups that contain less than 100 cells, it's less reliable).
```{r prepare_data_heatmaps}
# Get number of cells per lineage and cluster
nbcells <- seur_thym@meta.data %>%
  as_tibble() %>%
  group_by(tcell_lineage, clusters_integrated_data) %>%
  dplyr::count() %>%
  ungroup() %>%
  mutate(group=paste(tcell_lineage, clusters_integrated_data, sep="|")) %>%
  select(group, n)
groups_with_min_cells <- nbcells %>% filter(n>=100) %>% pull(group)

# general dataframe for annotation
df_annot <- data.frame("names"=nbcells$group) %>%
  separate(names, c("tcell_lineage", "clusters_integrated_data"), remove=F) %>%
  left_join(enframe(cols_integrated) %>% dplyr::rename(col_clusters=value), by=join_by(clusters_integrated_data==name)) %>%
  left_join(enframe(cols_lineages) %>% dplyr::rename(col_lineages=value), by=join_by(tcell_lineage==name)) %>%
  mutate(clusters_integrated_data=factor(clusters_integrated_data, levels=0:17)) %>%
  column_to_rownames("names")
```

Then, we can plot the heatmap (full heatmap).
```{r heatmap_full, fig.width=12, fig.height=12}
pdf("~/Projects/phd/data/figures/chapter_02/figs_unfinished/section1/ch2_fig7_mtn_clusters_integrated.pdf", width=17, height=15)
plot_mtn_heatmap(mtn_matrix=mtn, vector_groups_with_min_cells = groups_with_min_cells, df_annotation = df_annot, df_nbcells=nbcells)
dev.off()
```


## Plot metaneighbor Tconv vs Tinn
Plot metaneighbor result in a more readable format to compare Tconv vs Tinn.
```{r heatmap_tconv_tinn, fig.width=8, fig.height=8}
mtn_conv_inn <- mtn[str_detect(rownames(mtn),"CD4|CD8"),str_detect(colnames(mtn),"iNKT|MAIT|GD")]

plot_mtn_heatmap(mtn_matrix=mtn_conv_inn, vector_groups_with_min_cells = groups_with_min_cells, df_annotation = df_annot)
```




# METANEIGHBOR BY "BRANCH"

## Prepare data
```{r mtn_branches_prepare_data}
print(seur_thym)

# create a new metadata column for "branches"
seur_thym@meta.data
seur_thym@meta.data$clusters_branch <- dplyr::case_when(
  seur_thym@meta.data$clusters_integrated_data %in% c(0,1,2) ~ "pre-selection",
  seur_thym@meta.data$clusters_integrated_data %in% c(3,11)  ~ "CD4 branch",
  seur_thym@meta.data$clusters_integrated_data %in% c(9,10)  ~ "CD8 branch",
  seur_thym@meta.data$clusters_integrated_data %in% c(4,5,7) ~ "agonist branch",
  seur_thym@meta.data$clusters_integrated_data == 6          ~ "GD branch",
  seur_thym@meta.data$clusters_integrated_data == 8          ~ "IFN signaling",
  seur_thym@meta.data$clusters_integrated_data %in% c(12:17) ~ "effector",
  .default = NA
)
table(seur_thym@meta.data[,c("clusters_integrated_data", "clusters_branch")], useNA="ifany") # sanity check


# get count matrix
se_thym2 <- SummarizedExperiment(
  assays=seur_thym@assays[["RNA"]]@counts,
  colData=seur_thym@meta.data[,c("tcell_lineage", "donor_id", "clusters_branch")]
  )
```


## Run metaneighbor
```{r run_mtn_branches}
mtn2 <- MetaNeighborUS(
  var_genes=VariableFeatures(seur_thym),
  dat=se_thym2,
  study_id=se_thym2$donor_id,
  cell_type=se_thym2$tcell_lineage,
  fast_version=T
  )
```

```{r plot_mtn_branches_fullheatmap, fig.width=14, fig.height=14}
heatmap.2(mtn2,
          # trace
          trace="none",
          # dendrogram
          # Rowv=FALSE,
          # Colv=FALSE,
          # dendrogram="none",
          # superimpose a density histogram on color key
          density.info="none",
          # color scale
          col=rev(colorRampPalette(brewer.pal(11,"RdYlBu"))(100)),
          breaks=seq(0,1,length=101),
          key.xlab = "AUROC",
          key.title="",
          keysize = 1.2,
          # text labels
          main="",
          cexRow=0.6,
          cexCol=0.6,
          # margins
          margins=c(9,9))
```

## Plot metaneighbor full heatmap
First, let's obtain the number of cells per group (we are not interested in groups that contain less than 100 cells, it's less reliable).
```{r prepare_data_heatmaps_branches}
# Get number of cells per lineage and cluster
nbcells2 <- seur_thym@meta.data %>%
  as_tibble() %>%
  group_by(tcell_lineage, clusters_branch) %>%
  dplyr::count() %>%
  ungroup() %>%
  mutate(group=paste(tcell_lineage, clusters_branch, sep="|")) %>%
  select(group, n)
groups_with_min_cells2 <- nbcells2 %>% filter(n>=100) %>% pull(group)

# general dataframe for annotation
df_annot2 <- data.frame("names"=nbcells2$group) %>%
  separate(names, c("tcell_lineage", "clusters_branch"), sep="\\|", remove=F) %>%
  mutate(col_clusters=case_when(
    clusters_branch=="pre-selection" ~"#f4c40f",
    clusters_branch=="CD4 branch"    ~"#e09351",
    clusters_branch=="CD8 branch"    ~"#5a97c1",
    clusters_branch=="agonist branch"~"#2b9b81",
    clusters_branch=="GD branch"     ~"#92c051",
    clusters_branch=="IFN signaling" ~"#17154f",
    clusters_branch=="effector"      ~"#a40000"
  )) %>%
  left_join(enframe(cols_lineages) %>% dplyr::rename(col_lineages=value), by=join_by(tcell_lineage==name)) %>%
  mutate(clusters_branch=factor(clusters_branch, levels=c("pre-selection", "GD branch", "agonist branch", "CD4 branch", "CD8 branch", "IFN signaling", "effector"))) %>%
  column_to_rownames("names") %>%
  # for the function
  dplyr::rename(clusters_integrated_data=clusters_branch)

```

Then, we can plot the heatmap (full heatmap).
```{r heatmap_full, fig.width=12, fig.height=12}
plot_mtn_heatmap(mtn_matrix=mtn2, vector_groups_with_min_cells = groups_with_min_cells2, df_annotation = df_annot2, df_nbcells=nbcells2)
```




# SESSION INFO
```{r}
sessionInfo()
```


