---
title: "Chapter X - R Notebook"
author: "Salomé Carcy"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true

---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 5, fig.height = 5,
                      warning=FALSE, message=FALSE,
                      root.dir = "~/Projects/phd/")
```


# IMPORT

## Import librairies
```{r import-librairies}
library(ggplot2)
library(RColorBrewer)
library(cowplot)
library(tidyverse)
library(dplyr)
library(Seurat)
library(SCpubr)
library(patchwork)

source("~/Projects/phd/scripts/colors_universal.R")
```

## Import data
```{r import-data}
seur_integrated <- readRDS("./data/seurat_objects/seurat_human_integrated_object_23_12_01.rds")
seur_pbmc <- subset(seur_integrated, subset=tissue=="PBMC")
```




# FUNCTIONS
```{r define-functions}
featureplot_save <- function(seurobj=seur_integrated, genes=c("CD4", "CD8A", "CD8B")){
  for(gene in genes){
    print(gene)
    p <- do_FeaturePlot(
      sample = seurobj,
      slot="data",
      cells.highlight=rownames(seurobj@meta.data[seurobj@meta.data$tissue=="PBMC",]),
      features = gene,
      use_viridis=T,
      viridis.palette = "B",
      legend.position="right",
      order=T,
      pt.size=6,
      raster=T,
      raster.dpi = 2048
    )
    ggsave(filename=paste0("~/Projects/phd/data/figures/chapter_03/figs_unfinished/section1/ch3_fig1_", gene, ".pdf"),
           plot=p,
           width=8, height=7)
  }
}


plot_DotPlot <- function(seurobj, group, features, scaling=T){
  # get plot data
  p <- Seurat::DotPlot(
    seurobj,
    group.by=group,
    features=features,
    scale=scaling
  )
  # plot in personalized way
  p <- ggplot(p$data, aes(x=id, y=features.plot, fill=avg.exp.scaled, size=pct.exp))+
    geom_point(color="black", shape=21)+
    # scale_fill_gradient2(low=scales::muted("blue"), high=scales::muted("red"), name="z-score\nnormalized\navg expression")+
    scale_size_continuous(range=c(0,6), limits=c(0,100), name="%cells\nexpressing\ngene")+
    theme_bw()+
    theme(axis.text.y=element_text(face="italic"))+
    labs(y="", x="")
  # different color scale if scaled or not
  if(scaling==T){
    p <- p + scale_fill_gradient2(low=scales::muted("blue"), high=scales::muted("red"), name="z-score\nnormalized\navg expression")
  } else{
    p <- p + viridis::scale_fill_viridis(option="B", direction=-1, name="normalized\navg expression")
  }
  return(p)
}


lineage_genesignature <- function(seuratobj, genesignature, file_name="no", ordercells=T){
  minimum_genesignature <- min(seuratobj@meta.data[,genesignature])
  minimum_to_plot <- ifelse(minimum_genesignature<0, 0, minimum_genesignature)
  p <- ggrastr::rasterise(
    do_FeaturePlot(
    seuratobj,
    features = genesignature,
    min.cutoff=minimum_to_plot,
    use_viridis = T,
    viridis.palette = "E",
    legend.position="right",
    order=ordercells,
    pt.size = 1.5,
    raster=F
    # raster.dpi=2048,
    # pt.size=10
  ),
  layers="Point",
  dpi=300
  )
  print(p)
  if(file_name != "no"){
    ggsave(paste0("~/Projects/phd/data/figures/chapter_03/figs_unfinished/section1/", file_name),
           plot=p,
           width=9, height=6)
  }
}

plot_beeswarm <- function(seurobj=seur_pbmc, feature, ytitle, file_name="no"){
  p <- ggplot(
    as.data.frame(seurobj@meta.data),
    aes_string(x="clusters_integrated_data", y=feature, fill="clusters_integrated_data")
  ) +
    # geom_violin(alpha=0.5, width=1.5,color=NA) +
    ggbeeswarm::geom_quasirandom(shape = 21, size=2, width = .4, color="black", show.legend = F, stroke=0.1)+
    scale_fill_manual(values=cols_integrated)+
    labs(x="", y=ytitle)+
    ylim(c(0, max(seurobj@meta.data[,feature])))+
    theme_classic()+
    theme(
      legend.position="none",
      axis.line = element_blank(),
      panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.5)
    )
  p <- ggrastr::rasterise(p, layers="Point", dpi=300)
  print(p)
  if(file_name != "no"){
    ggsave(paste0("~/Projects/phd/data/figures/chapter_03/figs_unfinished/section1/ch3_fig1_literature_signatures/", file_name),
           plot=p,
           width=8, height=5)
  }
  # return(p)
}

```




# ANALYSIS

## Integrated clustering separated by tissue
```{r fig1_clusters_by_tissue, fig.width=20, fig.height=5}
# Highlight thymocytes
do_DimPlot(
    sample = seur_integrated,
    group.by="clusters_integrated_data",
    split.by="tissue",
    colors.use = cols_integrated,
    legend.position="none",
    pt.size=5,
    raster=T,
    raster.dpi = 2048
)
# ggsave("~/Projects/phd/data/figures/chapter_03/figs_unfinished/section1/ch3_fig1_clusters_by_tissue.pdf", width=18, height=6)
```


## FeaturePlot with CD4 and CD8
```{r umap-CD4-CD8}
genes_to_plot <- c("CD4", "CD8A", "CD8B", "ZBTB7B", "CD40LG", "RUNX3", "LINC02446")

# Expression of key Tconv genes
for(gene in genes_to_plot){
  # print(gene)
  print(do_FeaturePlot(
      sample = seur_integrated,
      slot="data",
      cells.highlight=rownames(seur_integrated@meta.data[seur_integrated@meta.data$tissue=="PBMC",]),
      features = gene,
      use_viridis=T,
      viridis.palette = "B",
      legend.position="right",
      order=T,
      pt.size=1
    ))
}

# save CD4, CD8A, CD8B for figure 1
# featureplot_save()
```


## DotPlot
The marker genes were chosen based on:
- [Sallusto et al. 1999](https://www.nature.com/articles/44385)
- [Terekhova et al. 2023](https://www.sciencedirect.com/science/article/pii/S1074761323004533)
- [Rose et al. 2023](https://www.nature.com/articles/s42003-023-04747-9)
```{r dotplot-marker-genes}
genes_dotplot <- c(
  # ------
  # Prolif
  "MKI67",
  "CDK1",
  # ------
  # Naive GD
  # "TRDC", "TRGC1", "TRGC2",
  # "CD24",
  # "CD38",
  # ------
  # IFN signaling
  "ISG15",
  "STAT1",
  "IFI6",
  # ------
  # Tnaive & Tcm
  "CCR7",
  "SELL",
  "TCF7",
  "LEF1",
  "CD27",
  # ------
  # Tcm
  "AHR",
  # "GPR15",
  "CCR4",
  # "IL2",
  # "IL4",
  "IL4R",
  "CCR6",
  "RORC",
  # "ICOS",
  # "IL2RA",
  # ------
  # Tcm & Tem
  "CD69",
  "IL7R",
  # ------
  # Tem & Temra
  "GZMK",
  "CCL5",
  "EOMES",
  "KLRG1",
  # ------
  # Temra
  "PRF1",
  "GNLY",
  "CX3CR1",
  "GZMB",
  "CCL4",
  "ZNF683",
  # "CXCR3", # very little expression... (same in Terekhova)
  "IKZF2",
  "TYROBP",
  "NCAM1", # CD56
  "B3GAT1" # CD57
  # "IDH2",
  # "COX8A",
  # "CYC1"
  )

# get list of cell id to plot (remove clusters with less than 100 cells)
plot_DotPlot(
  seurobj=seur_integrated[,rownames(seur_integrated@meta.data[seur_integrated@meta.data$tissue=="PBMC",])],
  group="clusters_integrated_data",
  features=genes_dotplot,
  scaling=T
  )
# ggsave("~/Projects/phd/data/figures/chapter_03/figs_unfinished/section1/ch3_fig1_markergenes.pdf", width=6, height=6)
```


## Cell hashing distribution by cluster
```{r fig1_cellhashing_nbcells, fig.height=6}
# Barplot of nb of cells
p1 <- seur_integrated@meta.data %>%
  as_data_frame() %>%
  filter(tissue=="PBMC") %>%
  group_by(clusters_integrated_data, tcell_lineage) %>%
  dplyr::count() %>%
  ggplot(aes(x=clusters_integrated_data, y=n, fill=tcell_lineage))+
  geom_bar(stat="identity")+
  scale_fill_manual(values=cols_lineages, name="Cell hashing")+
  scale_y_continuous(expand = c(0, 0))+
  labs(x="", y="# T cells")+
  theme_cowplot()

p2 <- plot_DotPlot(
  seurobj=seur_integrated[,rownames(seur_integrated@meta.data[seur_integrated@meta.data$tissue=="PBMC",])],
  group="clusters_integrated_data",
  features=c(
    # CD4
    "CD4",
    "ZBTB7B",
    "CD40LG",
    # Treg
    "IKZF4",
    "FOXP3",
    "CTLA4",
    "IL2RA",
    # CD8
    "CD8A",
    "CD8B",
    "RUNX3",
    "LINC02446",
    # Tinn
    "ZBTB16",
    "TRGC1"
  ),
  scaling=T
)

# combine
p1 / p2
# ggsave("~/Projects/phd/data/figures/chapter_03/figs_unfinished/section1/ch3_fig1_cellhashing_per_cluster.pdf", width=8, height=6)
```



## Score litterature gene signatures
First prepare the gene signatures
```{r gene-signatures-prep}
gene_signatures_df <- list()

# ROSE ET AL.
genesCD8.rose <- readxl::read_excel("./data/litterature_gene_signatures/rose_supp_clusmodulesCD8.xlsx", sheet=1)
genesCD4.rose <- readxl::read_excel("./data/litterature_gene_signatures/rose_supp_clusmodulesCD4.xlsx", sheet=1)
# Clean up and extract signatures
# head(genesCD4.rose)
# head(genesCD8.rose)
roseCD4 <- genesCD4.rose %>%
  select(SYMBOL, k.5.cluster) %>%
  filter(!is.na(SYMBOL)) %>% # the few NA symbols are miRNA or uncharacterized loci
  dplyr::rename(gene=SYMBOL, geneprogram=k.5.cluster) %>%
  mutate(geneprogram=ifelse(geneprogram==1, "Rose | CD4_modul1_Tem/cm",
                     ifelse(geneprogram==2, "Rose | CD4_modul2_Tcm/em",
                     ifelse(geneprogram==3, "Rose | CD4_modul3_Tem/cm",
                     ifelse(geneprogram==4, "Rose | CD4_modul4_Tem",
                     ifelse(geneprogram==5, "Rose | CD4_modul5_Tnaive", "?"))))))
roseCD8 <- genesCD8.rose %>%
  select(SYMBOL, k.5.cluster) %>%
  filter(!is.na(SYMBOL)) %>% # the few NA symbols are miRNA or uncharacterized loci
  dplyr::rename(gene=SYMBOL, geneprogram=k.5.cluster) %>%
  mutate(geneprogram=ifelse(geneprogram==1, "Rose | CD8_modul1_Temra",
                     ifelse(geneprogram==2, "Rose | CD8_modul2_Tcm/em",
                     ifelse(geneprogram==3, "Rose | CD8_modul3_Tem/emra",
                     ifelse(geneprogram==4, "Rose | CD8_modul4_Tnaive",
                     ifelse(geneprogram==5, "Rose | CD8_modul5_Tnaive/cm", "?"))))))
rosetotal <- rbind(roseCD4, roseCD8) %>%
  mutate(geneprogram_cat=case_when(
    geneprogram=="Rose | CD4_modul1_Tem/cm"    ~"rose_Tem",
    geneprogram=="Rose | CD4_modul2_Tcm/em"    ~"rose_Tcm",
    geneprogram=="Rose | CD4_modul3_Tem/cm"    ~"rose_Tem",
    geneprogram=="Rose | CD4_modul4_Tem"       ~"rose_Tem",
    geneprogram=="Rose | CD4_modul5_Tnaive"    ~"rose_Tnaive",
    geneprogram=="Rose | CD8_modul1_Temra"     ~"rose_Temra",
    geneprogram=="Rose | CD8_modul2_Tcm/em"    ~"rose_Tcm",
    geneprogram=="Rose | CD8_modul3_Tem/emra"  ~"rose_Tem",
    geneprogram=="Rose | CD8_modul4_Tnaive"    ~"rose_Tnaive",
    geneprogram=="Rose | CD8_modul5_Tnaive/cm" ~"rose_Tnaive"
  ))
head(rosetotal)

gene_signatures_df[["rose"]] <- rosetotal


# TEREKHOVA ET AL.
# genes.terekhova_general <- readxl::read_excel("./data/litterature_gene_signatures/terekhova_supptable.xlsx", sheet=1)
genes.terekhova_cd4 <- readxl::read_excel("./data/litterature_gene_signatures/terekhova_supptable.xlsx", sheet=3)
genes.terekhova_cd8 <- readxl::read_excel("./data/litterature_gene_signatures/terekhova_supptable.xlsx", sheet=5)
# genes.terekhova_gdt <- readxl::read_excel("./data/litterature_gene_signatures/terekhova_supptable.xlsx", sheet=7)
# Clean up and extract signatures
# head(genes.terekhova_cd4)
terekhovadf <- cbind(genes.terekhova_cd4, genes.terekhova_cd8) %>%
  pivot_longer(cols=everything(), names_to="geneprogram", values_to = "gene") %>%
  mutate(geneprogram=paste0("Terekhova | ", geneprogram),
         geneprogram_cat=case_when(
           geneprogram=="Terekhova | CD4_Tnaive"               ~ "terekhova_Tnaive",
           geneprogram=="Terekhova | CD4_Tnaive_IFN"           ~ "terekhova_Tnaive_IFN",
           geneprogram=="Terekhova | CD4_Tfh"                  ~ "terekhova_Tcm",
           # geneprogram=="Terekhova | CD4_Th1"                  ~ "terekhova_Tcm",
           geneprogram=="Terekhova | CD4_Th1/Th17"             ~ "terekhova_Tem",
           geneprogram=="Terekhova | CD4_Th17"                 ~ "terekhova_Tem",
           geneprogram=="Terekhova | CD4_Th22"                 ~ "terekhova_Tem",
           geneprogram=="Terekhova | CD4_Th2"                  ~ "terekhova_Tcm",
           geneprogram=="Terekhova | CD4_HLADRpos_memory"      ~ "terekhova_Tem",
           # geneprogram=="Terekhova | CD4_Exhausted-like_memory"~ "terekhova_Tcm",
           geneprogram=="Terekhova | CD4_Terminal_effector"    ~ "terekhova_Temra",
           geneprogram=="Terekhova | CD4_Temra"                ~ "terekhova_Temra",
           # geneprogram=="Terekhova | CD4_Treg_cytotoxic"       ~ "terekhova_Treg",
           geneprogram=="Terekhova | CD4_Treg_naive"           ~ "terekhova_Treg",
           # geneprogram=="Terekhova | CD4_Treg_memory"          ~ "terekhova_Treg",
           # geneprogram=="Terekhova | CD4_Treg_KLRB1+_RORC+"    ~ "terekhova_Treg",
           geneprogram=="Terekhova | CD8_Tnaive"        ~ "terekhova_Tnaive",
           geneprogram=="Terekhova | CD8_Tnaive_IFN"    ~ "terekhova_Tnaive_IFN",
           geneprogram=="Terekhova | CD8_Tcm_CCR4pos"   ~ "terekhova_Tcm",
           geneprogram=="Terekhova | CD8_Tcm_CCR4neg"   ~ "terekhova_Tcm",
           # geneprogram=="Terekhova | CD8_Trm"           ~ "terekhova_Trm",
           # geneprogram=="Terekhova | CD8_Tmem_KLRC2pos" ~ "terekhova_other",
           geneprogram=="Terekhova | CD8_Tem_GZMKpos"   ~ "terekhova_Tem",
           geneprogram=="Terekhova | CD8_HLADRpos"      ~ "terekhova_Tem",
           geneprogram=="Terekhova | CD8_proliferative" ~ "terekhova_Tprolif",
           geneprogram=="Terekhova | CD8_Tem_GZMBpos"   ~ "terekhova_Temra",
           geneprogram=="Terekhova | CD8_Temra"         ~ "terekhova_Temra",
           geneprogram=="Terekhova | CD8_NKTlike"       ~ "terekhova_Temra"
         ))
gene_signatures_df[["terekhova"]] <- terekhovadf


# Get it all in lists
longdf <- bind_rows(gene_signatures_df, .id="dataset") %>%
  filter(gene %in% rownames(seur_integrated)) %>%
  filter(!is.na(geneprogram_cat))
# table(longdf$geneprogram_cat, useNA="ifany")

gene_signatures_list <- list()
for(gp in unique(longdf$geneprogram_cat)){
  print(gp)
  genes_in_program <- longdf %>%
    filter(geneprogram_cat==gp & gene %in% rownames(seur_integrated)) %>%
    pull(gene)
  print(length(unique(genes_in_program)))
  gene_signatures_list[[gp]] <- unique(genes_in_program)
}
print(lengths(gene_signatures_list))
```

Score those gene signature on seurat pbmc object.
```{r gene-signatures-score}
# seur_pbmc@meta.data
# seur_pbmc@meta.data[,32:43] <- NULL
seur_pbmc <- AddModuleScore(seur_pbmc, name = names(gene_signatures_list), features=gene_signatures_list, seed=1)

# Remove the annoying numbers that are being added
colnames(seur_pbmc@meta.data)[32:42] <- names(gene_signatures_list)
```

Plot these gene signatures on UMAP
```{r gene-signatures-plot, fig.height=6, fig.width=6}
for(gp in names(gene_signatures_list)){
  print(gp)
  lineage_genesignature(
    seur_pbmc,
    gp,
    # file_name = paste0("ch3_fig1_literature_signatures/", gp, ".pdf"),
    ordercells = T
    )
}
```

Plot these gene signatures on violin plots
```{r gene-signatures-plot}
for(gp in names(gene_signatures_list)){
  print(gp)
  plot_beeswarm(
    feature = gp,
    # file_name = paste0("ggswarm_", gp, ".pdf")
    ytitle = gp
  )
}
```





# SESSION INFO
```{r}
sessionInfo()
```


