---
title: "Chapter 3 - CD8 blood figure"
author: "Salomé Carcy"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true

---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 5, fig.height = 5,
                      warning=FALSE, message=FALSE,
                      root.dir = "~/Projects/phd/")
```


# IMPORT

## Import librairies
```{r import-librairies}
library(ggplot2)
library(RColorBrewer)
library(cowplot)
library(tidyverse)
library(dplyr)
library(Seurat)
library(SCpubr)
library(patchwork)
library(SummarizedExperiment)
library(MetaNeighbor)

source("~/Projects/phd/scripts/colors_universal.R")
```

## Import data
```{r import-data}
seur_integrated <- readRDS("./data/seurat_objects/seurat_human_integrated_object_23_12_01.rds")
DimPlot(seur_integrated, reduction="umap_integrated", cols = cols_integrated)

seur_pbmc_cd8 <- readRDS("./data/seurat_objects/pbmc_objects/seurat_pbmc_cd8_clean.rds")
```




# FUNCTIONS
```{r define-functions}
lineage_dimplot <- function(seurobj, col_vector, file_name="no"){
  p <- do_DimPlot(
    seurobj,
    group.by="clusters_per_lineage",
    colors.use = col_vector,
    legend.position = "right",
    # raster = T,
    # raster.dpi=2048,
    pt.size = 0.5#,
    # border.size=1.5
  )+
    theme(
      # axis.title=element_text(size=8),
      # axis.text=element_text(size=6),
      legend.key.size = unit(0.4, "cm"),
      legend.title=element_text(size=8),
      legend.text = element_text(size=6)
    )
  if(file_name != "no"){
    ggsave(paste0("~/Projects/phd/data/figures/chapter_03/figs_unfinished/section1/", file_name),
           plot=p,
           width=4, height=2.5, units="in")
  }
}


lineage_clustersperdonor <- function(seurobj, col_vector, file_name="no", figwidth=6){
  p <- seurobj@meta.data %>%
    group_by(Donor, clusters_per_lineage) %>%
    dplyr::count() %>%
    group_by(Donor) %>%
    mutate(sum_per_donor=sum(n),
           freq=n*100/sum_per_donor) %>%
    mutate(Donor=factor(Donor, levels=c(8, 9, 10, 12, 13, 5, 11, 6, 7))) %>%
  ggplot(aes(x=factor(Donor, levels=sort(unique(Donor))), y=freq, fill=clusters_per_lineage))+
    geom_bar(stat="identity")+
    scale_fill_manual(values=col_vector)+
    labs(x="Donor", y="% cells", fill="")+
    theme_cowplot()#+
  # theme(axis.text=element_text(size=5/.pt),
  #       axis.title=element_text(size=7/.pt),
  #       legend.text = element_text(size=5/.pt))
  if(file_name != "no"){
    ggsave(paste0("~/Projects/phd/data/figures/chapter_03/figs_unfinished/section1/", file_name),
           plot=p,
           width=figwidth, height=5, units="in")
  }
}


plot_DotPlot <- function(seurobj, group, features, scaling=T){
  # get plot data
  p <- Seurat::DotPlot(
    seurobj,
    group.by=group,
    features=features,
    scale=scaling
  )
  # plot in personalized way
  p <- ggplot(p$data, aes(x=id, y=features.plot, fill=avg.exp.scaled, size=pct.exp))+
    geom_point(color="black", shape=21)+
    # scale_fill_gradient2(low=scales::muted("blue"), high=scales::muted("red"), name="z-score\nnormalized\navg expression")+
    scale_size_continuous(range=c(0,6), limits=c(0,100), name="%cells\nexpressing\ngene")+
    theme_bw()+
    theme(
      axis.text.y=element_text(face="italic"),
      axis.text.x=element_text(angle=45, hjust=1)
    )+
    labs(y="", x="")
  # different color scale if scaled or not
  if(scaling==T){
    p <- p + scale_fill_gradient2(low=scales::muted("blue"), high=scales::muted("red"), name="z-score\nnormalized\navg expression")
  } else{
    p <- p + viridis::scale_fill_viridis(option="B", direction=-1, name="normalized\navg expression")
  }
  return(p)
}

lineage_nebulosagenes <- function(seuratobj, genes_vector, pgrid_ncol=2, pgrid_size=c(10,30), file_name="no"){
  
  pnebulosa <- list()
  for(gene in genes_vector) {
    p <- SCpubr::do_NebulosaPlot(seuratobj, features = gene, sequential.palette="YlOrRd")
    p <- ggrastr::rasterise(p, layers="Point", dpi=200)
    pnebulosa[[gene]] <- p
  }
  pgrid <- plot_grid(plotlist=pnebulosa, ncol=pgrid_ncol)
  print(pgrid)
  
  if(file_name != "no"){
    ggsave(paste0("~/Projects/phd/data/figures/chapter_03/figs_unfinished/section1/", file_name),
           plot=pgrid,
           width=pgrid_size[1], height=pgrid_size[2])
  }
}
```




# UMAP and distribution per donor

## Umap
```{r umap-cd8}
lineage_dimplot(seurobj = seur_pbmc_cd8, col_vector=cols_pbmc_cd8, file_name="ch3_fig3_cd8_umap.pdf")
```


## Distribution per donor
```{r barplot_per_donor}
# cluster comp per donor
lineage_clustersperdonor(seurobj = seur_pbmc_cd8, col_vector = cols_pbmc_cd8, file_name = "ch3_fig3_cd8_clustersperdonor.pdf")
```








# GENES OF INTEREST
## Dotplot genes on interest
The marker genes were chosen based on:
- [Sallusto et al. 1999](https://www.nature.com/articles/44385)
- [Terekhova et al. 2023](https://www.sciencedirect.com/science/article/pii/S1074761323004533)
- [Rose et al. 2023](https://www.nature.com/articles/s42003-023-04747-9)
```{r dotplot-marker-genes, fig.height=5}
genes_dotplot <- c(
  # CD4
  # "CD8A",
  # "CD8B",
  # "RUNX3",
  "LINC02446",
  # ------
  # IFN signaling
  # "ISG15",
  # "STAT1",
  # "IFI6",
  # ------
  # Tnaive & Tcm
  "CCR7",
  "SELL",
  "TCF7",
  "LEF1",
  "CD27",
  # ------
  # Tcm
  # "AHR",
  # "GPR15",
  "CCR4",
  # "IL2",
  # "IL4",
  "IL4R",
  "CCR6",
  # "RORC",
  # "ICOS",
  # "IL2RA",
  # ------
  # Tcm & Tem
  "AHR",
  "CD69",
  "IL7R",
  # ------
  # Tem & Temra
  "GZMK",
  "CCL5",
  "EOMES",
  "KLRG1",
  # ------
  # Temra
  "PRF1",
  "GNLY",
  "CX3CR1",
  "GZMB",
  "CCL4",
  "ZNF683",
  # # "CXCR3", # very little expression... (same in Terekhova)
  # "IKZF2",
  "TYROBP",
  "NCAM1", # CD56
  "B3GAT1", # CD57
  # "IDH2",
  # "COX8A",
  # "CYC1"
  # ------
  # Proliferating
  "CDK1",
  "MKI67"
  )

# get list of cell id to plot (remove clusters with less than 100 cells)
plot_DotPlot(
  seurobj=seur_pbmc_cd8,
  group="clusters_per_lineage",
  features=genes_dotplot,
  scaling=F
  )
# ggsave("~/Projects/phd/data/figures/chapter_03/figs_unfinished/section1/ch3_fig3_cd8_markergenes.pdf", width=5, height=6)
```


## Nebulosa plot genes of interest
```{r, fig.height=8, fig.width=6}
lineage_nebulosagenes(
  seur_pbmc_cd8,
  genes_vector=c("CCR7", "SELL", "IL4R", "CD69", "GZMK", "GZMB"),
  # file_name="ch3_fig3_cd8_nebulosaplots.pdf",
  pgrid_ncol=2,
  pgrid_size=c(10, 15)
)
```




# METANEIGHBOR

## Metaneighbor thymocytes x pbmcs

Prepare data.
```{r prepare-cd8-mtn}
# Get only CD8s
seur_cd8 <- subset(seur_integrated, subset=tcell_lineage=="CD8")
table(seur_cd8@meta.data[,c("clusters_per_lineage", "donor_id")], useNA="ifany") # sanity check

# rename clusters_per_lineage for easier interpretation of metaneighbor
seur_cd8@meta.data$clusters_per_lineage_annotated <- case_when(
  seur_cd8@meta.data$clusters_per_lineage=="CD8_thymus_c0" ~ "CD8 DPp (c0)",
  seur_cd8@meta.data$clusters_per_lineage=="CD8_thymus_c1" ~ "GNG4+ CD8aa1 (c1)",
  seur_cd8@meta.data$clusters_per_lineage=="CD8_thymus_c2" ~ "ZNF683+ CD8aa2 (c2)",
  seur_cd8@meta.data$clusters_per_lineage=="CD8_thymus_c3" ~ "CD8 CCR9+ (c3)",
  seur_cd8@meta.data$clusters_per_lineage=="CD8_thymus_c4" ~ "CD8 CCR7+ (c4)",
  seur_cd8@meta.data$clusters_per_lineage=="CD8_thymus_c5" ~ "CD8 unknown (c5)",
  seur_cd8@meta.data$clusters_per_lineage=="CD8_pbmc_c0" ~ "PBMC naive (c0)",
  seur_cd8@meta.data$clusters_per_lineage=="CD8_pbmc_c1" ~ "PBMC Tcm (c1)",
  seur_cd8@meta.data$clusters_per_lineage=="CD8_pbmc_c2" ~ "PBMC Tem (c2)",
  seur_cd8@meta.data$clusters_per_lineage=="CD8_pbmc_c3" ~ "PBMC Temra (c3)",
  seur_cd8@meta.data$clusters_per_lineage=="CD8_pbmc_c4" ~ "PBMC Tprolif (c4)"
)
# table(seur_cd8@meta.data[,c("clusters_per_lineage_annotated", "clusters_per_lineage")], useNA="ifany")

# get count matrix
se_cd8 <- SummarizedExperiment(
  assays=seur_cd8@assays[["RNA"]]@counts,
  colData=seur_cd8@meta.data[,c("clusters_per_lineage_annotated", "donor_id", "tissue")]
  )

# get CD8-specific HVGs
# seur_thym_cd8 <- readRDS("./data/seurat_objects/thymus_objects/seurat_filtered_harmony_02_15_23_thymus.cd8.RDS")
# seur_pbmc_cd8 <- readRDS("./data/seurat_objects/pbmc_objects/blood.CD8_03_16_23.RDS")
# hvg_cd8_joint <- union(VariableFeatures(seur_thym_cd8), VariableFeatures(seur_pbmc_cd8))
hvg_cd8 <- VariableFeatures(FindVariableFeatures(seur_cd8))
length(intersect(VariableFeatures(seur_integrated), hvg_cd8)) # 1,243 genes
# length(intersect(hvg_cd8, hvg_cd8_joint))
```

Run metaneighbor with donors as "study" (to check for reproducibility across donors).
```{r run_mtn_branches, fig.width=14, fig.height=14}
mtn <- MetaNeighborUS(
  # var_genes=VariableFeatures(seur_integrated),
  var_genes=hvg_cd8,
  dat=se_cd8,
  study_id=se_cd8$donor_id,
  cell_type=se_cd8$clusters_per_lineage_annotated,
  fast_version=T
  )

# quick heatmap
heatmap.2(mtn,
          # trace
          trace="none",
          # dendrogram
          # Rowv=FALSE,
          # Colv=FALSE,
          # dendrogram="none",
          # superimpose a density histogram on color key
          density.info="none",
          # color scale
          col=rev(colorRampPalette(brewer.pal(11,"RdYlBu"))(100)),
          breaks=seq(0,1,length=101),
          key.xlab = "AUROC",
          key.title="",
          keysize = 1.2,
          # text labels
          main="",
          cexRow=0.6,
          cexCol=0.6,
          # margins
          margins=c(9,9))
```
To note: the PBMC Tem (c2) cluster doesn't have the highest AUROC scores sometimes:
- donors 7, 8, 9, 10, 11 good
- donors (5, 6, 12, 13) and (9, 10) good
- donors (7, 8, 11) and (5, 6, 12, 13) bad
Maybe correlated to the fact that (5, 6, 12, 13) have more naive and fewer effectors?...

Now re-run it with "tissue" as study_id (will be easier to read).
```{r, fig.width=8, fig.height=8}
mtn2 <- MetaNeighborUS(
  # var_genes=VariableFeatures(seur_integrated),
  var_genes=hvg_cd8,
  dat=se_cd8,
  study_id=se_cd8$tissue,
  cell_type=se_cd8$clusters_per_lineage_annotated,
  fast_version=T
  )

# quick heatmap
heatmap.2(mtn2,
          # trace
          trace="none",
          # dendrogram
          # Rowv=FALSE,
          # Colv=FALSE,
          # dendrogram="none",
          # superimpose a density histogram on color key
          density.info="none",
          # color scale
          col=rev(colorRampPalette(brewer.pal(11,"RdYlBu"))(100)),
          breaks=seq(0,1,length=101),
          key.xlab = "AUROC",
          key.title="",
          # keysize = 0.5,
          # text labels
          main="",
          cexRow=0.6,
          cexCol=0.6,
          # margins
          margins=c(10,10)
          )
```




```{r plot-cd8-mtn, fig.width=10, fig.height=10}
library(reshape2)
# library(ggrepel)
cd8_mtn_order <- c(
  "Thymus|CD8 DP (c0)",
  "Thymus|CD8 CCR9+ (c3)",
  "Thymus|CD8 CCR7+ (c4)",
  "PBMC|PBMC naive (c0)",
  "Thymus|CD8 unknown (c5)",
  "PBMC|PBMC Tcm (c1)",
  "Thymus|GNG4+ CD8aa1 (c1)",
  "Thymus|ZNF683+ CD8aa2 (c2)",
  "PBMC|PBMC Tem (c2)",
  "PBMC|PBMC Temra (c3)",
  "PBMC|PBMC Temra (c4)"
)
mtn.df <- melt(mtn2) %>%
  mutate(Var1=factor(Var1, levels=cd8_mtn_order),
         Var2=factor(Var2, levels=cd8_mtn_order)) %>%
  dplyr::rename(auroc=value)


# BUBBLE PLOT
ggplot(mtn.df, aes(x=Var1, y=Var2)) +
  geom_point(aes(size = abs(auroc-0.5), color= auroc))+
  geom_text(data=mtn.df %>% filter(auroc>0.8 | auroc < 0.1) %>% mutate(across("auroc", \(x) round(x,2))), aes(label=auroc), color="white", size=3)+
  scale_size_continuous(limits=c(0,1), breaks=seq(0,1, by=0.2), range = c(0, 20), guide="none")+
  scale_color_gradient2(low="#2166ac", mid="white", high="#a50f15", midpoint=0.5, limits=c(0,1), name="AUROC", breaks=seq(0,1, by=0.2))+
  labs(x="",y="", size="AUROC")+
  theme_cowplot()+
  theme(legend.position="bottom",
        legend.key.width = unit(0.8, 'cm'),
        axis.text = element_text(size=8),
        axis.title=element_text(size=8),
        axis.text.x=element_text(angle=45, hjust=1),
        
  )
# ggsave("~/Projects/phd/data/figures/chapter_03/figs_unfinished/section1/ch3_fig3_cd8_mtn_thymuspbmc.pdf",
#        width=7, height=7)
```



# SESSION INFO
```{r}
sessionInfo()
```


