---
title: "Chapter 3 - CD4 blood figure"
author: "Salomé Carcy"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true

---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 5, fig.height = 5,
                      warning=FALSE, message=FALSE,
                      root.dir = "~/Projects/phd/")
```


# IMPORT

## Import librairies
```{r import-librairies}
library(ggplot2)
library(cowplot)
library(SCpubr)
library(gplots)
library(RColorBrewer)
library(tidyverse)
library(dplyr)
library(Seurat)
library(SummarizedExperiment)
library(MetaNeighbor)

source("~/Projects/phd/scripts/colors_universal.R")
```

## Import data
```{r import-data}
seur_integrated <- readRDS("./data/seurat_objects/seurat_human_integrated_object_23_12_01.rds")
DimPlot(seur_integrated, reduction="umap_integrated", cols = cols_integrated)

seur_pbmc_cd4 <- readRDS("./data/seurat_objects/pbmc_objects/seurat_pbmc_cd4_clean.rds")
```




# FUNCTIONS
```{r define-functions}
lineage_dimplot <- function(seurobj, col_vector, file_name="no"){
  p <- do_DimPlot(
    seurobj,
    group.by="clusters_per_lineage",
    colors.use = col_vector,
    legend.position = "right",
    # raster = T,
    # raster.dpi=2048,
    pt.size = 0.5#,
    # border.size=1.5
  )+
    theme(
      # axis.title=element_text(size=8),
      # axis.text=element_text(size=6),
      legend.key.size = unit(0.4, "cm"),
      legend.title=element_text(size=8),
      legend.text = element_text(size=6)
    )
  if(file_name != "no"){
    ggsave(paste0("~/Projects/phd/data/figures/chapter_03/figs_unfinished/section1/", file_name),
           plot=p,
           width=4, height=2.5, units="in")
  }
}


lineage_clustersperdonor <- function(seurobj, col_vector, file_name="no", figwidth=6){
  p <- seurobj@meta.data %>%
    group_by(Donor, clusters_per_lineage) %>%
    dplyr::count() %>%
    group_by(Donor) %>%
    mutate(sum_per_donor=sum(n),
           freq=n*100/sum_per_donor) %>%
  ggplot(aes(x=factor(Donor, levels=sort(unique(Donor))), y=freq, fill=clusters_per_lineage))+
    geom_bar(stat="identity")+
    scale_fill_manual(values=col_vector)+
    labs(x="Donor", y="% cells", fill="")+
    theme_cowplot()#+
  # theme(axis.text=element_text(size=5/.pt),
  #       axis.title=element_text(size=7/.pt),
  #       legend.text = element_text(size=5/.pt))
  if(file_name != "no"){
    ggsave(paste0("~/Projects/phd/data/figures/chapter_03/figs_unfinished/section1/", file_name),
           plot=p,
           width=figwidth, height=5, units="in")
  }
}
```




# UMAP and distribution per donor

## Umap
```{r umap-cd4}
lineage_dimplot(seurobj = seur_pbmc_cd4, col_vector=cols_pbmc_cd4, file_name="ch3_fig2_cd4_umap.pdf")
```


## Distribution per donor
```{r barplot_per_donor}
# cluster comp per donor
lineage_clustersperdonor(seurobj = seur_pbmc_cd4, col_vector = cols_pbmc_cd4, file_name = "ch3_fig2_cd4_clustersperdonor.pdf")
```




# METANEIGHBOR

## Metaneighbor thymocytes x pbmcs

Prepare data.
```{r prepare-cd4-mtn}
# Get only CD4s
seur_cd4 <- subset(seur_integrated, subset=tcell_lineage=="CD4")
table(seur_cd4@meta.data[,c("clusters_per_lineage", "donor_id")], useNA="ifany") # sanity check

# rename clusters_per_lineage for easier interpretation of metaneighbor
seur_cd4@meta.data$clusters_per_lineage_annotated <- case_when(
  seur_cd4@meta.data$clusters_per_lineage=="CD4_thymus_c0" ~ "CD4 ISP (c0)",
  seur_cd4@meta.data$clusters_per_lineage=="CD4_thymus_c1" ~ "CD4 ISPp (c1)",
  seur_cd4@meta.data$clusters_per_lineage=="CD4_thymus_c2" ~ "DP (c2)",
  seur_cd4@meta.data$clusters_per_lineage=="CD4_thymus_c3" ~ "CD4 CCR9+ (c3)",
  seur_cd4@meta.data$clusters_per_lineage=="CD4_thymus_c4" ~ "CD4 CCR7+ (c4)",
  seur_cd4@meta.data$clusters_per_lineage=="CD4_thymus_c5" ~ "CD4 agonist (c5)",
  seur_cd4@meta.data$clusters_per_lineage=="CD4_thymus_c6" ~ "Thymus Treg (c6)",
  seur_cd4@meta.data$clusters_per_lineage=="CD4_pbmc_c0" ~ "PBMC naive (c0)",
  seur_cd4@meta.data$clusters_per_lineage=="CD4_pbmc_c1" ~ "PBMC naive (c1)",
  seur_cd4@meta.data$clusters_per_lineage=="CD4_pbmc_c2" ~ "PBMC Tcm (c2)",
  seur_cd4@meta.data$clusters_per_lineage=="CD4_pbmc_c3" ~ "PBMC Treg (c3)",
  seur_cd4@meta.data$clusters_per_lineage=="CD4_pbmc_c4" ~ "PBMC Tem (c4)",
  seur_cd4@meta.data$clusters_per_lineage=="CD4_pbmc_c5" ~ "PBMC Temra (c5)"
)
# table(seur_cd4@meta.data[,c("clusters_per_lineage_annotated", "clusters_per_lineage")], useNA="ifany")

# get count matrix
se_cd4 <- SummarizedExperiment(
  assays=seur_cd4@assays[["RNA"]]@counts,
  colData=seur_cd4@meta.data[,c("clusters_per_lineage_annotated", "donor_id", "tissue")]
  )

# get CD4-specific HVGs
# seur_thym_cd4 <- readRDS("./data/seurat_objects/thymus_objects/seurat_filtered_harmony_02_15_23_thymus.cd4.RDS")
# seur_pbmc_cd4 <- readRDS("./data/seurat_objects/pbmc_objects/blood.CD4_03_16_23.RDS")
# hvg_cd4_joint <- union(VariableFeatures(seur_thym_cd4), VariableFeatures(seur_pbmc_cd4))
hvg_cd4 <- VariableFeatures(FindVariableFeatures(seur_cd4))
length(intersect(VariableFeatures(seur_integrated), hvg_cd4))
# length(intersect(hvg_cd4, hvg_cd4_joint))
```

Run metaneighbor with donors as "study" (to check for reproducibility across donors).
```{r run_mtn_branches, fig.width=14, fig.height=14}
mtn <- MetaNeighborUS(
  # var_genes=VariableFeatures(seur_integrated),
  var_genes=hvg_cd4,
  dat=se_cd4,
  study_id=se_cd4$donor_id,
  cell_type=se_cd4$clusters_per_lineage_annotated,
  fast_version=T
  )

# quick heatmap
heatmap.2(mtn,
          # trace
          trace="none",
          # dendrogram
          # Rowv=FALSE,
          # Colv=FALSE,
          # dendrogram="none",
          # superimpose a density histogram on color key
          density.info="none",
          # color scale
          col=rev(colorRampPalette(brewer.pal(11,"RdYlBu"))(100)),
          breaks=seq(0,1,length=101),
          key.xlab = "AUROC",
          key.title="",
          keysize = 1.2,
          # text labels
          main="",
          cexRow=0.6,
          cexCol=0.6,
          # margins
          margins=c(9,9))
```


Now re-run it with "tissue" as study_id (will be easier to read).
```{r, fig.width=8, fig.height=8}
mtn2 <- MetaNeighborUS(
  # var_genes=VariableFeatures(seur_integrated),
  var_genes=hvg_cd4,
  dat=se_cd4,
  study_id=se_cd4$tissue,
  cell_type=se_cd4$clusters_per_lineage_annotated,
  fast_version=T
  )

# quick heatmap
heatmap.2(mtn2,
          # trace
          trace="none",
          # dendrogram
          # Rowv=FALSE,
          # Colv=FALSE,
          # dendrogram="none",
          # superimpose a density histogram on color key
          density.info="none",
          # color scale
          col=rev(colorRampPalette(brewer.pal(11,"RdYlBu"))(100)),
          breaks=seq(0,1,length=101),
          key.xlab = "AUROC",
          key.title="",
          # keysize = 0.5,
          # text labels
          main="",
          cexRow=0.6,
          cexCol=0.6,
          # margins
          margins=c(7,7)
          )
```




```{r plot-cd4-mtn, fig.width=10, fig.height=10}
library(reshape2)
# library(ggrepel)
cd4_mtn_order <- c(
  "Thymus|CD4 ISP (c0)",
  "Thymus|CD4 ISPp (c1)",
  "Thymus|DP (c2)",
  "Thymus|CD4 CCR9+ (c3)",
  "Thymus|CD4 CCR7+ (c4)",
  "PBMC|PBMC naive (c0)",
  "PBMC|PBMC naive (c1)",
  "Thymus|CD4 agonist (c5)",
  "Thymus|Thymus Treg (c6)",
  "PBMC|PBMC Treg (c3)",
  "PBMC|PBMC Tcm (c2)",
  "PBMC|PBMC Tem (c4)",
  "PBMC|PBMC Temra (c5)"
)
mtn.df <- melt(mtn2) %>%
  mutate(Var1=factor(Var1, levels=cd4_mtn_order),
         Var2=factor(Var2, levels=cd4_mtn_order)) %>%
  dplyr::rename(auroc=value)


# BUBBLE PLOT
# library(scales)
ggplot(mtn.df, aes(x=Var1, y=Var2)) +
  geom_point(aes(size = abs(auroc-0.5), color= auroc))+
  geom_text(data=mtn.df %>% filter(auroc>0.8 | auroc < 0.1) %>% mutate(across("auroc", \(x) round(x,2))), aes(label=auroc), color="white", size=3)+
  scale_size_continuous(limits=c(0,1), breaks=seq(0,1, by=0.2), range = c(0, 20), guide="none")+
  scale_color_gradient2(low="#2166ac", mid="white", high="#a50f15", midpoint=0.5, limits=c(0,1), name="AUROC", breaks=seq(0,1, by=0.2))+
  labs(x="",y="", size="AUROC")+
  theme_cowplot()+
  theme(legend.position="bottom",
        legend.key.width = unit(0.8, 'cm'),
        axis.text = element_text(size=8),
        axis.title=element_text(size=8),
        axis.text.x=element_text(angle=45, hjust=1),
        
  )
ggsave("~/Projects/phd/data/figures/chapter_03/figs_unfinished/section1/ch3_fig2_cd4_mtn_thymuspbmc.pdf",
       width=7, height=7)

```


## Analysis 2
```{r analysis-2, fig.width=10, fig.height=10}
```

## Analysis 1
```{r analysis-1}

```


## Analysis 2
```{r analysis-2}

```

## Analysis 3
```{r analysis-3}

```


## Analysis 4
```{r analysis-4}

```




# SESSION INFO
```{r}
sessionInfo()
```


