---
title: "Chapter 3 - gdT figure"
author: "Salomé Carcy"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true

---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 5, fig.height = 5,
                      warning=FALSE, message=FALSE,
                      root.dir = "~/Projects/phd/")
```


# IMPORT

## Import librairies
```{r import-librairies}
library(ggplot2)
library(RColorBrewer)
library(cowplot)
library(tidyverse)
library(dplyr)
library(Seurat)
library(SCpubr)
library(patchwork)
library(SummarizedExperiment)
library(reshape2)
library(MetaNeighbor)
library(patchwork)

source("~/Projects/phd/scripts/colors_universal.R")
```

## Import data
```{r import-data}
seur_integrated <- readRDS("./data/seurat_objects/seurat_human_integrated_object_23_12_01.rds")
DimPlot(seur_integrated, reduction="umap_integrated", cols = cols_integrated)

seur_pbmc_gdt <- readRDS("./data/seurat_objects/pbmc_objects/seurat_pbmc_gdt_clean.rds")
```




# FUNCTIONS
```{r define-functions}
lineage_dimplot <- function(seurobj, col_vector, file_name="no"){
  p <- do_DimPlot(
    seurobj,
    group.by="clusters_per_lineage",
    colors.use = col_vector,
    legend.position = "right",
    # raster = T,
    # raster.dpi=2048,
    pt.size = 0.5#,
    # border.size=1.5
  )+
    theme(
      # axis.title=element_text(size=8),
      # axis.text=element_text(size=6),
      legend.key.size = unit(0.4, "cm"),
      legend.title=element_text(size=8),
      legend.text = element_text(size=6)
    )
  if(file_name != "no"){
    ggsave(paste0("~/Projects/phd/data/figures/chapter_03/figs_unfinished/section1/", file_name),
           plot=p,
           width=4, height=2.5, units="in")
  }
}


lineage_clustersperdonor <- function(seurobj, col_vector, file_name="no", figwidth=6){
  p <- seurobj@meta.data %>%
    group_by(Donor, clusters_per_lineage) %>%
    dplyr::count() %>%
    group_by(Donor) %>%
    mutate(sum_per_donor=sum(n),
           freq=n*100/sum_per_donor) %>%
  ggplot(aes(x=factor(Donor, levels=sort(unique(Donor))), y=freq, fill=clusters_per_lineage))+
    geom_bar(stat="identity")+
    scale_fill_manual(values=col_vector)+
    labs(x="Donor", y="% cells", fill="")+
    theme_cowplot()#+
  # theme(axis.text=element_text(size=5/.pt),
  #       axis.title=element_text(size=7/.pt),
  #       legend.text = element_text(size=5/.pt))
  if(file_name != "no"){
    ggsave(paste0("~/Projects/phd/data/figures/chapter_03/figs_unfinished/section1/", file_name),
           plot=p,
           width=figwidth, height=5, units="in")
  }
}


plot_DotPlot <- function(seurobj, group, features, scaling=T){
  # get plot data
  p <- Seurat::DotPlot(
    seurobj,
    group.by=group,
    features=features,
    scale=scaling
  )
  # plot in personalized way
  p <- ggplot(p$data, aes(x=id, y=features.plot, fill=avg.exp.scaled, size=pct.exp))+
    geom_point(color="black", shape=21)+
    # scale_fill_gradient2(low=scales::muted("blue"), high=scales::muted("red"), name="z-score\nnormalized\navg expression")+
    scale_size_continuous(range=c(0,6), limits=c(0,100), name="%cells\nexpressing\ngene")+
    theme_bw()+
    theme(
      axis.text.y=element_text(face="italic"),
      axis.text.x=element_text(angle=45, hjust=1)
    )+
    labs(y="", x="")
  # different color scale if scaled or not
  if(scaling==T){
    p <- p + scale_fill_gradient2(low=scales::muted("blue"), high=scales::muted("red"), name="z-score\nnormalized\navg expression")
  } else{
    p <- p + viridis::scale_fill_viridis(option="B", direction=-1, name="normalized\navg expression")
  }
  return(p)
}


lineage_nebulosagenes <- function(seuratobj, genes_vector, pgrid_ncol=2, pgrid_size=c(10,30), file_name="no"){
  
  pnebulosa <- list()
  for(gene in genes_vector) {
    p <- SCpubr::do_NebulosaPlot(seuratobj, features = gene, sequential.palette="YlOrRd")
    p <- ggrastr::rasterise(p, layers="Point", dpi=200)
    pnebulosa[[gene]] <- p
  }
  pgrid <- plot_grid(plotlist=pnebulosa, ncol=pgrid_ncol)
  print(pgrid)
  
  if(file_name != "no"){
    ggsave(paste0("~/Projects/phd/data/figures/chapter_03/figs_unfinished/section1/", file_name),
           plot=pgrid,
           width=pgrid_size[1], height=pgrid_size[2])
  }
}
```




# UMAP and distribution per donor

## Umap
```{r umap-gdt}
lineage_dimplot(seurobj = seur_pbmc_gdt, col_vector=cols_pbmc_gdt, file_name="ch3_fig8_gdt_umap.pdf")
```


## Distribution per donor
```{r barplot_per_donor}
# cluster comp per donor
lineage_clustersperdonor(seurobj = seur_pbmc_gdt, col_vector = cols_pbmc_gdt, file_name = "ch3_fig8_gdt_clustersperdonor.pdf", figwidth = 7)
```




# GENES OF INTEREST
## Dotplot genes on interest
The marker genes were chosen based on:
- [Sallusto et al. 1999](https://www.nature.com/articles/44385)
- [Terekhova et al. 2023](https://www.sciencedirect.com/science/article/pii/S1074761323004533)
- [Rose et al. 2023](https://www.nature.com/articles/s42003-023-04747-9)
```{r dotplot-marker-genes, fig.height=5, message=F, warning=F}
genes_dotplot <- c(
  "ZBTB16",
  # ------
  # Tnaive
  "CCR7",
  "SELL",
  "TCF7",
  "LEF1",
  # ------
  # GDT stages
  "CD27",
  "KLRB1",
  "IL18R1",
  # ------
  # Coreceptors
  "CD4",
  "CD8A",
  "CD8B",
  # ------
  # Stage 3 GDTs
  # "CEBPD",
  # ------
  # GDT1
  "TBX21",
  "EOMES",
  "IFNG",
  "TNF",
  "CCR5",
  "IL2RB",
  "SLAMF7",
  # ------
  # GDT17
  "RORC",
  "RORA",
  "CCR6",
  "IL7R",
  "IL23R",
  "KLRG1",
  # ------
  # Cytotoxicity
  "PRF1",
  "GNLY",
  "GZMA",
  "GZMK",
  "GZMB",
  "CX3CR1",
  "NCAM1" # CD56
  )

# get list of cell id to plot (remove clusters with less than 100 cells)
plot_DotPlot(
  seurobj=seur_pbmc_gdt,
  group="clusters_per_lineage",
  features=genes_dotplot,
  scaling=F
  )
# ggsave("~/Projects/phd/data/figures/chapter_03/figs_unfinished/section1/ch3_fig6_gdt_markergenes.pdf", width=5, height=6)
```


## Nebulosa plot genes of interest
```{r, fig.height=8, fig.width=6}
lineage_nebulosagenes(
  seur_pbmc_gdt,
  genes_vector=c("ZBTB16", "CD27", "KLRB1", "IL18R1", "GZMK", "GZMB"),
  # file_name="ch3_fig6_gdt_nebulosaplots.pdf",
  pgrid_ncol=2,
  pgrid_size=c(10, 15)
)
```




# METANEIGHBOR

## Metaneighbor thymocytes x pbmcs

Prepare data.
```{r prepare-gdt-mtn}
# Get only GDTs
seur_gdt <- subset(seur_integrated, subset=tcell_lineage=="GD")
table(seur_gdt@meta.data[,c("clusters_per_lineage", "donor_id")], useNA="ifany") # sanity check

# rename clusters_per_lineage for easier interpretation of metaneighbor
seur_gdt@meta.data$clusters_per_lineage_annotated <- case_when(
  seur_gdt@meta.data$clusters_per_lineage=="GDT_thymus_c0" ~ "GDT DP (c0)",
  seur_gdt@meta.data$clusters_per_lineage=="GDT_thymus_c1" ~ "GDT TCRrearrange (c1)",
  seur_gdt@meta.data$clusters_per_lineage=="GDT_thymus_c2" ~ "GDT TCRrearrange (c2)",
  seur_gdt@meta.data$clusters_per_lineage=="GDT_thymus_c3" ~ "GDT TCRsignaling (c3)",
  seur_gdt@meta.data$clusters_per_lineage=="GDT_thymus_c4" ~ "GDT ISPp (c4)",
  seur_gdt@meta.data$clusters_per_lineage=="GDT_thymus_c5" ~ "GDT immature (c5)",
  seur_gdt@meta.data$clusters_per_lineage=="GDT_thymus_c6" ~ "GDT immature (c6)",
  seur_gdt@meta.data$clusters_per_lineage=="GDT_thymus_c7" ~ "GDT memory (c7)",
  seur_gdt@meta.data$clusters_per_lineage=="GDT_pbmc_c0" ~ "GDT naive (c0)",
  seur_gdt@meta.data$clusters_per_lineage=="GDT_pbmc_c1" ~ "GDT (c1)",
  seur_gdt@meta.data$clusters_per_lineage=="GDT_pbmc_c2" ~ "GDT (c2)",
  seur_gdt@meta.data$clusters_per_lineage=="GDT_pbmc_c3" ~ "GDT GZMB+ (c3)"
)
# table(seur_gdt@meta.data[,c("clusters_per_lineage_annotated", "clusters_per_lineage")], useNA="ifany")

# get count matrix
se_gdt <- SummarizedExperiment(
  assays=seur_gdt@assays[["RNA"]]@counts,
  colData=seur_gdt@meta.data[,c("clusters_per_lineage_annotated", "donor_id", "tissue")]
  )

# get GDT-specific HVGs
# seur_thym_gdt <- readRDS("./data/seurat_objects/thymus_objects/seurat_filtered_harmony_02_15_23_thymus.gdt.RDS")
# seur_pbmc_gdt <- readRDS("./data/seurat_objects/pbmc_objects/blood.GDT_03_16_23.RDS")
# hvg_gdt_joint <- union(VariableFeatures(seur_thym_gdt), VariableFeatures(seur_pbmc_gdt))
hvg_gdt <- VariableFeatures(FindVariableFeatures(seur_gdt))
length(intersect(VariableFeatures(seur_integrated), hvg_gdt)) # 1028 genes
# length(intersect(hvg_gdt, hvg_gdt_joint))
```

Run metaneighbor with donors as "study" (to check for reproducibility across donors).
```{r run_mtn_branches, fig.width=14, fig.height=14}
mtn <- MetaNeighborUS(
  # var_genes=VariableFeatures(seur_integrated),
  var_genes=hvg_gdt,
  dat=se_gdt,
  study_id=se_gdt$donor_id,
  cell_type=se_gdt$clusters_per_lineage_annotated,
  fast_version=T
  )

# quick heatmap
heatmap.2(mtn,
          # trace
          trace="none",
          # dendrogram
          # Rowv=FALSE,
          # Colv=FALSE,
          # dendrogram="none",
          # superimpose a density histogram on color key
          density.info="none",
          # color scale
          col=rev(colorRampPalette(brewer.pal(11,"RdYlBu"))(100)),
          breaks=seq(0,1,length=101),
          key.xlab = "AUROC",
          key.title="",
          keysize = 1.2,
          # text labels
          main="",
          cexRow=0.6,
          cexCol=0.6,
          # margins
          margins=c(9,9))
```
To note: the PBMC GDT (c0-2) don't always have the best AUROC with each other (is it because it's overclustered?).

Now re-run it with "tissue" as study_id (will be easier to read).
```{r, fig.width=8, fig.height=8}
# test by removing PBMC cluster c0
# se_gdt_test <- se_gdt[,rownames(colData(se_gdt)[colData(se_gdt)$clusters_per_lineage_annotated != "GDT (c1)",])]

mtn2 <- MetaNeighborUS(
  # var_genes=VariableFeatures(seur_integrated),
  var_genes=hvg_gdt,
  dat=se_gdt,
  study_id=se_gdt$tissue,
  cell_type=se_gdt$clusters_per_lineage_annotated,
  fast_version=T
  )

# quick heatmap
heatmap.2(mtn2,
          # trace
          trace="none",
          # dendrogram
          # Rowv=FALSE,
          # Colv=FALSE,
          # dendrogram="none",
          # superimpose a density histogram on color key
          density.info="none",
          # color scale
          col=rev(colorRampPalette(brewer.pal(11,"RdYlBu"))(100)),
          breaks=seq(0,1,length=101),
          key.xlab = "AUROC",
          key.title="",
          # keysize = 0.5,
          # text labels
          main="",
          cexRow=0.6,
          cexCol=0.6,
          # margins
          margins=c(10,10)
          )
```




```{r plot-gdt-mtn, fig.width=10, fig.height=10}
# library(ggrepel)
gdt_mtn_order <- c(
  "Thymus|GDT DP (c0)",
  "Thymus|GDT CD8aa (c1)",
  "Thymus|GDT CCR9+ (c2)",
  "Thymus|GDT CCR7+ (c3)",
  "Thymus|GDT CCR7+ (c4)",
  "Thymus|GDT CCR7+ (c5)",
  "Thymus|GDT memory (c6)",
  "PBMC|GDT (c0)",
  "PBMC|GDT (c1)",
  "PBMC|GDT (c2)",
  "PBMC|GDT GZMB+ (c3)"
)

# get nb of cells
ncells.df <- colData(se_gdt) %>%
  as_data_frame() %>%
  group_by(tissue, clusters_per_lineage_annotated) %>%
  dplyr::count() %>%
  mutate(var=paste(tissue, clusters_per_lineage_annotated, sep="|")) %>%
  ungroup() %>%
  dplyr::select(n, var)

# final df for plotting
mtn.df <- melt(mtn2) %>%
  dplyr::rename(auroc=value) %>%
  left_join(ncells.df, by=c("Var1"="var")) %>%
  dplyr::rename(Var1_ncells=n) %>%
  left_join(ncells.df, by=c("Var2"="var")) %>%
  dplyr::rename(Var2_ncells=n) %>%
  mutate(Var1=factor(Var1, levels=gdt_mtn_order),
         Var2=factor(Var2, levels=gdt_mtn_order))
  
# barplots
bp.x <- ggplot(data=mtn.df%>% select(Var1,Var1_ncells) %>% distinct(),
               aes(x=Var1, y=Var1_ncells))+
  geom_bar(stat="identity", fill="#bdbdbd") + theme_cowplot()+
  scale_x_discrete(position="top")+
  labs(y="# cells")+
  theme(axis.text = element_text(size=8),
        # axis.text.x = element_text(size = 8, angle=45, hjust=0),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.x = element_blank(),
        axis.line.x=element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y=element_text(size=8),
        legend.position = "none")

bp.y <- ggplot(data=mtn.df%>% select(Var2,Var2_ncells) %>% distinct(),
               aes(x=Var2, y=Var2_ncells))+
  geom_bar(stat="identity", fill="#bdbdbd") +
  scale_x_discrete(position="top") +
  labs(y="# cells")+ coord_flip() + theme_cowplot()+
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size=8),
        axis.text.y=element_blank(),
        axis.ticks.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_text(size=8),
        axis.line.y=element_blank(),
        legend.position = "none")


# BUBBLE PLOT
hm.clean <- ggplot(mtn.df, aes(x=Var1, y=Var2)) +
  geom_point(aes(size = abs(auroc-0.5), color= auroc))+
  geom_text(data=mtn.df %>% filter(auroc>0.8 | auroc < 0.1) %>% mutate(across("auroc", \(x) round(x,2))), aes(label=auroc), color="white", size=3)+
  scale_size_continuous(limits=c(0,1), breaks=seq(0,1, by=0.2), range = c(0, 20), guide="none")+
  scale_color_gradient2(low="#2166ac", mid="white", high="#a50f15", midpoint=0.5, limits=c(0,1), name="AUROC", breaks=seq(0,1, by=0.2))+
  labs(x="",y="", size="AUROC")+
  theme_cowplot()+
  theme(legend.position="bottom",
        legend.key.width = unit(0.8, 'cm'),
        axis.text = element_text(size=8),
        axis.title=element_text(size=8),
        axis.text.x=element_text(angle=45, hjust=1)
  )
(bp.x+plot_spacer() + plot_layout(widths = c(5, 1))) / (hm.clean + bp.y + plot_layout(widths = c(5, 1))) + plot_layout(heights = c(1, 5))
# ggsave("~/Projects/phd/data/figures/chapter_03/figs_unfinished/section1/ch3_fig6_gdt_mtn_thymuspbmc.pdf",
#        width=8, height=8)
```




# SESSION INFO
```{r}
sessionInfo()
```


