---
title: "Chapter 3 - gdT figure"
author: "Salomé Carcy"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true

---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 5, fig.height = 5,
                      warning=FALSE, message=FALSE,
                      root.dir = "~/Projects/phd/")
```


# IMPORT

## Import librairies
```{r import-librairies}
library(ggplot2)
library(RColorBrewer)
library(cowplot)
library(tidyverse)
library(dplyr)
library(Seurat)
library(SCpubr)
library(patchwork)
library(SummarizedExperiment)
library(reshape2)
library(MetaNeighbor)
library(patchwork)

source("~/Projects/phd/scripts/colors_universal.R")
```

## Import data
```{r import-data}
seur_integrated <- readRDS("./data/seurat_objects/seurat_human_integrated_object_23_12_01.rds")
DimPlot(seur_integrated, reduction="umap_integrated", cols = cols_integrated)

seur_pbmc_gdt <- readRDS("./data/seurat_objects/pbmc_objects/seurat_pbmc_gdt_clean.rds")
```




# FUNCTIONS
```{r define-functions}
lineage_dimplot <- function(seurobj, col_vector, file_name="no"){
  p <- do_DimPlot(
    seurobj,
    group.by="clusters_per_lineage",
    colors.use = col_vector,
    legend.position = "right",
    # raster = T,
    # raster.dpi=2048,
    pt.size = 0.5#,
    # border.size=1.5
  )+
    theme(
      # axis.title=element_text(size=8),
      # axis.text=element_text(size=6),
      legend.key.size = unit(0.4, "cm"),
      legend.title=element_text(size=8),
      legend.text = element_text(size=6)
    )
  if(file_name != "no"){
    ggsave(paste0("~/Projects/phd/data/figures/chapter_03/figs_unfinished/section1/", file_name),
           plot=p,
           width=4, height=2.5, units="in")
  }
}


lineage_clustersperdonor <- function(seurobj, col_vector, file_name="no", figwidth=6){
  p <- seurobj@meta.data %>%
    group_by(Donor, clusters_per_lineage) %>%
    dplyr::count() %>%
    group_by(Donor) %>%
    mutate(sum_per_donor=sum(n),
           freq=n*100/sum_per_donor) %>%
  ggplot(aes(x=factor(Donor, levels=sort(unique(Donor))), y=freq, fill=clusters_per_lineage))+
    geom_bar(stat="identity")+
    scale_fill_manual(values=col_vector)+
    labs(x="Donor", y="% cells", fill="")+
    theme_cowplot()#+
  # theme(axis.text=element_text(size=5/.pt),
  #       axis.title=element_text(size=7/.pt),
  #       legend.text = element_text(size=5/.pt))
  if(file_name != "no"){
    ggsave(paste0("~/Projects/phd/data/figures/chapter_03/figs_unfinished/section1/", file_name),
           plot=p,
           width=figwidth, height=5, units="in")
  }
}


plot_DotPlot <- function(seurobj, group, features, scaling=T){
  # get plot data
  p <- Seurat::DotPlot(
    seurobj,
    group.by=group,
    features=features,
    scale=scaling
  )
  # plot in personalized way
  p <- ggplot(p$data, aes(x=id, y=features.plot, fill=avg.exp.scaled, size=pct.exp))+
    geom_point(color="black", shape=21)+
    # scale_fill_gradient2(low=scales::muted("blue"), high=scales::muted("red"), name="z-score\nnormalized\navg expression")+
    scale_size_continuous(range=c(0,6), limits=c(0,100), name="%cells\nexpressing\ngene")+
    theme_bw()+
    theme(
      axis.text.y=element_text(face="italic"),
      axis.text.x=element_text(angle=45, hjust=1)
    )+
    labs(y="", x="")
  # different color scale if scaled or not
  if(scaling==T){
    p <- p + scale_fill_gradient2(low=scales::muted("blue"), high=scales::muted("red"), name="z-score\nnormalized\navg expression")
  } else{
    p <- p + viridis::scale_fill_viridis(option="B", direction=-1, name="normalized\navg expression")
  }
  return(p)
}


lineage_nebulosagenes <- function(seuratobj, genes_vector, pgrid_ncol=2, pgrid_size=c(10,30), file_name="no"){
  
  pnebulosa <- list()
  for(gene in genes_vector) {
    p <- SCpubr::do_NebulosaPlot(seuratobj, features = gene, sequential.palette="YlOrRd")
    p <- ggrastr::rasterise(p, layers="Point", dpi=200)
    pnebulosa[[gene]] <- p
  }
  pgrid <- plot_grid(plotlist=pnebulosa, ncol=pgrid_ncol)
  print(pgrid)
  
  if(file_name != "no"){
    ggsave(paste0("~/Projects/phd/data/figures/chapter_03/figs_unfinished/section1/", file_name),
           plot=pgrid,
           width=pgrid_size[1], height=pgrid_size[2])
  }
}
```




# UMAP and distribution per donor

## Umap
```{r umap-gdt}
lineage_dimplot(seurobj = seur_pbmc_gdt, col_vector=cols_pbmc_gdt, file_name="ch3_fig8_gdt_umap.pdf")
```


## Distribution per donor
```{r barplot_per_donor}
# cluster comp per donor
lineage_clustersperdonor(seurobj = seur_pbmc_gdt, col_vector = cols_pbmc_gdt, file_name = "ch3_fig8_gdt_clustersperdonor.pdf", figwidth = 7)
```




# GENES OF INTEREST
## Dotplot genes on interest
The marker genes were chosen based on:
- [Sallusto et al. 1999](https://www.nature.com/articles/44385)
- [Terekhova et al. 2023](https://www.sciencedirect.com/science/article/pii/S1074761323004533)
- [Rose et al. 2023](https://www.nature.com/articles/s42003-023-04747-9)
```{r dotplot-marker-genes, fig.height=5, message=F, warning=F}
genes_dotplot <- c(
  "ZBTB16",
  # "ZNF683",
  # ------
  # GD naive
  "CCR7",
  "SELL",
  "TCF7",
  "LEF1",
  "IL7R",
  # "CD27",
  # ------
  # VD2VG9 GDT stages
  "CD4",
  "KLRB1",
  # "IL18R1",
  # ------
  # Coreceptors
  # "CD8A",
  # "CD8B",
  # ------
  # GDT1
  "TBX21",
  "EOMES",
  "IFNG",
  # "TNF",
  # "CCR5",
  "IL2RB",
  "SLAMF7",
  # ------
  # GDT17
  "RORC",
  "RORA",
  "CCR6",
  "IL23R",
  # ------
  # GZMK vs GZMB
  "GZMK",
  "GNLY",
  "GZMA",
  "PRF1",
  "GZMB",
  "CX3CR1",
  # "NCAM1", # CD56
  # -------
  # Prolif
  "MKI67",
  "CDK1"
  )

# get list of cell id to plot (remove clusters with less than 100 cells)
plot_DotPlot(
  seurobj=seur_pbmc_gdt,
  group="clusters_per_lineage",
  features=genes_dotplot,
  scaling=F
  )
# ggsave("~/Projects/phd/data/figures/chapter_03/figs_unfinished/section1/ch3_fig8_gdt_markergenes.pdf", width=5, height=6)
```


## Nebulosa plot genes of interest
```{r nebulosa-plot, fig.height=8, fig.width=6}
lineage_nebulosagenes(
  seur_pbmc_gdt,
  genes_vector=c("CCR7", "ZBTB16", "GZMK", "GZMB"),
  # file_name="ch3_fig8_gdt_nebulosaplots.pdf",
  pgrid_ncol=2,
  pgrid_size=c(10, 10)
)
```

## V delta gene usage
First clean up data a bit
```{r TCRdelta-usage}
gd.tcr <- seur_pbmc_gdt@meta.data %>%
  rownames_to_column("cellid") %>%
  as_tibble() %>%
  select(cellid, clusters_per_lineage, grep("_gene_Dominant", colnames(seur_pbmc_gdt@meta.data), value=T)) %>%
  dplyr::rename(TRGV = TCR_Alpha_Gamma_V_gene_Dominant,
                TRGJ = TCR_Alpha_Gamma_J_gene_Dominant,
                TRDV = TCR_Beta_Delta_V_gene_Dominant,
                TRDD = TCR_Beta_Delta_D_gene_Dominant,
                TRDJ = TCR_Beta_Delta_J_gene_Dominant) %>%
  na.omit(!clusters_per_lineage) %>%
  mutate(TRGV = str_remove(TRGV, pattern = "\\*[^.]*$"),
         TRGJ = str_remove(TRGJ, pattern = "\\*[^.]*$"),
         TRDV = str_remove(TRDV, pattern = "\\*[^.]*$"),
         TRDD = str_remove(TRDD, pattern = "\\*[^.]*$"),
         TRDJ = str_remove(TRDJ, pattern = "\\*[^.]*$")) %>%
  filter(!TRGV %in% grep("TRA", TRDV, value=T)) %>%
  filter(!TRDD %in% grep("TRB", TRDD, value=T)) %>%
  filter(!TRDJ %in% grep("TRB", TRDJ, value=T))
dim(gd.tcr)
```
We have TCR information for 1,569 cells out of 10,160. Let's highlight TRDV1, TRDV2, TRDV3
```{r gdt-tcr-highlight-trdv, fig.height=7, fig.width=9}
highlight_TCRusage <- function(seurobj=seur_pbmc_gdt, cells, ptitle, cellscolor){
  p <- do_DimPlot(
      sample = seurobj,
      cells.highlight = cells,
      plot.title = ptitle,
      legend.position = "none",
      na.value = "grey90",
      colors.use = cellscolor
    )
  p <- ggrastr::rasterise(p, layers="Point", dpi=300)
  return(p)
}

# highlight TRDV1, TRDV2, TRDV3 and TRGV9
plot_grid(
  highlight_TCRusage(cells = pull(gd.tcr[gd.tcr$TRDV=="TRDV1","cellid"]), ptitle = "TRDV1", cellscolor = "#a40000"),
  highlight_TCRusage(cells = pull(gd.tcr[gd.tcr$TRDV=="TRDV2","cellid"]), ptitle = "TRDV2", cellscolor = "blue"),
  highlight_TCRusage(cells = pull(gd.tcr[gd.tcr$TRDV=="TRDV2"&gd.tcr$TRGV=="TRGV9","cellid"]), ptitle = "VD2+VG9+", cellscolor = "lightblue"),
  highlight_TCRusage(cells = pull(gd.tcr[gd.tcr$TRDV=="TRDV3","cellid"]), ptitle = "TRDV3", cellscolor = "#318f49"),
  highlight_TCRusage(cells = pull(gd.tcr[gd.tcr$TRGV=="TRGV9","cellid"]), ptitle = "VG9", cellscolor = "gold"),
  nrow=2)
# ggsave("~/Projects/phd/data/figures/chapter_03/figs_unfinished/section1/ch3_fig8_gdt_tcr_trdv.pdf", width=9, height=6)

# highlight on integrated umap
# plot_grid(
#   highlight_TCRusage(seurobj=seur_integrated, cells = pull(gd.tcr[gd.tcr$TRDV=="TRDV1","cellid"]), ptitle = "TRDV1", cellscolor = "#a40000"),
#   # highlight_TCRusage(seurobj=seur_integrated, cells = pull(gd.tcr[gd.tcr$TRDV=="TRDV2","cellid"]), ptitle = "TRDV2", cellscolor = "blue"),
#   highlight_TCRusage(seurobj=seur_integrated, cells = pull(gd.tcr[gd.tcr$TRDV=="TRDV2"&gd.tcr$TRGV=="TRGV9","cellid"]), ptitle = "VD2+VG9+", cellscolor = "lightblue"),
#   # highlight_TCRusage(seurobj=seur_integrated, cells = pull(gd.tcr[gd.tcr$TRDV=="TRDV3","cellid"]), ptitle = "TRDV3", cellscolor = "#318f49"),
#   highlight_TCRusage(seurobj=seur_integrated, cells = pull(gd.tcr[gd.tcr$TRGV=="TRGV9","cellid"]), ptitle = "VG9", cellscolor = "gold"),
#   nrow=2)
```




# METANEIGHBOR

## Metaneighbor thymocytes x pbmcs

Prepare data.
```{r prepare-gdt-mtn}
# Get only GDTs
seur_gdt <- subset(seur_integrated, subset=tcell_lineage=="GD")
table(seur_gdt@meta.data[,c("clusters_per_lineage", "donor_id")], useNA="ifany") # sanity check

# rename clusters_per_lineage for easier interpretation of metaneighbor
seur_gdt@meta.data$clusters_per_lineage_annotated <- case_when(
  seur_gdt@meta.data$clusters_per_lineage=="GDT_thymus_c0" ~ "GDT DP (c0)",
  seur_gdt@meta.data$clusters_per_lineage=="GDT_thymus_c1" ~ "GDT TCRrearrange (c1)",
  seur_gdt@meta.data$clusters_per_lineage=="GDT_thymus_c2" ~ "GDT TCRrearrange (c2)",
  seur_gdt@meta.data$clusters_per_lineage=="GDT_thymus_c3" ~ "GDT TCRsignaling (c3)",
  seur_gdt@meta.data$clusters_per_lineage=="GDT_thymus_c4" ~ "GDT ISPp (c4)",
  seur_gdt@meta.data$clusters_per_lineage=="GDT_thymus_c5" ~ "GDT immature (c5)",
  seur_gdt@meta.data$clusters_per_lineage=="GDT_thymus_c6" ~ "GDT immature (c6)",
  seur_gdt@meta.data$clusters_per_lineage=="GDT_thymus_c7" ~ "GDT memory (c7)",
  seur_gdt@meta.data$clusters_per_lineage=="GDT_pbmc_c0" ~ "GDT naive (c0)",
  seur_gdt@meta.data$clusters_per_lineage=="GDT_pbmc_c1" ~ "GDT VD2VG9 GZMK+ (c1)",
  seur_gdt@meta.data$clusters_per_lineage=="GDT_pbmc_c2" ~ "GDT GZMK+ (c2)",
  seur_gdt@meta.data$clusters_per_lineage=="GDT_pbmc_c3" ~ "GDT GZMB+ (c3)",
  seur_gdt@meta.data$clusters_per_lineage=="GDT_pbmc_c4" ~ "GDT prolif (c4)"
)
# table(seur_gdt@meta.data[,c("clusters_per_lineage_annotated", "clusters_per_lineage")], useNA="ifany")

# get count matrix
se_gdt <- SummarizedExperiment(
  assays=seur_gdt@assays[["RNA"]]@counts,
  colData=seur_gdt@meta.data[,c("clusters_per_lineage_annotated", "donor_id", "tissue")]
  )

# get GDT-specific HVGs
# seur_thym_gdt <- readRDS("./data/seurat_objects/thymus_objects/seurat_filtered_harmony_02_15_23_thymus.gdt.RDS")
# seur_pbmc_gdt <- readRDS("./data/seurat_objects/pbmc_objects/blood.GDT_03_16_23.RDS")
# hvg_gdt_joint <- union(VariableFeatures(seur_thym_gdt), VariableFeatures(seur_pbmc_gdt))
hvg_gdt <- VariableFeatures(FindVariableFeatures(seur_gdt))
length(intersect(VariableFeatures(seur_integrated), hvg_gdt)) # 1116 genes
# length(intersect(hvg_gdt, hvg_gdt_joint))
```

Run metaneighbor with donors as "study" (to check for reproducibility across donors).
```{r run_mtn_branches, fig.width=14, fig.height=14}
mtn <- MetaNeighborUS(
  # var_genes=VariableFeatures(seur_integrated),
  var_genes=hvg_gdt,
  dat=se_gdt,
  study_id=se_gdt$donor_id,
  cell_type=se_gdt$clusters_per_lineage_annotated,
  fast_version=T
  )

# quick heatmap
heatmap.2(mtn,
          # trace
          trace="none",
          # dendrogram
          # Rowv=FALSE,
          # Colv=FALSE,
          # dendrogram="none",
          # superimpose a density histogram on color key
          density.info="none",
          # color scale
          col=rev(colorRampPalette(brewer.pal(11,"RdYlBu"))(100)),
          breaks=seq(0,1,length=101),
          key.xlab = "AUROC",
          key.title="",
          keysize = 1.2,
          # text labels
          main="",
          cexRow=0.6,
          cexCol=0.6,
          # margins
          margins=c(9,9))
```

Now re-run it with "tissue" as study_id (will be easier to read).
```{r, fig.width=8, fig.height=8}
mtn2 <- MetaNeighborUS(
  # var_genes=VariableFeatures(seur_integrated),
  var_genes=hvg_gdt,
  dat=se_gdt,
  study_id=se_gdt$tissue,
  cell_type=se_gdt$clusters_per_lineage_annotated,
  fast_version=T
  )

# quick heatmap
heatmap.2(mtn2,
          # trace
          trace="none",
          # dendrogram
          # Rowv=FALSE,
          # Colv=FALSE,
          # dendrogram="none",
          # superimpose a density histogram on color key
          density.info="none",
          # color scale
          col=rev(colorRampPalette(brewer.pal(11,"RdYlBu"))(100)),
          breaks=seq(0,1,length=101),
          key.xlab = "AUROC",
          key.title="",
          # keysize = 0.5,
          # text labels
          main="",
          cexRow=0.6,
          cexCol=0.6,
          # margins
          margins=c(10,10)
          )
```


```{r plot-gdt-mtn, fig.width=10, fig.height=10}
# library(ggrepel)
gdt_mtn_order <- c(
  "Thymus|GDT DP (c0)",
  "Thymus|GDT TCRrearrange (c1)",
  "PBMC|GDT naive (c0)",
  "Thymus|GDT TCRrearrange (c2)",
  "Thymus|GDT TCRsignaling (c3)",
  "Thymus|GDT ISPp (c4)",
  "PBMC|GDT prolif (c4)",
  "Thymus|GDT immature (c5)",
  "Thymus|GDT immature (c6)",
  "Thymus|GDT memory (c7)",
  "PBMC|GDT VD2VG9 GZMK+ (c1)",
  "PBMC|GDT GZMK+ (c2)",
  "PBMC|GDT GZMB+ (c3)"
)

# get nb of cells
ncells.df <- colData(se_gdt) %>%
  as_data_frame() %>%
  group_by(tissue, clusters_per_lineage_annotated) %>%
  dplyr::count() %>%
  mutate(var=paste(tissue, clusters_per_lineage_annotated, sep="|")) %>%
  ungroup() %>%
  dplyr::select(n, var)

# final df for plotting
mtn.df <- melt(mtn2) %>%
  dplyr::rename(auroc=value) %>%
  left_join(ncells.df, by=c("Var1"="var")) %>%
  dplyr::rename(Var1_ncells=n) %>%
  left_join(ncells.df, by=c("Var2"="var")) %>%
  dplyr::rename(Var2_ncells=n) %>%
  mutate(Var1=factor(Var1, levels=gdt_mtn_order),
         Var2=factor(Var2, levels=gdt_mtn_order))
  
# barplots
bp.x <- ggplot(data=mtn.df%>% select(Var1,Var1_ncells) %>% distinct(),
               aes(x=Var1, y=Var1_ncells))+
  geom_bar(stat="identity", fill="#bdbdbd") + theme_cowplot()+
  scale_x_discrete(position="top")+
  labs(y="# cells")+
  theme(axis.text = element_text(size=8),
        # axis.text.x = element_text(size = 8, angle=45, hjust=0),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.x = element_blank(),
        axis.line.x=element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y=element_text(size=8),
        legend.position = "none")

bp.y <- ggplot(data=mtn.df%>% select(Var2,Var2_ncells) %>% distinct(),
               aes(x=Var2, y=Var2_ncells))+
  geom_bar(stat="identity", fill="#bdbdbd") +
  scale_x_discrete(position="top") +
  labs(y="# cells")+ coord_flip() + theme_cowplot()+
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size=8),
        axis.text.y=element_blank(),
        axis.ticks.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_text(size=8),
        axis.line.y=element_blank(),
        legend.position = "none")


# BUBBLE PLOT
hm.clean <- ggplot(mtn.df, aes(x=Var1, y=Var2)) +
  geom_point(aes(size = abs(auroc-0.5), color= auroc))+
  geom_text(data=mtn.df %>% filter(auroc>0.8 | auroc < 0.1) %>% mutate(across("auroc", \(x) round(x,2))), aes(label=auroc), color="white", size=3)+
  scale_size_continuous(limits=c(0,1), breaks=seq(0,1, by=0.2), range = c(0, 20), guide="none")+
  scale_color_gradient2(low="#2166ac", mid="white", high="#a50f15", midpoint=0.5, limits=c(0,1), name="AUROC", breaks=seq(0,1, by=0.2))+
  labs(x="",y="", size="AUROC")+
  theme_cowplot()+
  theme(legend.position="bottom",
        legend.key.width = unit(0.8, 'cm'),
        axis.text = element_text(size=8),
        axis.title=element_text(size=8),
        axis.text.x=element_text(angle=45, hjust=1)
  )
(bp.x+plot_spacer() + plot_layout(widths = c(5, 1))) / (hm.clean + bp.y + plot_layout(widths = c(5, 1))) + plot_layout(heights = c(1, 5))

# ggsave(
#   # plot = hm.clean,
#   filename = "~/Projects/phd/data/figures/chapter_03/figs_unfinished/section1/ch3_fig8_gdt_mtn_thymuspbmc.pdf",
#   width = 9,
#   height = 9
# )
```


## Metaneighbor on equal number of cells
Let's try to randomly sample cells from each cluster, to have equal number of cells per cluster, and see how robust the metaneighbor results are. Let's run it on the list of HVGs.
```{r gdt-metaneighbor-subset-100sampling, fig.height=7, fig.width=7}
# Randomly sample x number of cells per cluster
randomcells_list <- list()
mtn_list <- list()
for(i in 1:1000){
  cat("\nRandom sampling #", i, "\n")
  randomcells_vector <- c()
  mtn_subset <- NULL
  
  # RANDOMLY SAMPLE CELLS
  for(cluster in unique(se_gdt$clusters_per_lineage_annotated)){
    # print(cluster)
    randomcells <- NULL

    # if it's a thymus cluster
    if (cluster %in% c(
      "GDT DP (c0)",
      "GDT TCRrearrange (c1)",
      "GDT TCRrearrange (c2)",
      "GDT TCRsignaling (c3)",
      "GDT ISPp (c4)",
      "GDT immature (c5)",
      "GDT immature (c6)",
      "GDT memory (c7)"
    )) {
      nbcells_to_sample <- 27 # 27
      allcells <- rownames(colData(se_gdt)[colData(se_gdt)$clusters_per_lineage_annotated==cluster,])
      # cat("Total nb of cells:", length(allcells), "\n")
      # set.seed(123)
      randomcells <- sample(allcells, nbcells_to_sample)
      # cat("Nb of sampled cells:", length(randomcells), "\n")
      # sanity check they all come from same cluster
      # cat("Cluster sampled cells come from:", unique(colData(se_gdt)[randomcells, "clusters_per_lineage_annotated"]), "\n\n")
    }
    
    # if it's a pbmc cluster
    else if (cluster %in% c(
      "GDT naive (c0)",
      "GDT VD2VG9 GZMK+ (c1)",
      "GDT GZMK+ (c2)",
      "GDT GZMB+ (c3)",
      "GDT prolif (c4)"
    )){
      nbcells_to_sample <- 27 # 56
      allcells <- rownames(colData(se_gdt)[colData(se_gdt)$clusters_per_lineage_annotated==cluster,])
      # cat("Total nb of cells:", length(allcells), "\n")
      # set.seed(123)
      randomcells <- sample(allcells, nbcells_to_sample)
      # cat("Nb of sampled cells:", length(randomcells), "\n")
      # cat("Cluster sampled cells come from:", unique(colData(se_gdt)[randomcells, "clusters_per_lineage_annotated"]), "\n\n")
      
    }
    
    randomcells_vector <- c(randomcells_vector, randomcells)
    # print(length(randomcells_vector))
  }
  
  # Subset Summarized Experiment object to the randomly selected cells
  se_gdt_subset <- se_gdt[,randomcells_vector]
  table(colData(se_gdt_subset)$clusters_per_lineage_annotated)
  
  
  # re-run metaneighbor
  mtn_subset <- MetaNeighborUS(var_genes=hvg_gdt,
                               dat=se_gdt_subset,
                               study_id=se_gdt_subset$tissue,
                               cell_type=se_gdt_subset$clusters_per_lineage_annotated,
                               fast_version=T)
  
  # save
  randomcells_list[[i]] <- randomcells_vector
  mtn_list[[i]] <- mtn_subset
}

length(randomcells_list)
length(mtn_list)

# sanity check that different cells were sampled at each round
table(randomcells_list[[1]] %in% randomcells_list[[2]])

# get mean and sd of all the metaneighbor matrices
mtn_mean <- apply(simplify2array(mtn_list), 1:2, mean)
mtn_sd <- apply(simplify2array(mtn_list), 1:2, sd)

# check the distribution of SD (very low SD across random samplings, which is good!)
ggplot(melt(mtn_sd), aes(x=value))+
  geom_histogram(bins=200, color="black", fill="white")+
  scale_x_continuous(expand=c(0,0), limits=c(0,1))+
  # scale_y_continuous(expand=c(0,0), limits=c(0,70))+
  labs(x="AUROC standard deviation", y="# AUROC values", title="Distribution of SD of AUROC values")+
  theme(panel.grid.minor=element_blank(),
        panel.grid.major=element_blank())
# ggsave("./data/figures/chapter_03/figs_unfinished/suppfigures/gdt_crosspecies_100randomcellsampling_sd_auroc_hist.pdf", width=7, height=6)

# get nb of cells in subsetted data
ncells.df_sampling <- colData(se_gdt_subset) %>%
  as_data_frame() %>%
  group_by(tissue, clusters_per_lineage_annotated) %>%
  dplyr::count() %>%
  mutate(var=paste(tissue, clusters_per_lineage_annotated, sep="|")) %>%
  ungroup() %>%
  dplyr::select(n, var)
```

```{r gdt-metaneighbor-subset-100sampling-plot, fig.height=7, fig.width=7}
# pdf(
#   file = "./data/figures/chapter_04/figs_unfinished/suppfigures/gdt_crosspecies_100randomcellsampling_mean_auroc_heatmap.pdf",
#   width = 5.5,
#   height = 5.5
#   )
# heatmap
heatmap.2(
  mtn_mean,
  trace = "none",
  # superimpose a density histogram on color key
  density.info = "none",
  # color scale
  col = rev(colorRampPalette(brewer.pal(11, "RdYlBu"))(100)),
  breaks = seq(0, 1, length = 101),
  key.xlab = "AUROC",
  key.title = "",
  # text labels
  main = "Thymic vs PBMC GDT",
  cexRow = 0.6,
  cexCol = 0.6,
  # margins
  margins = c(7, 7)
)
# dev.off()
```

```{r,fig.width=10, fig.height=10}
# final df for plotting
mtn.df_sampling <- melt(mtn_mean) %>%
  dplyr::rename(auroc=value) %>%
  left_join(ncells.df_sampling, by=c("Var1"="var")) %>%
  dplyr::rename(Var1_ncells=n) %>%
  left_join(ncells.df_sampling, by=c("Var2"="var")) %>%
  dplyr::rename(Var2_ncells=n) %>%
  mutate(Var1=factor(Var1, levels=gdt_mtn_order),
         Var2=factor(Var2, levels=gdt_mtn_order))
  
# barplots
bp.x_sampling <- ggplot(data=mtn.df_sampling%>% select(Var1,Var1_ncells) %>% distinct(),
               aes(x=Var1, y=Var1_ncells))+
  geom_bar(stat="identity", fill="#bdbdbd") + theme_cowplot()+
  scale_x_discrete(position="top")+
  labs(y="# cells")+
  theme(axis.text = element_text(size=8),
        # axis.text.x = element_text(size = 8, angle=45, hjust=0),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.x = element_blank(),
        axis.line.x=element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y=element_text(size=8),
        legend.position = "none")

bp.y_sampling <- ggplot(data=mtn.df_sampling%>% select(Var2,Var2_ncells) %>% distinct(),
               aes(x=Var2, y=Var2_ncells))+
  geom_bar(stat="identity", fill="#bdbdbd") +
  scale_x_discrete(position="top") +
  labs(y="# cells")+ coord_flip() + theme_cowplot()+
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size=8),
        axis.text.y=element_blank(),
        axis.ticks.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_text(size=8),
        axis.line.y=element_blank(),
        legend.position = "none")


# BUBBLE PLOT
hm.clean_sampling <- ggplot(mtn.df_sampling, aes(x=Var1, y=Var2)) +
  geom_point(aes(size = abs(auroc-0.5), color= auroc))+
  geom_text(data=mtn.df_sampling %>% filter(auroc>0.8 | auroc < 0.1) %>% mutate(across("auroc", \(x) round(x,2))), aes(label=auroc), color="white", size=3)+
  scale_size_continuous(limits=c(0,1), breaks=seq(0,1, by=0.2), range = c(0, 20), guide="none")+
  scale_color_gradient2(low="#2166ac", mid="white", high="#a50f15", midpoint=0.5, limits=c(0,1), name="AUROC", breaks=seq(0,1, by=0.2))+
  labs(x="",y="", size="AUROC")+
  theme_cowplot()+
  theme(legend.position="bottom",
        legend.key.width = unit(0.8, 'cm'),
        axis.text = element_text(size=8),
        axis.title=element_text(size=8),
        axis.text.x=element_text(angle=45, hjust=1)
  )

(bp.x_sampling+plot_spacer() + plot_layout(widths = c(5, 1))) / (hm.clean_sampling + bp.y_sampling + plot_layout(widths = c(5, 1))) + plot_layout(heights = c(1, 5))

# ggsave(
#   # plot = hm.clean,
#   filename = "~/Projects/phd/data/figures/chapter_03/figs_unfinished/section1/ch3_fig8_gdt_mtn_thymuspbmc_equalnbcells_1000samples.pdf",
#   width = 9,
#   height = 9
# )
```




# SESSION INFO
```{r}
sessionInfo()
```


