---
title: "Chapter 3 - iNKT blood figure"
author: "Salomé Carcy"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true

---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 5, fig.height = 5,
                      warning=FALSE, message=FALSE,
                      root.dir = "~/Projects/phd/")
```


# IMPORT

## Import librairies
```{r import-librairies}
library(ggplot2)
library(RColorBrewer)
library(cowplot)
library(tidyverse)
library(dplyr)
library(Seurat)
library(SCpubr)
library(patchwork)
library(SummarizedExperiment)
library(MetaNeighbor)
library(patchwork)

source("~/Projects/phd/scripts/colors_universal.R")
```

## Import data
```{r import-data}
seur_integrated <- readRDS("./data/seurat_objects/seurat_human_integrated_object_23_12_01.rds")
DimPlot(seur_integrated, reduction="umap_integrated", cols = cols_integrated)

seur_pbmc_nkt <- readRDS("./data/seurat_objects/pbmc_objects/seurat_pbmc_nkt_clean.rds")
```




# FUNCTIONS
```{r define-functions}
lineage_dimplot <- function(seurobj, col_vector, file_name="no"){
  p <- do_DimPlot(
    seurobj,
    group.by="clusters_per_lineage",
    colors.use = col_vector,
    legend.position = "right",
    # raster = T,
    # raster.dpi=2048,
    pt.size = 0.5#,
    # border.size=1.5
  )+
    theme(
      # axis.title=element_text(size=8),
      # axis.text=element_text(size=6),
      legend.key.size = unit(0.4, "cm"),
      legend.title=element_text(size=8),
      legend.text = element_text(size=6)
    )
  if(file_name != "no"){
    ggsave(paste0("~/Projects/phd/data/figures/chapter_03/figs_unfinished/section1/", file_name),
           plot=p,
           width=4, height=2.5, units="in")
  }
}


lineage_clustersperdonor <- function(seurobj, col_vector, file_name="no", figwidth=6){
  p <- seurobj@meta.data %>%
    group_by(Donor, clusters_per_lineage) %>%
    dplyr::count() %>%
    group_by(Donor) %>%
    mutate(sum_per_donor=sum(n),
           freq=n*100/sum_per_donor) %>%
    mutate(Donor=factor(Donor, levels=c(8, 9, 10, 12, 13, 5, 11, 6, 7))) %>%
  ggplot(aes(x=factor(Donor, levels=sort(unique(Donor))), y=freq, fill=clusters_per_lineage))+
    geom_bar(stat="identity")+
    scale_fill_manual(values=col_vector)+
    labs(x="Donor", y="% cells", fill="")+
    theme_cowplot()#+
  # theme(axis.text=element_text(size=5/.pt),
  #       axis.title=element_text(size=7/.pt),
  #       legend.text = element_text(size=5/.pt))
  if(file_name != "no"){
    ggsave(paste0("~/Projects/phd/data/figures/chapter_03/figs_unfinished/section1/", file_name),
           plot=p,
           width=figwidth, height=5, units="in")
  }
}


plot_DotPlot <- function(seurobj, group, features, scaling=T){
  # get plot data
  p <- Seurat::DotPlot(
    seurobj,
    group.by=group,
    features=features,
    scale=scaling
  )
  # plot in personalized way
  p <- ggplot(p$data, aes(x=id, y=features.plot, fill=avg.exp.scaled, size=pct.exp))+
    geom_point(color="black", shape=21)+
    # scale_fill_gradient2(low=scales::muted("blue"), high=scales::muted("red"), name="z-score\nnormalized\navg expression")+
    scale_size_continuous(range=c(0,6), limits=c(0,100), name="%cells\nexpressing\ngene")+
    theme_bw()+
    theme(
      axis.text.y=element_text(face="italic"),
      axis.text.x=element_text(angle=45, hjust=1)
    )+
    labs(y="", x="")
  # different color scale if scaled or not
  if(scaling==T){
    p <- p + scale_fill_gradient2(low=scales::muted("blue"), high=scales::muted("red"), name="z-score\nnormalized\navg expression")
  } else{
    p <- p + viridis::scale_fill_viridis(option="B", direction=-1, name="normalized\navg expression")
  }
  return(p)
}

lineage_nebulosagenes <- function(seuratobj, genes_vector, pgrid_ncol=2, pgrid_size=c(10,30), file_name="no"){
  
  pnebulosa <- list()
  for(gene in genes_vector) {
    p <- SCpubr::do_NebulosaPlot(seuratobj, features = gene, sequential.palette="YlOrRd")
    p <- ggrastr::rasterise(p, layers="Point", dpi=200)
    pnebulosa[[gene]] <- p
  }
  pgrid <- plot_grid(plotlist=pnebulosa, ncol=pgrid_ncol)
  print(pgrid)
  
  if(file_name != "no"){
    ggsave(paste0("~/Projects/phd/data/figures/chapter_03/figs_unfinished/section1/", file_name),
           plot=pgrid,
           width=pgrid_size[1], height=pgrid_size[2])
  }
}
```




# UMAP and distribution per donor

## Umap
```{r umap-nkt}
lineage_dimplot(seurobj = seur_pbmc_nkt, col_vector=cols_pbmc_nkt, file_name="ch3_fig4_nkt_umap.pdf")
```


## Distribution per donor
```{r barplot_per_donor}
# cluster comp per donor
lineage_clustersperdonor(seurobj = seur_pbmc_nkt, col_vector = cols_pbmc_nkt, file_name = "ch3_fig4_nkt_clustersperdonor.pdf")
```




# GENES OF INTEREST
## Dotplot genes on interest
The marker genes were chosen based on:
- [Sallusto et al. 1999](https://www.nature.com/articles/44385)
- [Terekhova et al. 2023](https://www.sciencedirect.com/science/article/pii/S1074761323004533)
- [Rose et al. 2023](https://www.nature.com/articles/s42003-023-04747-9)
```{r dotplot-marker-genes, fig.height=5, message=F, warning=F}
genes_dotplot <- c(
  "ZBTB16",
  # ------
  # Tnaive
  "CCR7",
  "SELL",
  "TCF7",
  "LEF1",
  "CD27",
  # ------
  # CD4+ subset
  "CD4",
  "IL2RA",
  "AHR",
  "IL4R",
  "CCR4",
  "IL7R",
  # ------
  # AP1 signaling
  # "JUNB",
  # "JUN",
  # "FOS",
  # "CD69",
  # ------
  # CD4- subset
  "CD8A",
  "CD8B",
  "TBX21",
  "EOMES",
  "IFNG",
  "IL2RB",
  "IL15RA",
  "GZMK",
  "RORC",
  "RORA",
  "CCR6",
  # "IL23R",
  "KLRB1",
  "KLRG1",
  # "IL2RA",
  # ------
  # Temra
  "PRF1",
  "GNLY",
  "CX3CR1",
  "GZMB",
  # "TYROBP",
  "NCAM1", # CD56
  # -------
  # IL17 producing capacity
  "TGFBR1",
  "TGFBR2",
  "IL1R1",
  "IL1RAP",
  "IL23R"
  )

# get list of cell id to plot (remove clusters with less than 100 cells)
plot_DotPlot(
  seurobj=seur_pbmc_nkt,
  group="clusters_per_lineage",
  features=genes_dotplot,
  scaling=F
  )
# ggsave("~/Projects/phd/data/figures/chapter_03/figs_unfinished/section1/ch3_fig4_nkt_markergenes.pdf", width=5, height=7)
```


## Nebulosa plot genes of interest
```{r, fig.height=8, fig.width=6}
lineage_nebulosagenes(
  seur_pbmc_nkt,
  genes_vector=c("CCR7", "SELL", "ZBTB16", "KLRB1", "GZMK", "GZMB"),
  # file_name="ch3_fig4_nkt_nebulosaplots.pdf",
  pgrid_ncol=2,
  pgrid_size=c(10, 15)
)
```




# METANEIGHBOR

## Metaneighbor thymocytes x pbmcs

Prepare data.
```{r prepare-nkt-mtn}
# Get only NKTs
seur_nkt <- subset(seur_integrated, subset=tcell_lineage=="iNKT")
table(seur_nkt@meta.data[,c("clusters_per_lineage", "donor_id")], useNA="ifany") # sanity check

# rename clusters_per_lineage for easier interpretation of metaneighbor
seur_nkt@meta.data$clusters_per_lineage_annotated <- case_when(
  seur_nkt@meta.data$clusters_per_lineage=="iNKT_thymus_c0" ~ "iNKT CD8aa (c0)",
  seur_nkt@meta.data$clusters_per_lineage=="iNKT_thymus_c1" ~ "iNKT CCR9+ (c1)",
  seur_nkt@meta.data$clusters_per_lineage=="iNKT_thymus_c2" ~ "iNKT CCR7+ (c2)",
  seur_nkt@meta.data$clusters_per_lineage=="iNKT_thymus_c3" ~ "iNKT CCR9+ (c3)",
  seur_nkt@meta.data$clusters_per_lineage=="iNKT_thymus_c4" ~ "iNKT CD4+ mem (c4)",
  seur_nkt@meta.data$clusters_per_lineage=="iNKT_thymus_c5" ~ "iNKT CD4+ mem (c5)",
  seur_nkt@meta.data$clusters_per_lineage=="iNKT_thymus_c6" ~ "iNKT CD4- mem (c6)",
  seur_nkt@meta.data$clusters_per_lineage=="iNKT_pbmc_c0" ~ "iNKT CD4+ naive (c0)",
  seur_nkt@meta.data$clusters_per_lineage=="iNKT_pbmc_c1" ~ "iNKT CD4+ (c1)",
  seur_nkt@meta.data$clusters_per_lineage=="iNKT_pbmc_c2" ~ "iNKT CD4- (c2)",
  seur_nkt@meta.data$clusters_per_lineage=="iNKT_pbmc_c3" ~ "iNKT CD4- Temra (c3)"
)
# table(seur_nkt@meta.data[,c("clusters_per_lineage_annotated", "clusters_per_lineage")], useNA="ifany")

# get count matrix
se_nkt <- SummarizedExperiment(
  assays=seur_nkt@assays[["RNA"]]@counts,
  colData=seur_nkt@meta.data[,c("clusters_per_lineage_annotated", "donor_id", "tissue")]
  )

# get NKT-specific HVGs
# seur_thym_nkt <- readRDS("./data/seurat_objects/thymus_objects/seurat_filtered_harmony_02_15_23_thymus.nkt.RDS")
# seur_pbmc_nkt <- readRDS("./data/seurat_objects/pbmc_objects/blood.NKT_03_16_23.RDS")
# hvg_nkt_joint <- union(VariableFeatures(seur_thym_nkt), VariableFeatures(seur_pbmc_nkt))
hvg_nkt <- VariableFeatures(FindVariableFeatures(seur_nkt))
length(intersect(VariableFeatures(seur_integrated), hvg_nkt)) # 864 genes
# length(intersect(hvg_nkt, hvg_nkt_joint))
```

Run metaneighbor with donors as "study" (to check for reproducibility across donors).
```{r run_mtn_branches, fig.width=14, fig.height=14}
mtn <- MetaNeighborUS(
  # var_genes=VariableFeatures(seur_integrated),
  var_genes=hvg_nkt,
  dat=se_nkt,
  study_id=se_nkt$donor_id,
  cell_type=se_nkt$clusters_per_lineage_annotated,
  fast_version=T
  )

# quick heatmap
heatmap.2(mtn,
          # trace
          trace="none",
          # dendrogram
          # Rowv=FALSE,
          # Colv=FALSE,
          # dendrogram="none",
          # superimpose a density histogram on color key
          density.info="none",
          # color scale
          col=rev(colorRampPalette(brewer.pal(11,"RdYlBu"))(100)),
          breaks=seq(0,1,length=101),
          key.xlab = "AUROC",
          key.title="",
          keysize = 1.2,
          # text labels
          main="",
          cexRow=0.6,
          cexCol=0.6,
          # margins
          margins=c(9,9))
```
To note: the PBMC Temra (c3) between donors 11 and (5, 6, 7) don't have very high AUROCs, but at the same time donor 11 has the most Temra (the other donors have very very few Temra cells), so probably an issue of sampling (cell number).

Now re-run it with "tissue" as study_id (will be easier to read).
```{r, fig.width=8, fig.height=8}
mtn2 <- MetaNeighborUS(
  # var_genes=VariableFeatures(seur_integrated),
  var_genes=hvg_nkt,
  dat=se_nkt,
  study_id=se_nkt$tissue,
  cell_type=se_nkt$clusters_per_lineage_annotated,
  fast_version=T
  )

# quick heatmap
heatmap.2(mtn2,
          # trace
          trace="none",
          # dendrogram
          # Rowv=FALSE,
          # Colv=FALSE,
          # dendrogram="none",
          # superimpose a density histogram on color key
          density.info="none",
          # color scale
          col=rev(colorRampPalette(brewer.pal(11,"RdYlBu"))(100)),
          breaks=seq(0,1,length=101),
          key.xlab = "AUROC",
          key.title="",
          # keysize = 0.5,
          # text labels
          main="",
          cexRow=0.6,
          cexCol=0.6,
          # margins
          margins=c(10,10)
          )
```




```{r plot-nkt-mtn, fig.width=10, fig.height=10}
library(reshape2)
# library(ggrepel)
nkt_mtn_order <- c(
  "Thymus|iNKT CD8aa (c0)",
  "Thymus|iNKT CCR9+ (c1)",
  "Thymus|iNKT CCR9+ (c3)",
  "Thymus|iNKT CCR7+ (c2)",
  "PBMC|iNKT naive (c0)",
  "Thymus|iNKT CD4+ eff (c4)",
  "Thymus|iNKT CD4+ eff (c5)",
  "PBMC|iNKT Tcm (c1)",
  "Thymus|iNKT CD4- eff (c6)",
  "PBMC|iNKT Tem (c2)",
  "PBMC|iNKT Temra (c3)"
)

# get nb of cells
ncells.df <- seur_nkt@meta.data %>%
  group_by(tissue, clusters_per_lineage_annotated) %>%
  dplyr::count() %>%
  mutate(var=paste(tissue, clusters_per_lineage_annotated, sep="|")) %>%
  ungroup() %>%
  dplyr::select(n, var)

# final df for plotting
mtn.df <- melt(mtn2) %>%
  dplyr::rename(auroc=value) %>%
  left_join(ncells.df, by=c("Var1"="var")) %>%
  dplyr::rename(Var1_ncells=n) %>%
  left_join(ncells.df, by=c("Var2"="var")) %>%
  dplyr::rename(Var2_ncells=n) %>%
  mutate(Var1=factor(Var1, levels=nkt_mtn_order),
         Var2=factor(Var2, levels=nkt_mtn_order))
  
# barplots
# bp.x <- ggplot(data=mtn.df%>% select(Var1,Var1_ncells) %>% distinct(),
#                aes(x=Var1, y=Var1_ncells))+
#   geom_bar(stat="identity", fill="#bdbdbd") + theme_cowplot()+
#   scale_x_discrete(position="top")+
#   labs(y="# cells")+
#   theme(axis.text = element_text(size=8),
#         # axis.text.x = element_text(size = 8, angle=45, hjust=0),
#         axis.text.x=element_blank(),
#         axis.ticks.x=element_blank(),
#         axis.title.x = element_blank(),
#         axis.line.x=element_blank(),
#         axis.ticks.y = element_blank(),
#         axis.title.y=element_text(size=8),
#         legend.position = "none")
# 
# bp.y <- ggplot(data=mtn.df%>% select(Var2,Var2_ncells) %>% distinct(),
#                aes(x=Var2, y=Var2_ncells))+
#   geom_bar(stat="identity", fill="#bdbdbd") +
#   scale_x_discrete(position="top") +
#   labs(y="# cells")+ coord_flip() + theme_cowplot()+
#   theme(axis.title.y = element_blank(),
#         axis.text = element_text(size=8),
#         axis.text.y=element_blank(),
#         axis.ticks.y = element_blank(),
#         axis.ticks.x = element_blank(),
#         axis.title.x = element_text(size=8),
#         axis.line.y=element_blank(),
#         legend.position = "none")


# BUBBLE PLOT
hm.clean <- ggplot(mtn.df, aes(x=Var1, y=Var2)) +
  geom_point(aes(size = abs(auroc-0.5), color= auroc))+
  geom_text(data=mtn.df %>% filter(auroc>0.8 | auroc < 0.1) %>% mutate(across("auroc", \(x) round(x,2))), aes(label=auroc), color="white", size=3)+
  scale_size_continuous(limits=c(0,1), breaks=seq(0,1, by=0.2), range = c(0, 20), guide="none")+
  scale_color_gradient2(low="#2166ac", mid="white", high="#a50f15", midpoint=0.5, limits=c(0,1), name="AUROC", breaks=seq(0,1, by=0.2))+
  labs(x="",y="", size="AUROC")+
  theme_cowplot()+
  theme(legend.position="bottom",
        legend.key.width = unit(0.8, 'cm'),
        axis.text = element_text(size=8),
        axis.title=element_text(size=8),
        axis.text.x=element_text(angle=45, hjust=1)
  )
(bp.x+plot_spacer() + plot_layout(widths = c(5, 1))) / (hm.clean + bp.y + plot_layout(widths = c(5, 1))) + plot_layout(heights = c(1, 5))
# ggsave("~/Projects/phd/data/figures/chapter_03/figs_unfinished/section1/ch3_fig4_inkt_mtn_thymuspbmc.pdf",
#        width=8, height=8)
```



# SESSION INFO
```{r}
sessionInfo()
```


